---
title: "S1_Bayesian_Hierarchical_Analysis"
output: html_document
---

---
title: "Study 1 Bayesian Hierarchical Regression"

author: "Yu Liu"

date: "11th May 2026"

output:

  html_document:

    toc: true

    toc_float: true

    number_sections: true

---

# Setup

```{r setup, message=FALSE, warning=FALSE}
#install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
#cmdstanr::install_cmdstan()
#options(brms.backend = "cmdstanr")
library(tidyverse)

library(brms)

library(bayesplot)

library(tidybayes)

library(emmeans)

library(patchwork)

library(loo)

theme_set(theme_bw())

options(mc.cores = parallel::detectCores())

set.seed(1234)

```

# Load Data

```{r load-data}
d <- read_csv("S1_smallN_model_ready.csv")
glimpse(d)
summary(d)
```

# Factor Preparation

```{r factors}
d <- d %>%
  mutate(
    ID = factor(ID),
    CueType = factor(CueType,
      levels = c("NR", "R")),
    DurationF = factor(Duration,
      levels = c(
        "50ms",
        "100ms",
        "150ms",
        "200ms",
        "250ms",
        "300ms",
        "350ms")),
    Grouping = factor(Grouping),
    SessionF = factor(Session),
    SessionPhase = factor(
      SessionPhase,
      levels = c("Early", "Middle", "Late")
    ),
    logRT = log(RT)
  )

levels(d$DurationF)
levels(d$SessionF)
levels(d$CueType)
levels(d$Grouping)
```

# Initial Data Checks

## Trial Counts

```{r counts}
d %>%
  count(ID, CueType, DurationF)
```

## Signed Error Bias Check
Maximum response bias was 3.73deg (NR-cue at 150ms Encoding Duration), in comparison to SD (18.31-31.79 for R-cue, 49.67-58.49 for NR-cue), biases are tiny. Thus we do not consider the bias factor, and use abs error deg.

```{r signed-bias}
d %>%
  group_by(CueType, DurationF) %>%
  summarise(
    mean_signed = mean(SignedErr, na.rm = TRUE),
    sd_signed = sd(SignedErr, na.rm = TRUE),
    .groups = "drop"
  )
```

# Descriptive Figures

## Absolute Error by Duration and Cue

```{r descriptives-participant-cirSd-duration}
library(circular)

circSD_deg <- function(x) {
  circular::sd.circular(
    circular(x, units = "degrees")
  ) * (180/pi)
}

circ_summary <- d %>%
  group_by(ID, CueType, DurationF) %>%
  summarise(
    circSD = circSD_deg(SignedErr),
    meanAbs = mean(AbsErr),
    .groups = "drop"
  )

ggplot(circ_summary,
       aes(DurationF, circSD,
           colour = CueType,
           group = CueType)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  facet_wrap(~ID) +
  labs(
    y = "Circular SD (deg)",
    x = "Exposure Duration"
  ) +
  theme_bw()
```
```{r group-cirSD-Duration} 
group_circ <- d %>%
  group_by(CueType, DurationF) %>%
  summarise(
    circSD = circSD_deg(SignedErr),
    .groups = "drop"
  )

ggplot(group_circ,
       aes(DurationF, circSD,
           colour = CueType,
           group = CueType)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  labs(
    y = "Circular SD (deg)",
    x = "Exposure Duration"
  )
```

# Redundancy Gain Function
This plot isolates redundancy gain: Redundancy Gain = (NR error - R error)
Theoretically important because the primary research question concerns whether 
redundancy gain changes as exposure duration increases, rather than just whether overall performance improves with duration.
Encoding-rate/perceptual accounts predict that non-redundant items should 
gradually "catch up" with increased exposure duration, producing a decreasing 
redundancy-gain function
In contrast, mnemonic/post-encoding accounts predict that redundancy
gain should emerge rapidly and remain relatively stable across durations.

Thus, the shape of the redundancy-gain function directly visualizes
whether redundancy effects themselves are exposure-duration dependent.

Current descriptive pattern:

# NOTE:
This figure is descriptive only. Bayesian hierarchical models are required to 
formally assess the magnitude and uncertainty of the Cue × Duration interaction.

```{r redundancy-gain}
gain_d <- d %>%
  group_by(ID, DurationF, CueType) %>%
  summarise(
    mean_abs = mean(AbsErr),
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = CueType,
    values_from = mean_abs
  ) %>%
  mutate(
    RedundancyGain = NR - R
  )
ggplot(gain_d,
       aes(DurationF, RedundancyGain,
           group = ID)) +
  geom_line(alpha = .5) +
  geom_point(alpha = .7) +
  stat_summary(
    aes(group = 1),
    fun = mean,
    geom = "line",
    linewidth = 1.5,
    colour = "black"
  ) +
  stat_summary(
    aes(group = 1),
    fun = mean,
    geom = "point",
    size = 3,
    colour = "black"
  ) +
  labs(
    y = "Redundancy Gain (NR - R)",
    x = "Exposure Duration"
  )
```

# INITIAL BAYESIAN HIERARCHICAL PRECISION MODEL
Goal:
First inferential test of whether redundancy gain depends on exposure duration.

DV
- Absolute circular response error (AbsErr)

Fixed effects:
- CueType: tests overall redundancy gain (R vs NR)
- DurationF: tests whether overall precision changes across exposure durations
- CueType × DurationF: critical interaction term testing whether redundancy gain 
changes as a function of exposure duration
- SessionF: included as a fixed-effect control factor to account for 
practice/learning effects across sessions

Random effects:
- varying intercepts by participant
- varying CueType slopes by participant
Random-effects structure allows:
- baseline precision differences between participants
- participant-specific redundancy gain magnitude

Distributional family:
- Student-t likelihood was chosen because trial-level response-error 
distributions contain heavy tails and occasional large-error trials, which 
violate normality assumptions.

Current modelling rationale:

- Duration and Session are initially treated as categorical factors to avoid 
imposing assumptions about monotonicity or functional form.

Further analyses:
- continuous duration/session models
- logarithmic/asymptotic duration functions
- model comparison (categorical vs continuous formulations)
- practical magnitude assessment of Cue × Duration interaction

Theoretically:
- encoding-rate/perceptual accounts predict duration-dependent redundancy gain
- mnemonic/post-encoding accounts predict relatively stable redundancy gain 
across durations
Thus, CueType × DurationF is the primary theoretical parameter of interest.

```{r model1, message=FALSE}
m1 <- brm(
  AbsErr ~ CueType * DurationF + SessionF +
    (1 + CueType | ID),
  data = d,
  family = student(),
  prior = c(
    prior(normal(40, 20), class = Intercept),
    prior(normal(0, 10), class = b),
    prior(exponential(1), class = sigma),
    prior(exponential(1), class = sd)
  ),
  chains = 4,
  cores = 4,
  iter = 4000,
  warmup = 2000,
  control = list(
    adapt_delta = .95,
    max_treedepth = 15
  ),
  seed = 1234
)
```

# Model Summary (seed 1234)
 Family: student 
  Links: mu = identity 
Formula: AbsErr ~ CueType * DurationF + SessionF + (1 + CueType | ID) 
   Data: d (Number of observations: 12599) 
  Draws: 4 chains, each with iter = 4000; warmup = 2000; thin = 1;
         total post-warmup draws = 8000

Multilevel Hyperparameters:
~ID (Number of levels: 3) 
                        Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
sd(Intercept)               2.07      0.85     0.95     4.22 1.00     5399     5513
sd(CueTypeR)                1.68      0.78     0.66     3.65 1.00     4774     5356
cor(Intercept,CueTypeR)    -0.12      0.39    -0.78     0.64 1.00     5005     5291

Regression Coefficients:
                        Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
Intercept                  18.02      1.44    15.26    20.97 1.00     3038     4063
CueTypeR                   -6.97      1.28    -9.42    -4.32 1.00     2949     3974
DurationF100ms              1.09      0.83    -0.55     2.70 1.00     3094     4615
DurationF150ms             -2.38      0.83    -3.99    -0.76 1.00     3070     4613
DurationF200ms             -2.01      0.80    -3.56    -0.44 1.00     2883     4493
DurationF250ms             -1.66      0.78    -3.20    -0.12 1.00     3014     4723
DurationF300ms             -2.90      0.74    -4.35    -1.43 1.00     2770     4118
DurationF350ms             -2.46      0.77    -3.94    -0.96 1.00     2823     4334
SessionF2                   0.03      0.52    -0.98     1.02 1.00     4048     5713
SessionF3                  -0.96      0.49    -1.91    -0.02 1.00     3720     5482
SessionF4                  -1.10      0.48    -2.05    -0.18 1.00     3682     5628
SessionF5                  -1.15      0.49    -2.11    -0.19 1.00     3747     5029
SessionF6                  -1.27      0.48    -2.20    -0.31 1.00     3659     5577
SessionF7                  -0.44      0.49    -1.40     0.51 1.00     3620     5043
SessionF8                  -1.14      0.49    -2.09    -0.17 1.00     3703     5314
SessionF9                  -1.46      0.49    -2.42    -0.49 1.00     3679     5214
SessionF10                 -1.51      0.48    -2.46    -0.57 1.00     3863     5762
CueTypeR:DurationF100ms    -1.88      0.96    -3.74     0.01 1.00     3198     4557
CueTypeR:DurationF150ms     1.59      0.96    -0.28     3.44 1.00     3169     5032
CueTypeR:DurationF200ms     0.68      0.93    -1.12     2.49 1.00     2927     4508
CueTypeR:DurationF250ms    -0.10      0.90    -1.85     1.66 1.00     2873     4608
CueTypeR:DurationF300ms     1.74      0.87     0.02     3.46 1.00     2834     4754
CueTypeR:DurationF350ms     1.56      0.90    -0.21     3.35 1.00     2885     4588

Further Distributional Parameters:
      Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
sigma     8.52      0.13     8.27     8.77 1.00     7641     6496
nu        1.20      0.02     1.16     1.24 1.00     8467     5890

Draws were sampled using sample(hmc). For each parameter, Bulk_ESS
and Tail_ESS are effective sample size measures, and Rhat is the potential
scale reduction factor on split chains (at convergence, Rhat = 1).


```{r model1-summary}
m1 <- readRDS("m1_categorical_sessionfactor.rds")
summary(m1)
#saveRDS(m1, "m1_categorical_sessionfactor.rds")
#save.image("S1_analysis_workspace.RData")
```

# Posterior Diagnostics

```{r diagnostics}
plot(m1, variable = "^b_", regex = TRUE)
plot(m1, variable = "^sd_", regex = TRUE)
pp_check(m1, ndraws = 100)
```

# Conditional Effects

## Cue × Duration

```{r cond-effects}
conditional_effects(
  m1,
  effects = "DurationF:CueType"
)
```

# Posterior Estimates

```{r posterior-estimates}
fixef(m1)
```

# Model 2: Interaction Learning Model

Tests whether practice changes the Cue × Duration relation.

```{r model2, message=FALSE}

m2 <- brm(

  AbsErr ~ CueType * Duration * Session_z +

    (1 + CueType + Session_z | ID),

  data = d,

  family = student(),

  prior = c(

    prior(normal(40, 20), class = Intercept),

    prior(normal(0, 10), class = b),

    prior(exponential(1), class = sigma),

    prior(exponential(1), class = sd)

  ),

  chains = 4,

  cores = 4,

  iter = 4000,

  warmup = 2000,

  control = list(

    adapt_delta = .95,

    max_treedepth = 15

  ),

  seed = 5678

)

```

# Model 2 Summary

```{r model2-summary}

summary(m2)

```

# Model Comparison

```{r model-comparison}

loo_m1 <- loo(m1)

loo_m2 <- loo(m2)

loo_compare(loo_m1, loo_m2)

```

# Estimated Marginal Means

```{r emmeans}

emm <- emmeans(

  m1,

  ~ CueType * Duration

)

emm

```

# Posterior Redundancy Gain Estimates

```{r gain-estimates}

gain_emm <- contrast(

  emm,

  method = "revpairwise",

  by = "Duration"

)

gain_emm

```

# Session Learning Visualization

```{r session-learning}

ggplot(d,

       aes(Session, AbsErr,

           colour = CueType)) +

  geom_smooth(method = "loess", se = TRUE) +

  facet_wrap(~ID) +

  labs(

    y = "Absolute Error (deg)"

  )

```

# RT Model

Secondary analysis.

```{r rt-model, message=FALSE}

m_rt <- brm(

  logRT ~ CueType * Duration + Session_z +

    (1 + CueType | ID),

  data = d,

  family = gaussian(),

  prior = c(

    prior(normal(6, 1), class = Intercept),

    prior(normal(0, .2), class = b),

    prior(exponential(1), class = sigma),

    prior(exponential(1), class = sd)

  ),

  chains = 4,

  cores = 4,

  iter = 4000,

  control = list(

    adapt_delta = .95

  ),

  seed = 2468

)

```

# RT Summary

```{r rt-summary}

summary(m_rt)

```

# Save Models

```{r save-models}

saveRDS(m1, "m1_precision_main.rds")

saveRDS(m2, "m2_precision_learning.rds")

saveRDS(m_rt, "m_rt.rds")

```