---
title: "Study 1 Bayesian Hierarchical Regression"
author: "Yu Liu"
date: "11th May 2026"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

# Overview

Analysis order (small-N, each participant = one replication):

1. Load and prepare trial data
2. Simple descriptive plots (placeholder for manuscript figures)
3. Bayesian AbsErr primary models + theory contrasts
4. Bayesian cirSD primary (same design formula; bootstrap cell replicates)
5. Pre-mixture learning: AbsErr Session 1 vs 10 by duration
6. Pre-mixture learning: cirSD Session 1 vs 10 (nonparametric bootstrap companion)
7. Circular mixture: independent ML fit per cue × duration cell
   (signed-error distribution; no separate cirSD mixture step)

Appendix (below): frequentist sensitivity and older exploratory Bayes code.

Style note: code is written out step by step with comments. We avoid custom
helper functions so every scientific step is visible. Short `for` loops are used
only to repeat the same steps for AQ, HC, and YILIU.

Maps to `s1_smallN_analysis_discrete.Rmd` (SessionF instead of Session_z).

# Setup

```{r setup, message=FALSE, warning=FALSE}
# Core packages for this pipeline.
library(tidyverse)   # data wrangling + ggplot2
library(nlme)        # frequentist mixed models (Appendix)
library(brms)        # Bayesian regression
library(bayesplot)   # posterior predictive checks
library(tidybayes)   # gather posterior draws from emmeans
library(emmeans)     # estimated marginal means / contrasts
library(patchwork)   # combine ggplot panels
library(loo)         # Appendix model comparison only
library(circular)    # circular SD for signed colour errors

theme_set(theme_bw())
options(mc.cores = parallel::detectCores())
set.seed(1234)  # reproducible bootstrap / Stan seeds where used
```

# 1. Load and prepare data

```{r load-data}
# Model-ready trial table from MATLAB preprocessing.
d <- read_csv("S1_smallN_model_ready.csv", show_col_types = FALSE)

# Quick sanity checks: expected columns and no empty file.
stopifnot(nrow(d) > 0)
stopifnot(all(c("ID", "CueType", "Duration", "Session", "Session_z",
                "AbsErr", "SignedErr") %in% names(d)))

glimpse(d)
```

```{r prepare-factors}
# Set factor levels explicitly so contrasts and plots keep a fixed order.
d <- d %>%
  mutate(
    ID = factor(ID),
    CueType = factor(CueType, levels = c("NR", "R")),
    DurationF = factor(
      Duration,
      levels = c("50ms", "100ms", "150ms", "200ms", "250ms", "300ms", "350ms")
    ),
    Grouping = factor(Grouping),
    SessionF = factor(Session),
    SessionPhase = factor(SessionPhase, levels = c("Early", "Middle", "Late")),
    logRT = log(RT)
  )

# One data frame per currently complete participant.
# droplevels() drops unused ID levels so random-effect formulas stay clean.
d_AQ <- droplevels(filter(d, ID == "AQ"))
d_HC <- droplevels(filter(d, ID == "HC"))
d_YILIU <- droplevels(filter(d, ID == "YILIU"))

# Confirm design levels.
levels(d$CueType)
levels(d$DurationF)
levels(d$SessionF)
```

# 2. Descriptive figures (placeholder)

These plots are for exploration only. Manuscript figures will be updated later.

```{r descriptives-cirsd}
# Circular SD of signed errors by participant × cue × duration.
# sd.circular returns radians when units="degrees"; convert to degrees.
circ_summary <- d %>%
  group_by(ID, CueType, DurationF) %>%
  summarise(
    circSD = as.numeric(sd.circular(circular(SignedErr, units = "degrees"))) * (180 / pi),
    meanAbs = mean(AbsErr),
    .groups = "drop"
  )

ggplot(circ_summary, aes(DurationF, circSD, colour = CueType, group = CueType)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  facet_wrap(~ID) +
  labs(y = "Circular SD (deg)", x = "Encoding duration")

# Same summary pooled across participants (descriptive only; not primary inference).
group_circ <- d %>%
  group_by(CueType, DurationF) %>%
  summarise(
    circSD = as.numeric(sd.circular(circular(SignedErr, units = "degrees"))) * (180 / pi),
    .groups = "drop"
  )

ggplot(group_circ, aes(DurationF, circSD, colour = CueType, group = CueType)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  labs(y = "Circular SD (deg)", x = "Encoding duration")
```


# 3. Bayesian primary analysis: AbsErr

One Student-t model per participant:

`AbsErr ~ CueType * DurationF * Session_z + (1 | SessionF)`

- Cue and duration are categorical (cell means).
- Session_z is continuous practice (z-scored session index).
- Random intercept by SessionF absorbs leftover session bumps.
- ROPE for AbsErr differences: ±3 degrees (substantive “near zero”).

## 3.1 Fit / load models

```{r bayes-fit, message=FALSE, warning=FALSE}
# Region of practical equivalence for AbsErr contrasts (degrees).
rope_abs <- 3

# Session_z landmarks used later in contrasts.
# Early = mean of sessions 1–2; late = mean of sessions 9–10; average = 0.
early_z <- mean(unique(d$Session_z[d$Session %in% c(1, 2)]))
late_z  <- mean(unique(d$Session_z[d$Session %in% c(9, 10)]))
s1_z    <- unique(d$Session_z[d$Session == 1])
s10_z   <- unique(d$Session_z[d$Session == 10])
stopifnot(length(s1_z) == 1, length(s10_z) == 1)

# Ordered-duration linear contrast weights (7 levels: 50 ... 350 ms).
# Negative AbsErr contrast = error falls as duration increases.
w_linear <- c(-3, -2, -1, 0, 1, 2, 3)
w_350_minus_50 <- c(-1, 0, 0, 0, 0, 0, 1)

# Mildly regularising priors (AbsErr is typically tens of degrees).
bayes_priors <- c(
  prior(normal(30, 20), class = "Intercept"),
  prior(normal(0, 15), class = "b"),
  prior(exponential(1), class = "sigma"),
  prior(exponential(1), class = "sd"),
  prior(gamma(2, 0.1), class = "nu")
)

dir.create("bayes_primary", showWarnings = FALSE, recursive = TRUE)

# Fit each participant explicitly. brms `file=` caches the Stan fit on disk,
# so re-knitting loads the saved model unless the formula/data change.
m_AQ <- brm(
  AbsErr ~ CueType * DurationF * Session_z + (1 | SessionF),
  data = d_AQ,
  family = student(),
  prior = bayes_priors,
  chains = 4, cores = 4, iter = 4000, warmup = 2000,
  control = list(adapt_delta = 0.95, max_treedepth = 12),
  seed = 1234,
  file = "bayes_primary/m_AQ_abserr",
  file_refit = "on_change"
)

m_HC <- brm(
  AbsErr ~ CueType * DurationF * Session_z + (1 | SessionF),
  data = d_HC,
  family = student(),
  prior = bayes_priors,
  chains = 4, cores = 4, iter = 4000, warmup = 2000,
  control = list(adapt_delta = 0.95, max_treedepth = 12),
  seed = 1234,
  file = "bayes_primary/m_HC_abserr",
  file_refit = "on_change"
)

m_YILIU <- brm(
  AbsErr ~ CueType * DurationF * Session_z + (1 | SessionF),
  data = d_YILIU,
  family = student(),
  prior = bayes_priors,
  chains = 4, cores = 4, iter = 4000, warmup = 2000,
  control = list(adapt_delta = 0.95, max_treedepth = 12),
  seed = 1234,
  file = "bayes_primary/m_YILIU_abserr",
  file_refit = "on_change"
)

summary(m_AQ)
summary(m_HC)
summary(m_YILIU)
```

## 3.2 Theory contrasts (written out)

Scientific questions:

1. Redundancy gain at 50 ms: AbsErr(NR) − AbsErr(R)  (positive = R better)
2. Duration change by cue: 350−50 and linear trend at average session
3. Catch-up: does NR improve more than R from 50→350?  Defined as (R−NR) of that change so positive = NR improved more
4. Practice: late−early change in the 50→350 duration effect (not change in redundancy gain)

```{r bayes-contrasts, message=FALSE, warning=FALSE}
# Store one summary row per effect × participant.
# Columns: ID, effect, median, lower, upper, p_lt_0, p_gt_0, p_in_rope
contrast_rows <- list()
row_i <- 0

# Repeat the same contrast workflow for each saved model.
models <- list(AQ = m_AQ, HC = m_HC, YILIU = m_YILIU)

for (pid in names(models)) {
  model <- models[[pid]]

  # ---- 1) Redundancy gain by duration at average session (Session_z = 0) ----
  emm_gain <- emmeans(model, ~ CueType | DurationF, at = list(Session_z = 0))
  ct_gain <- contrast(emm_gain, method = list("NR - R" = c(1, -1)), by = "DurationF")
  draws_gain <- gather_emmeans_draws(ct_gain)

  for (dur in levels(d$DurationF)) {
    v <- draws_gain$.value[draws_gain$DurationF == dur]
    v <- v[is.finite(v)]
    stopifnot(length(v) > 0)
    row_i <- row_i + 1
    contrast_rows[[row_i]] <- tibble(
      ID = pid,
      effect = paste0("gain_", dur),
      median = median(v),
      lower = as.numeric(quantile(v, 0.025)),
      upper = as.numeric(quantile(v, 0.975)),
      p_lt_0 = mean(v < 0),
      p_gt_0 = mean(v > 0),
      p_in_rope = mean(abs(v) < rope_abs)
    )
  }

  # Also keep the focal 50 ms gain under a short name.
  v <- draws_gain$.value[draws_gain$DurationF == "50ms"]
  v <- v[is.finite(v)]
  row_i <- row_i + 1
  contrast_rows[[row_i]] <- tibble(
    ID = pid,
    effect = "gain_at_50ms",
    median = median(v),
    lower = as.numeric(quantile(v, 0.025)),
    upper = as.numeric(quantile(v, 0.975)),
    p_lt_0 = mean(v < 0),
    p_gt_0 = mean(v > 0),
    p_in_rope = mean(abs(v) < rope_abs)
  )

  # ---- 2) Duration effects by cue at average session ----
  emm_dur <- emmeans(model, ~ DurationF | CueType, at = list(Session_z = 0))
  ct_dur <- contrast(
    emm_dur,
    method = list(
      "linear_trend" = w_linear,
      "350ms - 50ms" = w_350_minus_50
    ),
    by = "CueType"
  )
  draws_dur <- gather_emmeans_draws(ct_dur)

  for (cue in c("NR", "R")) {
    for (cname in c("linear_trend", "350ms - 50ms")) {
      v <- draws_dur$.value[draws_dur$CueType == cue & draws_dur$contrast == cname]
      v <- v[is.finite(v)]
      stopifnot(length(v) > 0)
      row_i <- row_i + 1
      contrast_rows[[row_i]] <- tibble(
        ID = pid,
        effect = paste0(cue, ": ", cname),
        median = median(v),
        lower = as.numeric(quantile(v, 0.025)),
        upper = as.numeric(quantile(v, 0.975)),
        p_lt_0 = mean(v < 0),
        p_gt_0 = mean(v > 0),
        p_in_rope = mean(abs(v) < rope_abs)
      )
    }
  }

  # ---- 3) Catch-up at early / average / late session slices ----
  # First get 350-50 change for each cue at each Session_z slice.
  emm_dur_sess <- emmeans(
    model,
    ~ DurationF | CueType * Session_z,
    at = list(Session_z = c(early_z, 0, late_z))
  )
  ct_dur_sess <- contrast(
    emm_dur_sess,
    method = list("350ms - 50ms" = w_350_minus_50),
    by = c("CueType", "Session_z")
  )
  # ungroup(): tidybayes keeps contrast/CueType/Session_z groups; leaving them
  # in place makes pivot_wider keep Session_z as a key so early/late never
  # share a row and late - early becomes all NA (length(v) > 0 fails).
  draws_ds <- gather_emmeans_draws(ct_dur_sess) %>%
    ungroup() %>%
    select(.draw, CueType, Session_z, .value)

  # Label Session_z values as early / average / late (sorted numerically).
  sz_levels <- sort(unique(draws_ds$Session_z))
  stopifnot(length(sz_levels) == 3)
  draws_ds <- draws_ds %>%
    mutate(SessionSlice = c("early", "average", "late")[match(Session_z, sz_levels)])

  # Catch-up draw = duration_change_R - duration_change_NR
  # Positive => NR AbsErr fell more from 50 to 350 than R did.
  for (slice in c("early", "average", "late")) {
    wide <- draws_ds %>%
      filter(SessionSlice == slice) %>%
      select(.draw, CueType, .value) %>%
      pivot_wider(names_from = CueType, values_from = .value)
    v <- wide$R - wide$NR
    v <- v[is.finite(v)]
    stopifnot(length(v) > 0)
    row_i <- row_i + 1
    contrast_rows[[row_i]] <- tibble(
      ID = pid,
      effect = paste0("catchup_", slice),
      median = median(v),
      lower = as.numeric(quantile(v, 0.025)),
      upper = as.numeric(quantile(v, 0.975)),
      p_lt_0 = mean(v < 0),
      p_gt_0 = mean(v > 0),
      p_in_rope = mean(abs(v) < rope_abs)
    )
  }

  # Catch-up change with practice: late catch-up minus early catch-up.
  wide_e <- draws_ds %>%
    filter(SessionSlice == "early") %>%
    select(.draw, CueType, .value) %>%
    pivot_wider(names_from = CueType, values_from = .value) %>%
    mutate(catchup = R - NR)
  wide_l <- draws_ds %>%
    filter(SessionSlice == "late") %>%
    select(.draw, CueType, .value) %>%
    pivot_wider(names_from = CueType, values_from = .value) %>%
    mutate(catchup = R - NR)
  v <- wide_l$catchup[match(wide_e$.draw, wide_l$.draw)] - wide_e$catchup
  v <- v[is.finite(v)]
  stopifnot(length(v) > 0)
  row_i <- row_i + 1
  contrast_rows[[row_i]] <- tibble(
    ID = pid,
    effect = "catchup_late_minus_early",
    median = median(v),
    lower = as.numeric(quantile(v, 0.025)),
    upper = as.numeric(quantile(v, 0.975)),
    p_lt_0 = mean(v < 0),
    p_gt_0 = mean(v > 0),
    p_in_rope = mean(abs(v) < rope_abs)
  )

  # ---- 4) Practice: late - early change in 50→350 duration effect, by cue ----
  for (cue in c("NR", "R")) {
    wide_cue <- draws_ds %>%
      filter(CueType == cue, SessionSlice %in% c("early", "late")) %>%
      select(.draw, SessionSlice, .value) %>%
      pivot_wider(names_from = SessionSlice, values_from = .value)
    v <- wide_cue$late - wide_cue$early
    v <- v[is.finite(v)]
    stopifnot(length(v) > 0)
    row_i <- row_i + 1
    contrast_rows[[row_i]] <- tibble(
      ID = pid,
      effect = paste0(cue, ": duration_effect_late_minus_early"),
      median = median(v),
      lower = as.numeric(quantile(v, 0.025)),
      upper = as.numeric(quantile(v, 0.975)),
      p_lt_0 = mean(v < 0),
      p_gt_0 = mean(v > 0),
      p_in_rope = mean(abs(v) < rope_abs)
    )
  }
}

primary_summary <- bind_rows(contrast_rows) %>%
  arrange(effect, ID)

primary_summary

# Compact table of the main theory questions only.
theory_focus <- primary_summary %>%
  filter(
    effect %in% c(
      "gain_at_50ms",
      "NR: linear_trend", "R: linear_trend",
      "NR: 350ms - 50ms", "R: 350ms - 50ms",
      "catchup_early", "catchup_average", "catchup_late",
      "catchup_late_minus_early",
      "NR: duration_effect_late_minus_early",
      "R: duration_effect_late_minus_early"
    )
  ) %>%
  mutate(
    question = case_when(
      effect == "gain_at_50ms" ~ "1. redundancy advantage",
      grepl("linear_trend|350ms - 50ms", effect) ~ "2. duration change by cue",
      grepl("^catchup_", effect) ~ "3. NR catch-up vs R",
      grepl("duration_effect_late_minus_early", effect) ~ "4. practice change of duration slope",
      TRUE ~ "other"
    )
  ) %>%
  arrange(question, effect, ID)

theory_focus

# Persist AbsErr contrast tables (mirrors discrete script outputs).
dir.create("bayes_primary", showWarnings = FALSE, recursive = TRUE)
write_csv(theory_focus, "bayes_primary/theory_focus.csv")
write_csv(primary_summary, "bayes_primary/primary_summary.csv")

# Reading guide:
# - duration p_lt_0: P(AbsErr decreases with longer encoding)
# - catch-up p_gt_0: P(NR improves more than R)
# - practice p_in_rope: P(duration slope change within ±3 deg)
```

## 3.3 Posterior predictive checks

Student-t on AbsErr is for robust mean contrasts. Predictions can be <0 or >180,
so density overlays are often misleading. Use ECDF + clipped histograms.

```{r bayes-ppc, fig.width=10, fig.height=7}
# ECDF overlays (robust to extreme predictive draws).
p_ecdf_AQ <- pp_check(m_AQ, ndraws = 50, type = "ecdf_overlay") +
  coord_cartesian(xlim = c(0, 180)) +
  labs(title = "AQ: ECDF", x = "AbsErr (deg)")
p_ecdf_HC <- pp_check(m_HC, ndraws = 50, type = "ecdf_overlay") +
  coord_cartesian(xlim = c(0, 180)) +
  labs(title = "HC: ECDF", x = "AbsErr (deg)")
p_ecdf_YILIU <- pp_check(m_YILIU, ndraws = 50, type = "ecdf_overlay") +
  coord_cartesian(xlim = c(0, 180)) +
  labs(title = "YILIU: ECDF", x = "AbsErr (deg)")

# Histograms: clip yrep to [0, 180] for display only (bayesplot forbids NAs).
yrep_AQ <- pmin(pmax(posterior_predict(m_AQ, ndraws = 8), 0), 180)
yrep_HC <- pmin(pmax(posterior_predict(m_HC, ndraws = 8), 0), 180)
yrep_YILIU <- pmin(pmax(posterior_predict(m_YILIU, ndraws = 8), 0), 180)

p_hist_AQ <- bayesplot::ppc_hist(get_y(m_AQ), yrep_AQ, binwidth = 5) +
  coord_cartesian(xlim = c(0, 180)) +
  labs(title = "AQ: hist", x = "AbsErr (deg)")
p_hist_HC <- bayesplot::ppc_hist(get_y(m_HC), yrep_HC, binwidth = 5) +
  coord_cartesian(xlim = c(0, 180)) +
  labs(title = "HC: hist", x = "AbsErr (deg)")
p_hist_YILIU <- bayesplot::ppc_hist(get_y(m_YILIU), yrep_YILIU, binwidth = 5) +
  coord_cartesian(xlim = c(0, 180)) +
  labs(title = "YILIU: hist", x = "AbsErr (deg)")

(p_ecdf_AQ | p_ecdf_HC | p_ecdf_YILIU) /
  (p_hist_AQ | p_hist_HC | p_hist_YILIU)
```

## 3.4 Contrast plots

```{r bayes-contrast-plots, fig.width=10, fig.height=8}
# Pull the three plot-ready subsets from the summary table.
plot_duration <- primary_summary %>%
  filter(effect %in% c("NR: 350ms - 50ms", "R: 350ms - 50ms")) %>%
  mutate(CueType = if_else(grepl("^NR", effect), "NR", "R"))

plot_catch <- primary_summary %>%
  filter(effect %in% c("catchup_early", "catchup_average", "catchup_late")) %>%
  mutate(
    SessionSlice = factor(
      gsub("catchup_", "", effect),
      levels = c("early", "average", "late")
    )
  )

plot_practice <- primary_summary %>%
  filter(grepl("duration_effect_late_minus_early", effect)) %>%
  mutate(CueType = if_else(grepl("^NR", effect), "NR", "R"))

p1 <- ggplot(plot_duration, aes(CueType, median, ymin = lower, ymax = upper)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_hline(yintercept = c(-rope_abs, rope_abs), linetype = "dotted", colour = "grey40") +
  geom_pointrange() +
  facet_wrap(~ID) +
  labs(
    title = "50→350 ms AbsErr change by cue",
    subtitle = "Negative = lower error at 350 ms; dotted = ±3° ROPE",
    y = "Posterior median (deg)", x = "Cue"
  )

p2 <- ggplot(plot_catch, aes(SessionSlice, median, ymin = lower, ymax = upper, group = 1)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_hline(yintercept = c(-rope_abs, rope_abs), linetype = "dotted", colour = "grey40") +
  geom_line() +
  geom_pointrange() +
  facet_wrap(~ID) +
  labs(
    title = "NR catch-up across session slices",
    subtitle = "Positive = NR improved more than R (50→350)",
    y = "Posterior median (deg)", x = "Session slice"
  )

p3 <- ggplot(plot_practice, aes(CueType, median, ymin = lower, ymax = upper)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_hline(yintercept = c(-rope_abs, rope_abs), linetype = "dotted", colour = "grey40") +
  geom_pointrange() +
  facet_wrap(~ID) +
  labs(
    title = "Practice change in 50→350 duration effect",
    subtitle = "Late − early; near 0 / in ROPE = slope stable",
    y = "Posterior median (deg)", x = "Cue"
  )

p1 / p2 / p3
```

# 4. Bayesian primary analysis: cirSD (continuous session)

Parallel to AbsErr, but the outcome is **circular SD** of signed errors.
cirSD is a cell summary (one value per Cue × Duration × Session). To keep the
**same design formula** as AbsErr and identify residual variance, we use
**B bootstrap cirSD replicates per cell** (resample trials within the cell).

Model (same fixed + random structure as AbsErr continuous):

`cirSD_deg ~ CueType * DurationF * Session_z + (1 | SessionF)`

Mixture modelling is **not** repeated here: the mixture already targets the
signed-error distribution directly (finer than AbsErr or cirSD summaries).

## Build bootstrap cirSD dataset and fit / load models

```{r cirsd-bayes-data, message=FALSE, warning=FALSE}
rope_cir <- rope_abs

B_cir <- 40
set.seed(1234)

circ_sd_deg <- function(x) {
  x <- x[is.finite(x)]
  if (length(x) < 5) return(NA_real_)
  as.numeric(sd.circular(circular(x, units = "degrees"))) * (180 / pi)
}

build_cirsd_boot <- function(d_sub, B = B_cir) {
  rows <- list()
  i <- 0
  cues <- levels(d_sub$CueType)
  durs <- levels(d_sub$DurationF)
  sess <- levels(d_sub$SessionF)
  for (cue in cues) {
    for (dur in durs) {
      for (sess_i in sess) {
        idx <- which(
          d_sub$CueType == cue &
            d_sub$DurationF == dur &
            d_sub$SessionF == sess_i
        )
        x <- d_sub$SignedErr[idx]
        x <- x[is.finite(x)]
        stopifnot(length(x) >= 10)
        sz <- unique(d_sub$Session_z[idx])
        stopifnot(length(sz) == 1)
        for (b in seq_len(B)) {
          xb <- sample(x, replace = TRUE)
          i <- i + 1
          rows[[i]] <- tibble(
            CueType = cue,
            DurationF = dur,
            SessionF = sess_i,
            Session_z = sz,
            boot_id = b,
            cirSD_deg = circ_sd_deg(xb)
          )
        }
      }
    }
  }
  bind_rows(rows) %>%
    mutate(
      CueType = factor(CueType, levels = levels(d_sub$CueType)),
      DurationF = factor(DurationF, levels = levels(d_sub$DurationF)),
      SessionF = factor(SessionF, levels = levels(d_sub$SessionF))
    )
}

dir.create("bayes_primary", showWarnings = FALSE, recursive = TRUE)

cir_boot_files <- c(
  AQ = "bayes_primary/cirsd_boot_AQ.rds",
  HC = "bayes_primary/cirsd_boot_HC.rds",
  YILIU = "bayes_primary/cirsd_boot_YILIU.rds"
)
d_cir_AQ <- if (file.exists(cir_boot_files[["AQ"]])) {
  readRDS(cir_boot_files[["AQ"]])
} else {
  tmp <- build_cirsd_boot(d_AQ)
  saveRDS(tmp, cir_boot_files[["AQ"]])
  tmp
}
d_cir_HC <- if (file.exists(cir_boot_files[["HC"]])) {
  readRDS(cir_boot_files[["HC"]])
} else {
  tmp <- build_cirsd_boot(d_HC)
  saveRDS(tmp, cir_boot_files[["HC"]])
  tmp
}
d_cir_YILIU <- if (file.exists(cir_boot_files[["YILIU"]])) {
  readRDS(cir_boot_files[["YILIU"]])
} else {
  tmp <- build_cirsd_boot(d_YILIU)
  saveRDS(tmp, cir_boot_files[["YILIU"]])
  tmp
}

stopifnot(nrow(d_cir_AQ) == 2 * 7 * 10 * B_cir)

bayes_priors_cir <- c(
  prior(normal(40, 20), class = "Intercept"),
  prior(normal(0, 15), class = "b"),
  prior(exponential(1), class = "sigma"),
  prior(exponential(1), class = "sd"),
  prior(gamma(2, 0.1), class = "nu")
)

m_cir_AQ <- brm(
  cirSD_deg ~ CueType * DurationF * Session_z + (1 | SessionF),
  data = d_cir_AQ,
  family = student(),
  prior = bayes_priors_cir,
  chains = 4, cores = 4, iter = 4000, warmup = 2000,
  control = list(adapt_delta = 0.95, max_treedepth = 12),
  seed = 1234,
  file = "bayes_primary/m_AQ_cirsd",
  file_refit = "on_change"
)

m_cir_HC <- brm(
  cirSD_deg ~ CueType * DurationF * Session_z + (1 | SessionF),
  data = d_cir_HC,
  family = student(),
  prior = bayes_priors_cir,
  chains = 4, cores = 4, iter = 4000, warmup = 2000,
  control = list(adapt_delta = 0.95, max_treedepth = 12),
  seed = 1234,
  file = "bayes_primary/m_HC_cirsd",
  file_refit = "on_change"
)

m_cir_YILIU <- brm(
  cirSD_deg ~ CueType * DurationF * Session_z + (1 | SessionF),
  data = d_cir_YILIU,
  family = student(),
  prior = bayes_priors_cir,
  chains = 4, cores = 4, iter = 4000, warmup = 2000,
  control = list(adapt_delta = 0.95, max_treedepth = 12),
  seed = 1234,
  file = "bayes_primary/m_YILIU_cirsd",
  file_refit = "on_change"
)

summary(m_cir_AQ)
summary(m_cir_HC)
summary(m_cir_YILIU)
```

## 4.2 Theory contrasts (cirSD; same questions as AbsErr)

```{r cirsd-bayes-contrasts, message=FALSE, warning=FALSE}
contrast_rows_cir <- list()
row_i <- 0
models_cir <- list(AQ = m_cir_AQ, HC = m_cir_HC, YILIU = m_cir_YILIU)

for (pid in names(models_cir)) {
  model <- models_cir[[pid]]

  emm_gain <- emmeans(model, ~ CueType | DurationF, at = list(Session_z = 0))
  ct_gain <- contrast(emm_gain, method = list("NR - R" = c(1, -1)), by = "DurationF")
  draws_gain <- gather_emmeans_draws(ct_gain)

  for (dur in levels(d$DurationF)) {
    v <- draws_gain$.value[draws_gain$DurationF == dur]
    v <- v[is.finite(v)]
    stopifnot(length(v) > 0)
    row_i <- row_i + 1
    contrast_rows_cir[[row_i]] <- tibble(
      ID = pid, effect = paste0("gain_", dur),
      median = median(v),
      lower = as.numeric(quantile(v, 0.025)),
      upper = as.numeric(quantile(v, 0.975)),
      p_lt_0 = mean(v < 0), p_gt_0 = mean(v > 0),
      p_in_rope = mean(abs(v) < rope_cir)
    )
  }
  v <- draws_gain$.value[draws_gain$DurationF == "50ms"]
  v <- v[is.finite(v)]
  row_i <- row_i + 1
  contrast_rows_cir[[row_i]] <- tibble(
    ID = pid, effect = "gain_at_50ms",
    median = median(v),
    lower = as.numeric(quantile(v, 0.025)),
    upper = as.numeric(quantile(v, 0.975)),
    p_lt_0 = mean(v < 0), p_gt_0 = mean(v > 0),
    p_in_rope = mean(abs(v) < rope_cir)
  )

  emm_dur <- emmeans(model, ~ DurationF | CueType, at = list(Session_z = 0))
  ct_dur <- contrast(
    emm_dur,
    method = list("linear_trend" = w_linear, "350ms - 50ms" = w_350_minus_50),
    by = "CueType"
  )
  draws_dur <- gather_emmeans_draws(ct_dur)
  for (cue in c("NR", "R")) {
    for (cname in c("linear_trend", "350ms - 50ms")) {
      v <- draws_dur$.value[draws_dur$CueType == cue & draws_dur$contrast == cname]
      v <- v[is.finite(v)]
      stopifnot(length(v) > 0)
      row_i <- row_i + 1
      contrast_rows_cir[[row_i]] <- tibble(
        ID = pid, effect = paste0(cue, ": ", cname),
        median = median(v),
        lower = as.numeric(quantile(v, 0.025)),
        upper = as.numeric(quantile(v, 0.975)),
        p_lt_0 = mean(v < 0), p_gt_0 = mean(v > 0),
        p_in_rope = mean(abs(v) < rope_cir)
      )
    }
  }

  emm_dur_sess <- emmeans(
    model,
    ~ DurationF | CueType * Session_z,
    at = list(Session_z = c(early_z, 0, late_z))
  )
  ct_dur_sess <- contrast(
    emm_dur_sess,
    method = list("350ms - 50ms" = w_350_minus_50),
    by = c("CueType", "Session_z")
  )
  draws_ds <- gather_emmeans_draws(ct_dur_sess) %>% ungroup()
  sz_levels <- sort(unique(draws_ds$Session_z))
  stopifnot(length(sz_levels) == 3)
  draws_ds <- draws_ds %>%
    mutate(SessionSlice = c("early", "average", "late")[match(Session_z, sz_levels)]) %>%
    select(.draw, CueType, SessionSlice, .value)

  for (slice in c("early", "average", "late")) {
    wide <- draws_ds %>%
      filter(SessionSlice == slice) %>%
      pivot_wider(names_from = CueType, values_from = .value)
    v <- wide$R - wide$NR
    v <- v[is.finite(v)]
    stopifnot(length(v) > 0)
    row_i <- row_i + 1
    contrast_rows_cir[[row_i]] <- tibble(
      ID = pid, effect = paste0("catchup_", slice),
      median = median(v),
      lower = as.numeric(quantile(v, 0.025)),
      upper = as.numeric(quantile(v, 0.975)),
      p_lt_0 = mean(v < 0), p_gt_0 = mean(v > 0),
      p_in_rope = mean(abs(v) < rope_cir)
    )
  }

  wide_e <- draws_ds %>%
    filter(SessionSlice == "early") %>%
    pivot_wider(names_from = CueType, values_from = .value) %>%
    mutate(catchup = R - NR)
  wide_l <- draws_ds %>%
    filter(SessionSlice == "late") %>%
    pivot_wider(names_from = CueType, values_from = .value) %>%
    mutate(catchup = R - NR)
  v <- wide_l$catchup[match(wide_e$.draw, wide_l$.draw)] - wide_e$catchup
  v <- v[is.finite(v)]
  row_i <- row_i + 1
  contrast_rows_cir[[row_i]] <- tibble(
    ID = pid, effect = "catchup_late_minus_early",
    median = median(v),
    lower = as.numeric(quantile(v, 0.025)),
    upper = as.numeric(quantile(v, 0.975)),
    p_lt_0 = mean(v < 0), p_gt_0 = mean(v > 0),
    p_in_rope = mean(abs(v) < rope_cir)
  )

  for (cue in c("NR", "R")) {
    wc <- draws_ds %>%
      filter(CueType == cue, SessionSlice %in% c("early", "late")) %>%
      select(.draw, SessionSlice, .value) %>%
      pivot_wider(names_from = SessionSlice, values_from = .value)
    v <- wc$late - wc$early
    v <- v[is.finite(v)]
    stopifnot(length(v) > 0)
    row_i <- row_i + 1
    contrast_rows_cir[[row_i]] <- tibble(
      ID = pid,
      effect = paste0(cue, ": duration_effect_late_minus_early"),
      median = median(v),
      lower = as.numeric(quantile(v, 0.025)),
      upper = as.numeric(quantile(v, 0.975)),
      p_lt_0 = mean(v < 0), p_gt_0 = mean(v > 0),
      p_in_rope = mean(abs(v) < rope_cir)
    )
  }
}

primary_summary_cir <- bind_rows(contrast_rows_cir) %>% arrange(effect, ID)
primary_summary_cir

theory_focus_cir <- primary_summary_cir %>%
  filter(
    effect %in% c(
      "gain_at_50ms",
      "NR: linear_trend", "R: linear_trend",
      "NR: 350ms - 50ms", "R: 350ms - 50ms",
      "catchup_early", "catchup_average", "catchup_late",
      "catchup_late_minus_early",
      "NR: duration_effect_late_minus_early",
      "R: duration_effect_late_minus_early"
    )
  ) %>%
  mutate(
    question = case_when(
      effect == "gain_at_50ms" ~ "1. redundancy advantage",
      grepl("linear_trend|350ms - 50ms", effect) ~ "2. duration change by cue",
      grepl("^catchup_", effect) ~ "3. NR catch-up vs R",
      grepl("duration_effect_late_minus_early", effect) ~ "4. practice change of duration slope",
      TRUE ~ "other"
    )
  ) %>%
  arrange(question, effect, ID)

theory_focus_cir

write_csv(theory_focus_cir, "bayes_primary/theory_focus_cirsd.csv")
write_csv(primary_summary_cir, "bayes_primary/primary_summary_cirsd.csv")
```

## 4.3 Posterior predictive checks (cirSD)

```{r cirsd-bayes-ppc, fig.width=10, fig.height=4}
p_ecdf_cir_AQ <- pp_check(m_cir_AQ, ndraws = 50, type = "ecdf_overlay") +
  labs(title = "AQ: cirSD ECDF", x = "cirSD (deg)")
p_ecdf_cir_HC <- pp_check(m_cir_HC, ndraws = 50, type = "ecdf_overlay") +
  labs(title = "HC: cirSD ECDF", x = "cirSD (deg)")
p_ecdf_cir_YILIU <- pp_check(m_cir_YILIU, ndraws = 50, type = "ecdf_overlay") +
  labs(title = "YILIU: cirSD ECDF", x = "cirSD (deg)")

p_ecdf_cir_AQ | p_ecdf_cir_HC | p_ecdf_cir_YILIU
```

## 4.4 Contrast plots (cirSD)

```{r cirsd-bayes-plots, fig.width=10, fig.height=8}
plot_duration_cir <- primary_summary_cir %>%
  filter(effect %in% c("NR: 350ms - 50ms", "R: 350ms - 50ms")) %>%
  mutate(CueType = if_else(grepl("^NR", effect), "NR", "R"))

plot_catch_cir <- primary_summary_cir %>%
  filter(effect %in% c("catchup_early", "catchup_average", "catchup_late")) %>%
  mutate(
    SessionSlice = factor(
      gsub("catchup_", "", effect),
      levels = c("early", "average", "late")
    )
  )

plot_practice_cir <- primary_summary_cir %>%
  filter(grepl("duration_effect_late_minus_early", effect)) %>%
  mutate(CueType = if_else(grepl("^NR", effect), "NR", "R"))

p1c <- ggplot(plot_duration_cir, aes(CueType, median, ymin = lower, ymax = upper)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_hline(yintercept = c(-rope_cir, rope_cir), linetype = "dotted", colour = "grey40") +
  geom_pointrange() +
  facet_wrap(~ID) +
  labs(
    title = "50→350 ms cirSD change by cue",
    subtitle = "Negative = lower cirSD at 350 ms; dotted = ±ROPE",
    y = "Posterior median (deg)", x = "Cue"
  )

p2c <- ggplot(plot_catch_cir, aes(SessionSlice, median, ymin = lower, ymax = upper, group = 1)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_hline(yintercept = c(-rope_cir, rope_cir), linetype = "dotted", colour = "grey40") +
  geom_line() +
  geom_pointrange() +
  facet_wrap(~ID) +
  labs(
    title = "NR catch-up in cirSD across session slices",
    subtitle = "Positive = NR cirSD fell more than R (50→350)",
    y = "Posterior median (deg)", x = "Session slice"
  )

p3c <- ggplot(plot_practice_cir, aes(CueType, median, ymin = lower, ymax = upper)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_hline(yintercept = c(-rope_cir, rope_cir), linetype = "dotted", colour = "grey40") +
  geom_pointrange() +
  facet_wrap(~ID) +
  labs(
    title = "Practice change in 50→350 cirSD duration effect",
    subtitle = "Late − early",
    y = "Posterior median (deg)", x = "Cue"
  )

p1c / p2c / p3c
```

# 5. Pre-mixture 1: AbsErr Session 1 vs Session 10 by duration

Question: at fixed cue and duration, is AbsErr higher in Session 1 than Session 10?
Positive = worse (larger AbsErr) in Session 1. Focus on short durations (≤150 ms).

```{r abserr-s1-s10, message=FALSE, warning=FALSE}
s1s10_rows <- list()
row_i <- 0

for (pid in names(models)) {
  model <- models[[pid]]

  # Model-estimated AbsErr at S1 and S10 for every cue × duration cell.
  # Session_z order in `at` is s1 then s10; contrast c(1, -1) = S1 − S10.
  emm <- emmeans(
    model,
    ~ Session_z | CueType * DurationF,
    at = list(Session_z = c(s1_z, s10_z))
  )
  ct <- contrast(emm, method = list("S1 - S10" = c(1, -1)), by = c("CueType", "DurationF"))
  draws <- gather_emmeans_draws(ct)

  for (cue in c("NR", "R")) {
    for (dur in levels(d$DurationF)) {
      v <- draws$.value[draws$CueType == cue & draws$DurationF == dur]
      v <- v[is.finite(v)]
      stopifnot(length(v) > 0)
      dur_ms <- as.numeric(gsub("ms", "", dur))
      row_i <- row_i + 1
      s1s10_rows[[row_i]] <- tibble(
        ID = pid,
        CueType = cue,
        DurationF = dur,
        Duration_ms = dur_ms,
        short_duration = dur_ms <= 150,
        median = median(v),
        lower = as.numeric(quantile(v, 0.025)),
        upper = as.numeric(quantile(v, 0.975)),
        p_gt_0 = mean(v > 0),
        p_lt_0 = mean(v < 0),
        p_in_rope = mean(abs(v) < rope_abs)
      )
    }

    # Short-duration average (50/100/150): mean of the three cell contrasts per draw.
    short <- draws %>%
      filter(CueType == cue, DurationF %in% c("50ms", "100ms", "150ms")) %>%
      group_by(.draw) %>%
      summarise(.value = mean(.value), .groups = "drop")
    v <- short$.value[is.finite(short$.value)]
    row_i <- row_i + 1
    s1s10_rows[[row_i]] <- tibble(
      ID = pid,
      CueType = cue,
      DurationF = "short_mean",
      Duration_ms = NA_real_,
      short_duration = TRUE,
      median = median(v),
      lower = as.numeric(quantile(v, 0.025)),
      upper = as.numeric(quantile(v, 0.975)),
      p_gt_0 = mean(v > 0),
      p_lt_0 = mean(v < 0),
      p_in_rope = mean(abs(v) < rope_abs)
    )
  }
}

abserr_s1_s10 <- bind_rows(s1s10_rows)
abserr_s1_s10_cell <- abserr_s1_s10 %>%
  filter(DurationF != "short_mean") %>%
  arrange(ID, CueType, Duration_ms)
abserr_s1_s10_short <- abserr_s1_s10 %>%
  filter(DurationF == "short_mean") %>%
  arrange(ID, CueType)

abserr_s1_s10_cell
abserr_s1_s10_short

write_csv(abserr_s1_s10_cell, "bayes_primary/abserr_s1_vs_s10_by_duration.csv")
write_csv(abserr_s1_s10_short, "bayes_primary/abserr_s1_vs_s10_short.csv")
```

```{r abserr-s1-s10-plot, fig.width=10, fig.height=5}
ggplot(
  abserr_s1_s10_cell,
  aes(Duration_ms, median, ymin = lower, ymax = upper, colour = CueType, group = CueType)
) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_hline(yintercept = c(-rope_abs, rope_abs), linetype = "dotted", colour = "grey40") +
  geom_line() +
  geom_pointrange() +
  facet_wrap(~ID) +
  labs(
    title = "AbsErr Session 1 − Session 10 by duration",
    subtitle = "Positive = higher AbsErr in Session 1; dotted = ±3° ROPE",
    x = "Encoding duration (ms)",
    y = "Posterior median difference (deg)"
  )
```

# 6. Pre-mixture 2: cirSD Session 1 vs Session 10 (bootstrap)

Companion in circular-SD units. For each participant × cue × duration:
compute cirSD in Session 1 and Session 10, then bootstrap S1 − S10.

```{r cirsd-s1-s10, message=FALSE, warning=FALSE}
B <- 1000  # bootstrap replicates
set.seed(1234)

ids <- c("AQ", "HC", "YILIU")
cues <- c("NR", "R")
durs <- levels(d$DurationF)

cirsd_rows <- list()
row_i <- 0

for (pid in ids) {
  for (cue in cues) {
    for (dur in durs) {
      # Trials in this cell for Session 1 and Session 10.
      x1 <- d$SignedErr[d$ID == pid & d$CueType == cue & d$DurationF == dur & d$Session == 1]
      x10 <- d$SignedErr[d$ID == pid & d$CueType == cue & d$DurationF == dur & d$Session == 10]
      x1 <- x1[is.finite(x1)]
      x10 <- x10[is.finite(x10)]

      # Need enough trials for a stable circular SD.
      if (length(x1) < 5 || length(x10) < 5) {
        row_i <- row_i + 1
        cirsd_rows[[row_i]] <- tibble(
          ID = pid, CueType = cue, DurationF = dur,
          Duration_ms = as.numeric(gsub("ms", "", dur)),
          estimate = NA_real_, lower = NA_real_, upper = NA_real_,
          boot_se = NA_real_, n1 = length(x1), n10 = length(x10),
          excludes_0 = NA
        )
        next
      }

      # Observed cirSD difference (degrees).
      sd1 <- as.numeric(sd.circular(circular(x1, units = "degrees"))) * (180 / pi)
      sd10 <- as.numeric(sd.circular(circular(x10, units = "degrees"))) * (180 / pi)
      obs <- sd1 - sd10

      # Bootstrap: resample each session's trials with replacement.
      boot_vals <- numeric(B)
      for (b in seq_len(B)) {
        s1_b <- sample(x1, replace = TRUE)
        s10_b <- sample(x10, replace = TRUE)
        sd1_b <- as.numeric(sd.circular(circular(s1_b, units = "degrees"))) * (180 / pi)
        sd10_b <- as.numeric(sd.circular(circular(s10_b, units = "degrees"))) * (180 / pi)
        boot_vals[b] <- sd1_b - sd10_b
      }
      boot_vals <- boot_vals[is.finite(boot_vals)]

      lo <- as.numeric(quantile(boot_vals, 0.025))
      hi <- as.numeric(quantile(boot_vals, 0.975))

      row_i <- row_i + 1
      cirsd_rows[[row_i]] <- tibble(
        ID = pid, CueType = cue, DurationF = dur,
        Duration_ms = as.numeric(gsub("ms", "", dur)),
        estimate = obs,
        lower = lo,
        upper = hi,
        boot_se = sd(boot_vals),
        n1 = length(x1),
        n10 = length(x10),
        excludes_0 = (lo > 0 & hi > 0) | (lo < 0 & hi < 0)
      )
    }
  }
}

cirsd_s1_s10 <- bind_rows(cirsd_rows) %>%
  arrange(ID, CueType, Duration_ms)

cirsd_s1_s10
write_csv(cirsd_s1_s10, "bayes_primary/cirsd_s1_vs_s10_bootstrap.csv")
```

```{r cirsd-s1-s10-short, message=FALSE, warning=FALSE}
# Short-duration companion: mean of cirSD(50/100/150) in S1 minus same mean in S10.
# Bootstrap resamples trials within each duration × session cell, then re-averages.
cirsd_short_rows <- list()
row_i <- 0
short_durs <- c("50ms", "100ms", "150ms")

for (pid in ids) {
  for (cue in cues) {
    # Observed mean cirSD across the three short durations.
    sd1_vec <- numeric(3)
    sd10_vec <- numeric(3)
    ok <- TRUE
    for (j in seq_along(short_durs)) {
      dur <- short_durs[j]
      x1 <- d$SignedErr[d$ID == pid & d$CueType == cue & d$DurationF == dur & d$Session == 1]
      x10 <- d$SignedErr[d$ID == pid & d$CueType == cue & d$DurationF == dur & d$Session == 10]
      x1 <- x1[is.finite(x1)]
      x10 <- x10[is.finite(x10)]
      if (length(x1) < 5 || length(x10) < 5) {
        ok <- FALSE
        break
      }
      sd1_vec[j] <- as.numeric(sd.circular(circular(x1, units = "degrees"))) * (180 / pi)
      sd10_vec[j] <- as.numeric(sd.circular(circular(x10, units = "degrees"))) * (180 / pi)
    }

    if (!ok) {
      row_i <- row_i + 1
      cirsd_short_rows[[row_i]] <- tibble(
        ID = pid, CueType = cue,
        estimate = NA_real_, lower = NA_real_, upper = NA_real_,
        boot_se = NA_real_, excludes_0 = NA
      )
      next
    }

    obs <- mean(sd1_vec) - mean(sd10_vec)

    boot_vals <- numeric(B)
    for (b in seq_len(B)) {
      sd1_b <- numeric(3)
      sd10_b <- numeric(3)
      for (j in seq_along(short_durs)) {
        dur <- short_durs[j]
        x1 <- d$SignedErr[d$ID == pid & d$CueType == cue & d$DurationF == dur & d$Session == 1]
        x10 <- d$SignedErr[d$ID == pid & d$CueType == cue & d$DurationF == dur & d$Session == 10]
        x1 <- x1[is.finite(x1)]
        x10 <- x10[is.finite(x10)]
        s1_b <- sample(x1, replace = TRUE)
        s10_b <- sample(x10, replace = TRUE)
        sd1_b[j] <- as.numeric(sd.circular(circular(s1_b, units = "degrees"))) * (180 / pi)
        sd10_b[j] <- as.numeric(sd.circular(circular(s10_b, units = "degrees"))) * (180 / pi)
      }
      boot_vals[b] <- mean(sd1_b) - mean(sd10_b)
    }
    boot_vals <- boot_vals[is.finite(boot_vals)]
    lo <- as.numeric(quantile(boot_vals, 0.025))
    hi <- as.numeric(quantile(boot_vals, 0.975))

    row_i <- row_i + 1
    cirsd_short_rows[[row_i]] <- tibble(
      ID = pid, CueType = cue,
      estimate = obs, lower = lo, upper = hi,
      boot_se = sd(boot_vals),
      excludes_0 = (lo > 0 & hi > 0) | (lo < 0 & hi < 0)
    )
  }
}

cirsd_short <- bind_rows(cirsd_short_rows) %>% arrange(ID, CueType)
cirsd_short
write_csv(cirsd_short, "bayes_primary/cirsd_s1_vs_s10_short_bootstrap.csv")
```

```{r cirsd-s1-s10-plot, fig.width=10, fig.height=5}
ggplot(
  cirsd_s1_s10,
  aes(Duration_ms, estimate, ymin = lower, ymax = upper, colour = CueType, group = CueType)
) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_line() +
  geom_pointrange() +
  facet_wrap(~ID) +
  labs(
    title = "cirSD Session 1 − Session 10 by duration",
    subtitle = "Bootstrap 95% CI; positive = higher circular variability in Session 1",
    x = "Encoding duration (ms)",
    y = "cirSD difference (deg)"
  )
```

# 7. Circular mixture: per-cell ML (target + guess + swap)

Bays-style mixture on signed errors, one independent ML fit per cue × duration
cell (14 cells × participant; sessions pooled). Always 3 unique non-target
colours, so the swap term is three equal von Mises bumps:

`p(θ) = p_target · VM(0, κ) + (β/3) Σ_{j=1..3} VM(μ_j, κ) + γ · Uniform`

- Shared κ for target and colour-swap
- Stick-break order: estimate β (swap) first, then γ from the leftover
  (pure guesses expected rare); p_target = 1 − β − γ
- Fits from `Rscript s1_mixture_cells.R` → `mixture_cells/all_cell_params.csv`
  (Part B of that script is a single worked example; Part C loops all cells)

```{r mixture-load}
mix_file <- "mixture_cells/all_cell_params.csv"
stopifnot(file.exists(mix_file))

mix_cells <- read_csv(mix_file, show_col_types = FALSE) %>%
  mutate(
    CueType = factor(CueType, levels = c("NR", "R")),
    DurationF = factor(
      DurationF,
      levels = c("50ms", "100ms", "150ms", "200ms", "250ms", "300ms", "350ms")
    ),
    Duration_ms = as.numeric(gsub("ms", "", as.character(DurationF)))
  )

# Basic safety checks.
stopifnot(all(c("AQ", "HC", "YILIU") %in% mix_cells$ID))
stopifnot(all(mix_cells$convergence == 0 | is.na(mix_cells$convergence)))
stopifnot(all(mix_cells$p_target + mix_cells$gamma + mix_cells$beta > 0.99))
stopifnot(all(mix_cells$p_target + mix_cells$gamma + mix_cells$beta < 1.01))

mix_cells %>%
  select(ID, CueType, DurationF, n_trials, kappa, circSD_deg,
         p_target, gamma, beta, nll, convergence) %>%
  mutate(across(c(kappa, circSD_deg, p_target, gamma, beta, nll), ~ round(.x, 3))) %>%
  arrange(ID, CueType, DurationF)
```

```{r mixture-proportion-plot, fig.width=10, fig.height=6}
# Stacked response proportions (target / swap / guess) — must sum to 1.
prop_long <- mix_cells %>%
  select(ID, CueType, Duration_ms, p_target, gamma, beta) %>%
  pivot_longer(
    c(p_target, gamma, beta),
    names_to = "component",
    values_to = "proportion"
  ) %>%
  mutate(
    component = factor(
      component,
      levels = c("p_target", "beta", "gamma"),
      labels = c("Target", "Colour swap", "Guess")
    )
  )

ggplot(prop_long, aes(Duration_ms, proportion, fill = component)) +
  geom_col(width = 35, colour = "grey20", linewidth = 0.2) +
  facet_grid(CueType ~ ID) +
  scale_x_continuous(breaks = seq(50, 350, by = 50)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 1)) +
  labs(
    title = "Mixture response proportions by cue and duration",
    subtitle = "Independent ML fits per cell; sessions pooled",
    x = "Encoding duration (ms)",
    y = "Response proportion",
    fill = "Component"
  ) +
  theme(legend.position = "bottom")
```

```{r mixture-kappa-plot, fig.width=10, fig.height=4}
# κ is precision, not a proportion — plot separately.
ggplot(mix_cells, aes(Duration_ms, kappa, colour = CueType, group = CueType)) +
  geom_line() +
  geom_point(size = 2) +
  facet_wrap(~ID, scales = "free_y") +
  scale_x_continuous(breaks = seq(50, 350, by = 50)) +
  labs(
    title = "Shared von Mises precision (κ) by cue and duration",
    x = "Encoding duration (ms)",
    y = "κ"
  )
```

```{r mixture-weight-lines, fig.width=10, fig.height=6}
ggplot(prop_long, aes(Duration_ms, proportion, colour = CueType, group = CueType)) +
  geom_line() +
  geom_point() +
  facet_grid(component ~ ID, scales = "free_y") +
  scale_x_continuous(breaks = seq(50, 350, by = 100)) +
  labs(
    title = "Mixture weights by cue and duration",
    x = "Encoding duration (ms)",
    y = "Proportion"
  )
```


## Appendix

Legacy / sensitivity material. Not rewritten in the linear style of Sections 1–7;
kept for reference and frequentist checks. Primary claims come from Sections 3–7.

# Sensitivity Check: Participant-Level Mixed Models

Each currently complete participant (AQ, HC, YILIU) is analysed independently.
Cue type and encoding duration are categorical predictors, session progression
is continuous, and the random session intercept accounts for clustering of
trials within sessions.

```{r participant-level-models}
d_AQ <- droplevels(filter(d, ID == "AQ"))
d_HC <- droplevels(filter(d, ID == "HC"))
d_YILIU <- droplevels(filter(d, ID == "YILIU"))

m_AQ <- lme(
  fixed = AbsErr ~ CueType * DurationF * Session_z,
  random = ~ 1 | SessionF,
  data = d_AQ,
  method = "REML"
)

m_HC <- lme(
  fixed = AbsErr ~ CueType * DurationF * Session_z,
  random = ~ 1 | SessionF,
  data = d_HC,
  method = "REML"
)

m_YILIU <- lme(
  fixed = AbsErr ~ CueType * DurationF * Session_z,
  random = ~ 1 | SessionF,
  data = d_YILIU,
  method = "REML"
)

summary(m_AQ)
anova(m_AQ, type = "marginal")

summary(m_HC)
anova(m_HC, type = "marginal")

summary(m_YILIU)
anova(m_YILIU, type = "marginal")
```

## Descriptive Means and Variability

Observed absolute-error means and SDs, plus circular SD of signed errors, are
summarised by cue type, encoding duration, and session phase. Means quantify
average error magnitude. Absolute-error SD and circular SD quantify response
variability / precision.

```{r descriptive-means-and-sd}
library(circular)

circSD_deg <- function(x) {
  as.numeric(
    circular::sd.circular(circular(x, units = "degrees"))
  ) * (180 / pi)
}

descriptive_summary <- d %>%
  filter(ID %in% c("AQ", "HC", "YILIU")) %>%
  group_by(ID, CueType, DurationF, SessionPhase) %>%
  summarise(
    n = n(),
    mean_AbsErr = mean(AbsErr),
    sd_AbsErr = sd(AbsErr),
    circSD = circSD_deg(SignedErr),
    .groups = "drop"
  ) %>%
  arrange(ID, SessionPhase, CueType, DurationF)

descriptive_summary

descriptive_gain <- descriptive_summary %>%
  select(ID, CueType, DurationF, SessionPhase, mean_AbsErr, circSD) %>%
  pivot_wider(
    names_from = CueType,
    values_from = c(mean_AbsErr, circSD)
  ) %>%
  mutate(
    mean_gain = mean_AbsErr_NR - mean_AbsErr_R,
    circSD_gain = circSD_NR - circSD_R
  )

descriptive_gain
```

## Catch-Up Across Session Progression

Early sessions are Sessions 1–2; late sessions are Sessions 9–10; average
session uses `Session_z = 0`.

For each session slice:

- `350ms - 50ms` is estimated separately for NR and R. Negative values mean
  lower absolute error at 350 ms.
- Catch-up is `(R change) - (NR change)`. Positive values mean NR improved more
  than R, as predicted by a perceptual encoding account.
- Catch-up progression is `late catch-up - early catch-up`.

```{r participant-catch-up-contrasts}
early_z <- mean(unique(d$Session_z[d$Session %in% c(1, 2)]))
late_z <- mean(unique(d$Session_z[d$Session %in% c(9, 10)]))
session_slices <- c(early = early_z, average = 0, late = late_z)

catch_up_by_session <- function(model) {
  duration_by_cue <- emmeans(
    model,
    ~ DurationF | CueType * Session_z,
    at = list(Session_z = unname(session_slices))
  )

  cue_changes <- contrast(
    duration_by_cue,
    method = list(
      "350ms - 50ms" = c(-1, 0, 0, 0, 0, 0, 1)
    ),
    by = c("CueType", "Session_z")
  )

  catch_up <- contrast(
    cue_changes,
    method = list(
      "R change - NR change" = c(-1, 1)
    ),
    by = "Session_z"
  )

  # Session_z order from `at` is early, average, late after numerical sorting.
  catch_up_progression <- contrast(
    catch_up,
    method = list(
      "Late catch-up - early catch-up" = c(-1, 0, 1)
    ),
    by = NULL
  )

  list(
    cue_changes = summary(cue_changes, infer = c(TRUE, TRUE)),
    catch_up = summary(catch_up, infer = c(TRUE, TRUE)),
    catch_up_progression = summary(
      catch_up_progression,
      infer = c(TRUE, TRUE)
    )
  )
}

catch_AQ <- catch_up_by_session(m_AQ)
catch_HC <- catch_up_by_session(m_HC)
catch_YILIU <- catch_up_by_session(m_YILIU)

catch_AQ$cue_changes
catch_AQ$catch_up
catch_AQ$catch_up_progression

catch_HC$cue_changes
catch_HC$catch_up
catch_HC$catch_up_progression

catch_YILIU$cue_changes
catch_YILIU$catch_up
catch_YILIU$catch_up_progression
```

## Model-Estimated Means With Uncertainty

Fitted absolute-error means are reported with standard errors and 95% CIs for
each cue × duration cell at early, average, and late sessions. These are
uncertainty estimates for the means, not response-distribution SDs.

```{r model-estimated-means}
model_means_by_session <- function(model, participant) {
  emmeans(
    model,
    ~ CueType * DurationF | Session_z,
    at = list(Session_z = unname(session_slices))
  ) %>%
    summary(infer = c(TRUE, FALSE)) %>%
    as_tibble() %>%
    mutate(
      ID = participant,
      SessionSlice = case_when(
        abs(Session_z - early_z) < 1e-8 ~ "Early",
        abs(Session_z - 0) < 1e-8 ~ "Average",
        abs(Session_z - late_z) < 1e-8 ~ "Late"
      )
    ) %>%
    select(
      ID, SessionSlice, CueType, DurationF,
      emmean, SE, lower.CL, upper.CL
    )
}

model_means <- bind_rows(
  model_means_by_session(m_AQ, "AQ"),
  model_means_by_session(m_HC, "HC"),
  model_means_by_session(m_YILIU, "YILIU")
)

model_means

model_gain <- model_means %>%
  select(ID, SessionSlice, CueType, DurationF, emmean) %>%
  pivot_wider(names_from = CueType, values_from = emmean) %>%
  mutate(gain = NR - R)

model_gain
```

## Cue-Specific Duration and Session Effects

The joint models above directly test redundancy gain. The following models
analyse NR-cue and R-cue trials separately so that duration-related improvement
can be assessed within each cue condition.

Each model tests:

- whether absolute error differs across encoding durations;
- whether absolute error changes across sessions;
- whether the duration function changes across sessions.

```{r cue-specific-models}
fit_cue_model <- function(data, cue) {
  cue_data <- droplevels(filter(data, CueType == cue))

  lme(
    fixed = AbsErr ~ DurationF * Session_z,
    random = ~ 1 | SessionF,
    data = cue_data,
    method = "REML"
  )
}

m_AQ_NR <- fit_cue_model(d_AQ, "NR")
m_AQ_R <- fit_cue_model(d_AQ, "R")

m_HC_NR <- fit_cue_model(d_HC, "NR")
m_HC_R <- fit_cue_model(d_HC, "R")

m_YILIU_NR <- fit_cue_model(d_YILIU, "NR")
m_YILIU_R <- fit_cue_model(d_YILIU, "R")

cue_models <- list(
  AQ_NR = m_AQ_NR,
  AQ_R = m_AQ_R,
  HC_NR = m_HC_NR,
  HC_R = m_HC_R,
  YILIU_NR = m_YILIU_NR,
  YILIU_R = m_YILIU_R
)

for (model_name in names(cue_models)) {
  cat("\n\n###", model_name, "\n\n")
  print(anova(cue_models[[model_name]], type = "marginal"))
}
```

## Does the Encoding-Duration Effect Reduce with Practice?

For each cue-specific model:

- `350ms - 50ms` estimates encoding-duration improvement. Negative values mean
  lower error at 350 ms.
- `Late change - early change` tests whether this duration effect changes with
  practice. A positive value means that an early 350-ms advantage becomes
  smaller during later sessions.

```{r cue-specific-duration-by-session-contrasts}
duration_session_test <- function(model) {
  duration_means <- emmeans(
    model,
    ~ DurationF | Session_z,
    at = list(Session_z = c(early_z, late_z))
  )

  duration_changes <- contrast(
    duration_means,
    method = list(
      "350ms - 50ms" = c(-1, 0, 0, 0, 0, 0, 1)
    ),
    by = "Session_z"
  )

  practice_change <- contrast(
    duration_changes,
    method = list(
      "Late change - early change" = c(-1, 1)
    ),
    by = NULL
  )

  list(
    duration_changes = summary(
      duration_changes,
      infer = c(TRUE, TRUE)
    ),
    practice_change = summary(
      practice_change,
      infer = c(TRUE, TRUE)
    )
  )
}

duration_session_results <- lapply(
  cue_models,
  duration_session_test
)

for (model_name in names(duration_session_results)) {
  cat("\n\n###", model_name, "\n\n")
  print(duration_session_results[[model_name]]$duration_changes)
  print(duration_session_results[[model_name]]$practice_change)
}
```



# Redundancy Gain Function
This plot isolates redundancy gain: Redundancy Gain = (NR error - R error)
Theoretically important because the primary research question concerns whether 
redundancy gain changes as exposure duration increases, rather than just whether overall performance improves with duration.
Encoding-rate/perceptual accounts predict that non-redundant items should 
gradually "catch up" with increased exposure duration, producing a decreasing 
redundancy-gain function
In contrast, mnemonic/post-encoding accounts predict that redundancy
gain should emerge rapidly and remain relatively stable across durations.

Thus, the shape of the redundancy-gain function directly visualizes
whether redundancy effects themselves are exposure-duration dependent.

Current descriptive pattern:

# NOTE:
This figure is descriptive only. Bayesian hierarchical models are required to 
formally assess the magnitude and uncertainty of the Cue × Duration interaction.

```{r redundancy-gain}
gain_d <- d %>%
  group_by(ID, DurationF, CueType) %>%
  summarise(
    mean_abs = mean(AbsErr),
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = CueType,
    values_from = mean_abs
  ) %>%
  mutate(
    RedundancyGain = NR - R
  )
ggplot(gain_d,
       aes(DurationF, RedundancyGain,
           group = ID)) +
  geom_line(alpha = .5) +
  geom_point(alpha = .7) +
  stat_summary(
    aes(group = 1),
    fun = mean,
    geom = "line",
    linewidth = 1.5,
    colour = "black"
  ) +
  stat_summary(
    aes(group = 1),
    fun = mean,
    geom = "point",
    size = 3,
    colour = "black"
  ) +
  labs(
    y = "Redundancy Gain (NR - R)",
    x = "Exposure Duration"
  )
```

# INITIAL BAYESIAN HIERARCHICAL PRECISION MODEL
Goal:
First inferential test of whether redundancy gain depends on exposure duration.

DV
- Absolute circular response error (AbsErr)

Fixed effects:
- CueType: tests overall redundancy gain (R vs NR)
- DurationF: tests whether overall precision changes across exposure durations
- CueType × DurationF: critical interaction term testing whether redundancy gain 
changes as a function of exposure duration
- SessionF: included as a fixed-effect control factor to account for 
practice/learning effects across sessions

Random effects:
- varying intercepts by participant
- varying CueType slopes by participant
Random-effects structure allows:
- baseline precision differences between participants
- participant-specific redundancy gain magnitude

Distributional family:
- Student-t likelihood was chosen because trial-level response-error 
distributions contain heavy tails and occasional large-error trials, which 
violate normality assumptions.

Current modelling rationale:

- Duration and Session are initially treated as categorical factors to avoid 
imposing assumptions about monotonicity or functional form.

Further analyses:
- continuous duration/session models
- logarithmic/asymptotic duration functions
- model comparison (categorical vs continuous formulations)
- practical magnitude assessment of Cue × Duration interaction

Theoretically:
- encoding-rate/perceptual accounts predict duration-dependent redundancy gain
- mnemonic/post-encoding accounts predict relatively stable redundancy gain 
across durations
Thus, CueType × DurationF is the primary theoretical parameter of interest.

```{r model1, message=FALSE}
m1 <- brm(
  AbsErr ~ CueType * DurationF + SessionF +
    (1 + CueType | ID),
  data = d,
  family = student(),
  prior = c(
    prior(normal(40, 20), class = Intercept),
    prior(normal(0, 10), class = b),
    prior(exponential(1), class = sigma),
    prior(exponential(1), class = sd)
  ),
  chains = 4,
  cores = 4,
  iter = 4000,
  warmup = 2000,
  control = list(
    adapt_delta = .95,
    max_treedepth = 15
  ),
  seed = 1234
)

saveRDS(m1, "m1_categorical_sessionfactor.rds")
```

# Model Summary (seed 1234)
 Family: student 
  Links: mu = identity 
Formula: AbsErr ~ CueType * DurationF + SessionF + (1 + CueType | ID) 
   Data: d (Number of observations: 12599) 
  Draws: 4 chains, each with iter = 4000; warmup = 2000; thin = 1;
         total post-warmup draws = 8000

Multilevel Hyperparameters:
~ID (Number of levels: 3) 
                        Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
sd(Intercept)               2.07      0.85     0.95     4.22 1.00     5399     5513
sd(CueTypeR)                1.68      0.78     0.66     3.65 1.00     4774     5356
cor(Intercept,CueTypeR)    -0.12      0.39    -0.78     0.64 1.00     5005     5291

Regression Coefficients:
                        Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
Intercept                  18.02      1.44    15.26    20.97 1.00     3038     4063
CueTypeR                   -6.97      1.28    -9.42    -4.32 1.00     2949     3974
DurationF100ms              1.09      0.83    -0.55     2.70 1.00     3094     4615
DurationF150ms             -2.38      0.83    -3.99    -0.76 1.00     3070     4613
DurationF200ms             -2.01      0.80    -3.56    -0.44 1.00     2883     4493
DurationF250ms             -1.66      0.78    -3.20    -0.12 1.00     3014     4723
DurationF300ms             -2.90      0.74    -4.35    -1.43 1.00     2770     4118
DurationF350ms             -2.46      0.77    -3.94    -0.96 1.00     2823     4334
SessionF2                   0.03      0.52    -0.98     1.02 1.00     4048     5713
SessionF3                  -0.96      0.49    -1.91    -0.02 1.00     3720     5482
SessionF4                  -1.10      0.48    -2.05    -0.18 1.00     3682     5628
SessionF5                  -1.15      0.49    -2.11    -0.19 1.00     3747     5029
SessionF6                  -1.27      0.48    -2.20    -0.31 1.00     3659     5577
SessionF7                  -0.44      0.49    -1.40     0.51 1.00     3620     5043
SessionF8                  -1.14      0.49    -2.09    -0.17 1.00     3703     5314
SessionF9                  -1.46      0.49    -2.42    -0.49 1.00     3679     5214
SessionF10                 -1.51      0.48    -2.46    -0.57 1.00     3863     5762
CueTypeR:DurationF100ms    -1.88      0.96    -3.74     0.01 1.00     3198     4557
CueTypeR:DurationF150ms     1.59      0.96    -0.28     3.44 1.00     3169     5032
CueTypeR:DurationF200ms     0.68      0.93    -1.12     2.49 1.00     2927     4508
CueTypeR:DurationF250ms    -0.10      0.90    -1.85     1.66 1.00     2873     4608
CueTypeR:DurationF300ms     1.74      0.87     0.02     3.46 1.00     2834     4754
CueTypeR:DurationF350ms     1.56      0.90    -0.21     3.35 1.00     2885     4588

Further Distributional Parameters:
      Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
sigma     8.52      0.13     8.27     8.77 1.00     7641     6496
nu        1.20      0.02     1.16     1.24 1.00     8467     5890

Draws were sampled using sample(hmc). For each parameter, Bulk_ESS
and Tail_ESS are effective sample size measures, and Rhat is the potential
scale reduction factor on split chains (at convergence, Rhat = 1).


```{r model1-summary}
m1 <- readRDS("m1_categorical_sessionfactor.rds")
summary(m1)
#save.image("S1_analysis_workspace.RData")
```

# Posterior Diagnostics

```{r diagnostics}
plot(m1, variable = "^b_", regex = TRUE)
plot(m1, variable = "^sd_", regex = TRUE)
pp_check(m1, ndraws = 100)
```

# Conditional Effects

## Cue × Duration

```{r cond-effects}
conditional_effects(
  m1,
  effects = "DurationF:CueType"
)
```

# Posterior Estimates

```{r posterior-estimates}
fixef(m1)
```

# Model 1 Sum coding: Summed reference

```{r Model1Sum}
contrasts(d$CueType) <- contr.sum(2)
contrasts(d$DurationF) <- contr.sum(7)
contrasts(d$SessionF) <- contr.sum(length(levels(d$SessionF)))

m1_sum <- brm(
  AbsErr ~ CueType * DurationF + SessionF +
    (1 + CueType | ID),
  data = d,
  family = student(),
  prior = c(
    prior(normal(40, 20), class = Intercept),
    prior(normal(0, 10), class = b),
    prior(exponential(1), class = sigma),
    prior(exponential(1), class = sd)
  ),
  chains = 4,
  cores = 4,
  iter = 4000,
  warmup = 2000,
  control = list(adapt_delta = .95, max_treedepth = 15),
  seed = 1234
)

saveRDS(m1_sum, "m1_sum.rds")
```

```{r model1sum-summary}
m1_sum <- readRDS("m1_sum.rds")
summary(m1_sum)

#save.image("S1_analysis_workspace.RData")
```


```{r cond-effects}
conditional_effects(
  m1_sum
)
```

```{r posterior-estimates}
fixef(m1_sum)
emm_cue_duration <- emmeans(
  m1_sum,
  ~ CueType * DurationF
)

emm_cue_duration

gain_by_duration <- contrast(
  emm_cue_duration,
  method = list("NR - R" = c(1, -1)),
  by = "DurationF"
)

gain_by_duration

gain_df <- as.data.frame(confint(gain_by_duration))

ggplot(gain_df, aes(x = DurationF, y = estimate)) +
  geom_line(aes(group = 1), linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = lower.HPD, ymax = upper.HPD), width = .15) +
  labs(
    x = "Exposure Duration",
    y = "Estimated Redundancy Gain (NR - R)"
  ) +
  theme_bw()
```

# MODEL 2: CATEGORICAL INTERACTION-LEARNING MODEL

Goal: Tests whether the CueType × DurationF pattern changes across sessions.

Compared with m1_sum:
- m1_sum controls for SessionF as a categorical nuisance factor.
- m2_sum tests whether session/practice modulates the duration-dependence of 
redundancy gain.

Critical term:
CueType × DurationF × SessionF

Interpretation:
If this three-way interaction is meaningfully large, then practice changes how 
redundancy gain varies across exposure duration.

If small/uncertain, then practice may improve overall performance without 
substantially changing the redundancy-duration relation.

```{r model2, message=FALSE}
m2_sessionz <- brm(
  AbsErr ~ CueType * DurationF * Session_z +
    (1 + CueType | ID),
  data = d,
  family = student(),
  prior = c(
    prior(normal(40, 20), class = Intercept),
    prior(normal(0, 10), class = b),
    prior(exponential(1), class = sigma),
    prior(exponential(1), class = sd)
  ),
  chains = 4,
  cores = 4,
  iter = 4000,
  warmup = 2000,
  control = list(
    adapt_delta = 0.99,
    max_treedepth = 15
  ),
  seed = 2346
)

saveRDS(m2_sessionz, "m2_sessionz_cue_duration_learning.rds")
```

# Model 2 Summary

```{r model2-summary}
m2_sessionz <- readRDS("m2_sessionz_cue_duration_learning.rds")
summary(m2_sessionz)

```

# Estimated marginal means: Cue × Duration at average session
```{r}
emm_m2 <- emmeans(

  m2_sessionz,

  ~ CueType * DurationF,

  at = list(Session_z = 0)

)

emm_m2

confint(emm_m2)
```

# Posterior redundancy gain estimates at average session
```{r}
gain_m2 <- contrast(
  emm_m2,
  method = list("NR - R" = c(1, -1)),
  by = "DurationF"
)

gain_m2
confint(gain_m2)
```

# Convert redundancy gain estimates to dataframe and plot
```{r}
gain_m2_df <- as.data.frame(confint(gain_m2))

names(gain_m2_df)

ggplot(gain_m2_df, aes(x = DurationF, y = estimate)) +
  geom_line(aes(group = 1), linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = lower.HPD, ymax = upper.HPD),
    width = .15
  ) +
  labs(
    x = "Exposure Duration",
    y = "Estimated Redundancy Gain (NR - R)",
    title = "Posterior Redundancy Gain at Average Session"
  ) +
  theme_bw()
```


# Session learning visualization: predicted AbsErr across Session_z
```{r}
session_grid <- expand.grid(
  Session_z = seq(min(d$Session_z), max(d$Session_z), length.out = 100),
  CueType = levels(d$CueType),
  DurationF = levels(d$DurationF)
)

pred_session <- fitted(
  m2_sessionz,
  newdata = session_grid,
  re_formula = NA,
  summary = TRUE
)

session_pred_df <- bind_cols(
  session_grid,
  as.data.frame(pred_session)
)

ggplot(
  session_pred_df,
  aes(x = Session_z, y = Estimate, colour = CueType)
) +
  geom_line(linewidth = 1) +
  geom_ribbon(
    aes(ymin = Q2.5, ymax = Q97.5, fill = CueType),
    alpha = .15,
    colour = NA
  ) +
  facet_wrap(~DurationF) +
  labs(
    x = "Session progression (z-scored)",
    y = "Predicted Absolute Error (deg)",
    title = "Predicted Session Learning by Cue and Duration"
  ) +
  theme_bw()
```

# Optional: easier session plot averaged across duration
```{r}
session_grid_avg <- expand.grid(
  Session_z = seq(min(d$Session_z), max(d$Session_z), length.out = 100),
  CueType = levels(d$CueType),
  DurationF = levels(d$DurationF)
)

pred_session_avg <- fitted(
  m2_sessionz,
  newdata = session_grid_avg,
  re_formula = NA,
  summary = TRUE
)

session_avg_df <- bind_cols(
  session_grid_avg,
  as.data.frame(pred_session_avg)
) %>%
  group_by(Session_z, CueType) %>%
  summarise(
    Estimate = mean(Estimate),
    Q2.5 = mean(Q2.5),
    Q97.5 = mean(Q97.5),
    .groups = "drop"
  )

ggplot(
  session_avg_df,
  aes(x = Session_z, y = Estimate, colour = CueType)
) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(
    aes(ymin = Q2.5, ymax = Q97.5, fill = CueType),
    alpha = .15,
    colour = NA
  ) +
  labs(
    x = "Session progression (z-scored)",
    y = "Predicted Absolute Error (deg)",
    title = "Predicted Session Learning Averaged Across Durations"
  ) +
  theme_bw()
```

# Model comparison: m1_sum vs m2_sessionz
```{r}
loo_m1_sum <- loo(m1_sum)
loo_m2_sessionz <- loo(m2_sessionz)

loo_compare(loo_m1_sum, loo_m2_sessionz)
```

```{r}
loo_m1_sum <- loo(m1_sum, moment_match = TRUE)
loo_m2_sessionz <- loo(m2_sessionz, moment_match = TRUE)

loo_compare(loo_m1_sum, loo_m2_sessionz)
```

```{r}
saveRDS(loo_m1_sum, "loo_m1_sum.rds")
saveRDS(loo_m2_sessionz, "loo_m2_sessionz.rds")
saveRDS(gain_m2, "gain_m2_sessionz.rds")
```

# RT Model

Secondary analysis.

```{r rt-model, message=FALSE}

m_rt <- brm(

  logRT ~ CueType * Duration + Session_z +

    (1 + CueType | ID),

  data = d,

  family = gaussian(),

  prior = c(

    prior(normal(6, 1), class = Intercept),

    prior(normal(0, .2), class = b),

    prior(exponential(1), class = sigma),

    prior(exponential(1), class = sd)

  ),

  chains = 4,

  cores = 4,

  iter = 4000,

  control = list(

    adapt_delta = .95

  ),

  seed = 2468

)

```

# RT Summary

```{r rt-summary}

summary(m_rt)

```

# Save Models

```{r save-models}

saveRDS(m1, "m1_precision_main.rds")

saveRDS(m2, "m2_precision_learning.rds")

saveRDS(m_rt, "m_rt.rds")

```