---
title: "Study 1 AbsErr Bayes — discrete SessionF"
author: "Yu Liu"
date: "11th May 2026"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
editor_options: 
  markdown: 
    wrap: 72
---

# Overview

Companion to `s1_smallN_analysis.Rmd`, identical pipeline except
**session is categorical** (`SessionF`) instead of continuous
`Session_z`.

Model difference (Sections 3–4):

- Continuous: `AbsErr / cirSD ~ CueType * DurationF * Session_z + (1 | SessionF)`
- Discrete: **SignedErr** von Mises mixture ladder (Stan, Section 3); cirSD
  bootstrap + Student-t (Section 4, legacy summary)

Analysis order:

1.  Load and prepare trial data
2.  Descriptive figures (circular SD and RT by participant)
3.  Bayesian SignedErr VM mixture ladder (Stan: L0 / L1 / L2)
4.  Bayesian cirSD primary (same design; bootstrap cell replicates)
5.  Pre-mixture: AbsErr Session 1 vs 10 by duration
6.  Pre-mixture: cirSD Session 1 vs 10 (nonparametric bootstrap companion)
7.  Circular mixture cell fits — 3-comp (target + guess + swap)
8.  Circular mixture cell fits — 2-comp (target + guess; no swap)
9.  Simple von Mises κ comparison: 14-κ (cue × duration) vs 2-κ (cue only)
10. Mixture 3-comp: κ free per cell vs κ tied within cue (β, γ free per cell)
11. Mixture 2-comp: κ free per cell vs κ tied within cue (γ free per cell)

Style: linear commented code; loops over P01–P04 (AQ, HC, YILIU, YDL).

# Setup

```{r setup, message=FALSE, warning=FALSE}
# Core packages for this pipeline.
library(tidyverse)   # data wrangling + ggplot2
library(nlme)        # frequentist mixed models (Appendix)
library(brms)        # Bayesian regression
library(bayesplot)   # posterior predictive checks
library(tidybayes)   # gather posterior draws from emmeans
library(emmeans)     # estimated marginal means / contrasts
library(patchwork)   # combine ggplot panels
library(loo)         # LOO / WAIC model comparison
library(cmdstanr)    # Stan VM mixture ladder (Section 3)
library(posterior)   # Stan draw summaries
library(circular)    # circular SD for signed colour errors

theme_set(theme_bw())
options(mc.cores = parallel::detectCores())
set.seed(1234)  # reproducible bootstrap / Stan seeds where used
```

# 1. Load and prepare data

```{r load-data}
# Model-ready trial table from MATLAB preprocessing.
d <- read_csv("S1_smallN_model_ready.csv", show_col_types = FALSE)

# Quick sanity checks: expected columns and no empty file.
stopifnot(nrow(d) > 0)
stopifnot(all(c("ID", "CueType", "Duration", "Session", "Session_z",
                "AbsErr", "SignedErr") %in% names(d)))

glimpse(d)
```

```{r prepare-factors}
# Publication participant IDs (CSV / mixture files still use raw lab IDs).
id_raw <- c("AQ", "HC", "YILIU", "YDL")
id_pub <- c("P01", "P02", "P03", "P04")
names(id_pub) <- id_raw
participants <- unname(id_pub)

# Set factor levels explicitly so contrasts and plots keep a fixed order.
d <- d %>%
  mutate(
    ID = factor(id_pub[as.character(ID)], levels = participants),
    CueType = factor(CueType, levels = c("NR", "R")),
    DurationF = factor(
      Duration,
      levels = c("50ms", "100ms", "150ms", "200ms", "250ms", "300ms", "350ms")
    ),
    Grouping = factor(Grouping),
    SessionF = factor(Session),
    SessionPhase = factor(SessionPhase, levels = c("Early", "Middle", "Late")),
    logRT = log(RT)
  )

# One data frame per complete participant (object names = raw ID; ID column = P0x).
d_AQ <- droplevels(filter(d, ID == "P01"))
d_HC <- droplevels(filter(d, ID == "P02"))
d_YILIU <- droplevels(filter(d, ID == "P03"))
d_YDL <- droplevels(filter(d, ID == "P04"))
stopifnot(all(participants %in% levels(d$ID)))
stopifnot(n_distinct(d_YDL$Session) == 10)

# Confirm design levels.
levels(d$CueType)
levels(d$DurationF)
levels(d$SessionF)
levels(d$ID)
```

# 2. Descriptive figures

Four vertically stacked panels (one participant each), shared encoding-duration
axis. **Circular SD:** pooled cell estimate (diamond) with bootstrap 95% CI;
semi-transparent **AbsErr** trial points behind. **RT:** cell mean with bootstrap
95% CI; trial-level RT points behind.

```{r descriptives-helpers, include=FALSE}
# Circular SD of signed colour errors (degrees).
circ_sd_deg <- function(x) {
  x <- x[is.finite(x)]
  if (length(x) < 2) return(NA_real_)
  as.numeric(sd.circular(circular(x, units = "degrees"))) * (180 / pi)
}

# Bootstrap percentile CI for any cell statistic.
boot_cell_ci <- function(x, stat_fn, B = 2000, conf = 0.95) {
  x <- x[is.finite(x)]
  n <- length(x)
  if (n < 10) {
    est <- stat_fn(x)
    return(tibble(estimate = est, lower = NA_real_, upper = NA_real_))
  }
  obs <- stat_fn(x)
  boots <- vapply(seq_len(B), function(b) {
    stat_fn(sample(x, n, replace = TRUE))
  }, numeric(1))
  alpha <- (1 - conf) / 2
  tibble(
    estimate = obs,
    lower = unname(quantile(boots, alpha, na.rm = TRUE)),
    upper = unname(quantile(boots, 1 - alpha, na.rm = TRUE))
  )
}

cue_cols <- c("NR" = "#6B9AC4", "R" = "#C75146")
cue_x_offset <- c("NR" = -5, "R" = 5)

# Times New Roman for publication; fall back to Times if unavailable (e.g. Linux CI).
desc_font <- if (
  requireNamespace("systemfonts", quietly = TRUE) &&
    "Times New Roman" %in% unique(systemfonts::system_fonts()$family)
) {
  "Times New Roman"
} else {
  "Times"
}

theme_desc <- theme_bw(base_size = 11, base_family = desc_font) +
  theme(
    panel.grid.minor = element_blank(),
    strip.placement = "inside",
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", hjust = 0.5, vjust = 0),
    legend.position = "bottom",
    legend.title = element_text(face = "bold"),
    plot.title = element_blank(),
    text = element_text(family = desc_font),
    axis.title = element_text(family = desc_font),
    axis.text = element_text(family = desc_font),
    legend.text = element_text(family = desc_font)
  )

desc_x_jitter <- 2
# Pad x limits so cue offsets (+ jitter / error-bar caps) are not clipped at 50 / 350 ms.
desc_x_pad <- max(abs(cue_x_offset)) + desc_x_jitter / 2 + 6

desc_x_scale <- scale_x_continuous(
  breaks = seq(50, 350, by = 50),
  limits = c(50, 350),
  expand = expansion(mult = c(0.08, 0.08))
)

desc_x_scale_cir <- scale_x_continuous(
  breaks = seq(50, 350, by = 50),
  limits = c(50 - desc_x_pad, 350 + desc_x_pad),
  expand = expansion(mult = c(0, 0))
)
```

```{r descriptives-data, message=FALSE, warning=FALSE}
B_desc <- 2000L
set.seed(1234)

d_desc <- d %>%
  mutate(
    Duration_ms = as.numeric(gsub("ms", "", as.character(DurationF)))
  )

# Pooled cell cirSD + bootstrap 95% CI (main indicator).
circ_cell <- d_desc %>%
  group_by(ID, CueType, DurationF, Duration_ms) %>%
  group_modify(~ boot_cell_ci(.x$SignedErr, circ_sd_deg, B = B_desc)) %>%
  ungroup() %>%
  mutate(
    Duration_x = Duration_ms + unname(cue_x_offset[as.character(CueType)])
  )

d_desc_cir <- d_desc %>%
  mutate(
    Duration_x = Duration_ms + unname(cue_x_offset[as.character(CueType)])
  )

# Pooled cell mean RT + bootstrap 95% CI.
rt_cell <- d_desc %>%
  group_by(ID, CueType, DurationF, Duration_ms) %>%
  group_modify(~ boot_cell_ci(.x$RT, mean, B = B_desc)) %>%
  ungroup()
```

```{r descriptives-cirsd, fig.width=4.2, fig.height=9, message=FALSE, warning=FALSE}
# Precision: circular SD (degrees) by cue and duration.
ggplot() +
  geom_point(
    data = d_desc_cir,
    aes(x = Duration_x, y = AbsErr, colour = CueType),
    alpha = 0.20,
    size = 1,
    stroke = 0,
    position = position_jitter(width = desc_x_jitter, height = 0, seed = 1234)
  ) +
  geom_line(
    data = circ_cell,
    aes(x = Duration_x, y = estimate, colour = CueType, group = CueType),
    linewidth = 0.3,
    linetype = "dashed"
  ) +
  geom_errorbar(
    data = circ_cell,
    aes(x = Duration_x, ymin = lower, ymax = upper, colour = 'black'),
    width = 3,
    linewidth = 0.5
  ) +
  geom_point(
    data = circ_cell,
    aes(x = Duration_x, y = estimate, colour = CueType),
    shape = 23,
    fill = "white",
    size = 2.5,
    stroke = 0.45
  ) +
  facet_wrap(~ID, ncol = 1, scales = "fixed") +
  desc_x_scale_cir +
  scale_colour_manual(values = cue_cols, name = "Cue") +
  labs(x = "Encoding duration (ms)", y = "Circular SD (°)") +
  theme_desc
```

```{r descriptives-rt, fig.width=4.2, fig.height=9, message=FALSE, warning=FALSE}
# RT: mean response time with bootstrap 95% CI; trial-level points behind.
ggplot() +
  geom_point(
    data = d_desc,
    aes(x = Duration_ms, y = RT, colour = CueType),
    alpha = 0.10,
    size = 0.55,
    stroke = 0,
    position = position_jitter(width = 5, height = 0, seed = 1234)
  ) +
  geom_line(
    data = rt_cell,
    aes(x = Duration_ms, y = estimate, colour = CueType, group = CueType),
    linewidth = 0.35,
    linetype = "dashed"
  ) +
  geom_errorbar(
    data = rt_cell,
    aes(x = Duration_ms, ymin = lower, ymax = upper, colour = CueType),
    width = 6,
    linewidth = 0.35
  ) +
  geom_point(
    data = rt_cell,
    aes(x = Duration_ms, y = estimate, colour = CueType),
    shape = 21,
    fill = "white",
    size = 1.6,
    stroke = 0.45
  ) +
  facet_wrap(~ID, ncol = 1, scales = "fixed") +
  desc_x_scale +
  scale_colour_manual(values = cue_cols, name = "Cue") +
  labs(x = "Encoding duration (ms)", y = "Response time (ms)") +
  theme_desc
```

# 3. Bayesian primary analysis: von Mises mixture ladder (SignedErr)

Generative models on **SignedErr** (radians), fit with **Stan** (`cmdstanr`).
One hierarchical model per participant × level × κ structure.

| Level | Likelihood | Session nesting |
|-------|------------|-----------------|
| **L0** | VM(0, κ) only | κ: Cue×Duration (14κ) or Cue (2κ) + Session offsets |
| **L1** | VM + uniform guess (γ) | κ as L0; γ ~ Cue×Duration + Session |
| **L2** | VM + swap (3 bumps) + uniform | κ as L0; β, γ ~ Cue×Duration + Session |

Compare **2κ vs 14κ** within each level (LOO/WAIC). Compare **L0 vs L1 vs L2**
using the preferred κ structure per level (expect **2κ** at L1/L2 once γ and β are included).

**Runtime guide** (4 chains, 2000 warmup + 2000 sampling, ~4200 trials/participant):

- **L0:** ~5–15 min per fit → ~40–120 min for all 8 L0 fits (4 IDs × 2κ/14κ)
- **L1:** ~15–30 min per fit
- **L2:** ~5–10 min per fit → ~40–80 min for all 8 L2 fits (4 IDs × 2κ/14κ)

Fits cache under `bayes_primary_discrete/vm_stan/`. Console shows chain
progress (`refresh = 100`).

**Parallel fitting:** each level chunk fits 8 models (4 participants × 2κ/14κ)
in parallel. Default: `n_workers = floor(cores / chains)` with 4 chains per
fit (e.g. 10 cores → 2 workers × 4 chains). Set `vm_n_workers` in setup to
override (use `1` for strictly sequential models).

Reliability check: Sections 7–11 (MLE mixture cells, 14κ vs 2κ) unchanged.

```{r vm-ladder-setup, message=FALSE, warning=FALSE}
source("s1_vm_ladder_stan.R")
vm_init_plot_defaults()

dir.create("bayes_primary_discrete/vm_stan", showWarnings = FALSE, recursive = TRUE)

vm_raw_ids <- c("AQ", "HC", "YILIU", "YDL")
vm_id_pub <- c(AQ = "P01", HC = "P02", YILIU = "P03", YDL = "P04")

vm_data <- vm_participant_data(d)

# Stan settings (reduce iter for a quick smoke test if needed).
vm_chains <- 4
vm_iter_warmup <- 2000
vm_iter_sampling <- 2000
vm_refresh <- 100
# Parallel model workers (NULL = auto: floor(detectCores() / vm_chains)).
vm_n_workers <- NULL

vm_parallel_plan <- vm_stan_parallel_plan(
  chains_per_fit = vm_chains,
  n_workers = vm_n_workers
)
print_vm_stan_parallel_plan(vm_parallel_plan)
```

## 3.1 L0 — von Mises only (indicative κ)

```{r vm-ladder-L0, message=TRUE, warning=TRUE}
vm_L0_batch <- fit_vm_ladder_batch(
  vm_data = vm_data,
  level = "L0",
  raw_ids = vm_raw_ids,
  id_pub = vm_id_pub,
  kappa_structs = c("2k", "14k"),
  n_workers = vm_n_workers,
  chains = vm_chains,
  iter_warmup = vm_iter_warmup,
  iter_sampling = vm_iter_sampling,
  refresh = vm_refresh,
  seed = 1234
)
vm_L0 <- vm_L0_batch$fits
vm_L0_loo <- vm_L0_batch$compare %>% arrange(ID, delta_loo)

write_csv(vm_L0_loo, "bayes_primary_discrete/vm_ladder_L0_loo.csv")
saveRDS(list(fits = vm_L0, compare = vm_L0_loo), "bayes_primary_discrete/vm_ladder_L0_results.rds")

vm_L0_loo

# κ by Cue×Duration (posterior median), averaged over sessions — indicative only.
vm_L0_kappa <- bind_rows(lapply(vm_raw_ids, function(raw) {
  bind_rows(lapply(names(vm_L0[[raw]]), function(mid) {
    extract_vm_cell_params(vm_L0[[raw]][[mid]]) %>%
      mutate(ID = vm_id_pub[[raw]], model_id = mid)
  }))
}))

ggplot(
  vm_summarise_cells_plot(vm_L0_kappa, "kappa"),
  aes(Duration_ms, estimate, ymin = lower, ymax = upper, colour = CueType, group = CueType)
) +
  geom_line(linewidth = 0.4) +
  geom_errorbar(width = 8, linewidth = 0.35) +
  geom_point(size = 1.5) +
  facet_grid(ID ~ model_id) +
  scale_x_continuous(breaks = seq(50, 350, 50)) +
  scale_colour_manual(values = cue_cols) +
  labs(
    x = "Encoding duration (ms)",
    y = expression(paste("Posterior ", kappa, " (median + 95% CrI)")),
    title = "L0 (indicative): precision by cue and duration"
  ) +
  theme_desc
```

## 3.2 L1 — VM + uniform guess

```{r vm-ladder-L1, message=TRUE, warning=TRUE}
vm_L1_batch <- fit_vm_ladder_batch(
  vm_data = vm_data,
  level = "L1",
  raw_ids = vm_raw_ids,
  id_pub = vm_id_pub,
  kappa_structs = c("2k", "14k"),
  n_workers = vm_n_workers,
  chains = vm_chains,
  iter_warmup = vm_iter_warmup,
  iter_sampling = vm_iter_sampling,
  refresh = vm_refresh,
  seed = 2234
)
vm_L1 <- vm_L1_batch$fits
vm_L1_loo <- vm_L1_batch$compare %>% arrange(ID, delta_loo)

write_csv(vm_L1_loo, "bayes_primary_discrete/vm_ladder_L1_loo.csv")
saveRDS(list(fits = vm_L1, compare = vm_L1_loo), "bayes_primary_discrete/vm_ladder_L1_results.rds")

vm_L1_loo

# γ (guess rate) by cell — primary composition story at L1.
vm_L1_gamma <- bind_rows(lapply(vm_raw_ids, function(raw) {
  bind_rows(lapply(names(vm_L1[[raw]]), function(mid) {
    extract_vm_cell_params(vm_L1[[raw]][[mid]]) %>%
      mutate(ID = vm_id_pub[[raw]], model_id = mid)
  }))
}))

ggplot(
  vm_summarise_cells_plot(vm_L1_gamma, "gamma"),
  aes(Duration_ms, estimate, ymin = lower, ymax = upper, colour = CueType, group = CueType)
) +
  geom_line(linewidth = 0.4) +
  geom_errorbar(width = 8, linewidth = 0.35) +
  geom_point(size = 1.5) +
  facet_grid(ID ~ model_id) +
  scale_x_continuous(breaks = seq(50, 350, 50)) +
  scale_colour_manual(values = cue_cols) +
  labs(
    x = "Encoding duration (ms)",
    y = "Posterior γ (median + 95% CrI)",
    title = "L1: guessing by cue and duration (session-nested model)"
  ) +
  theme_desc

ggplot(
  vm_summarise_cells_plot(vm_L1_gamma, "kappa"),
  aes(Duration_ms, estimate, ymin = lower, ymax = upper, colour = CueType, group = CueType)
) +
  geom_line(linewidth = 0.4) +
  geom_errorbar(width = 8, linewidth = 0.35) +
  geom_point(size = 1.5) +
  facet_grid(ID ~ model_id) +
  scale_x_continuous(breaks = seq(50, 350, 50)) +
  scale_colour_manual(values = cue_cols) +
  labs(
    x = "Encoding duration (ms)",
    y = expression(paste("Posterior ", kappa, " (median + 95% CrI)")),
    title = "L1: precision by cue and duration (session-nested model)"
  ) +
  theme_desc
```

## 3.3 L2 — VM + swap + uniform guess

L1 comparison suggested **2κ** is sufficient once γ is included; L2 fits both 2κ and 14κ
for a direct LOO check with swap (β) in the model. Cached 2κ loads instantly if already fit.

```{r vm-ladder-L2, message=TRUE, warning=TRUE}
vm_L2_batch <- fit_vm_ladder_batch(
  vm_data = vm_data,
  level = "L2",
  raw_ids = vm_raw_ids,
  id_pub = vm_id_pub,
  kappa_structs = c("2k", "14k"),
  n_workers = vm_n_workers,
  chains = vm_chains,
  iter_warmup = vm_iter_warmup,
  iter_sampling = vm_iter_sampling,
  refresh = vm_refresh,
  seed = 3234
)
vm_L2 <- vm_L2_batch$fits
vm_L2_loo <- vm_L2_batch$compare %>% arrange(ID, delta_loo)

write_csv(vm_L2_loo, "bayes_primary_discrete/vm_ladder_L2_loo.csv")
saveRDS(list(fits = vm_L2, compare = vm_L2_loo), "bayes_primary_discrete/vm_ladder_L2_results.rds")

vm_L2_loo

vm_L2_params <- bind_rows(lapply(vm_raw_ids, function(raw) {
  bind_rows(lapply(names(vm_L2[[raw]]), function(mid) {
    extract_vm_cell_params(vm_L2[[raw]][[mid]]) %>%
      mutate(ID = vm_id_pub[[raw]], model_id = mid)
  }))
}))

ggplot(
  vm_summarise_cells_plot(vm_L2_params, "gamma"),
  aes(Duration_ms, estimate, ymin = lower, ymax = upper, colour = CueType, group = CueType)
) +
  geom_line(linewidth = 0.4) +
  geom_errorbar(width = 8, linewidth = 0.35) +
  geom_point(size = 1.2) +
  facet_grid(ID ~ model_id) +
  scale_x_continuous(breaks = seq(50, 350, 50)) +
  scale_colour_manual(values = cue_cols) +
  labs(
    x = "Encoding duration (ms)", y = "Posterior γ (median + 95% CrI)",
    title = "L2: guessing by cue and duration (session-nested model)"
  ) +
  theme_desc

ggplot(
  vm_summarise_cells_plot(vm_L2_params, "beta"),
  aes(Duration_ms, estimate, ymin = lower, ymax = upper, colour = CueType, group = CueType)
) +
  geom_line(linewidth = 0.4) +
  geom_errorbar(width = 8, linewidth = 0.35) +
  geom_point(size = 1.2) +
  facet_grid(ID ~ model_id) +
  scale_x_continuous(breaks = seq(50, 350, 50)) +
  scale_colour_manual(values = cue_cols) +
  labs(
    x = "Encoding duration (ms)", y = "Posterior β (median + 95% CrI)",
    title = "L2: swap by cue and duration (session-nested model)"
  ) +
  theme_desc

ggplot(
  vm_summarise_cells_plot(vm_L2_params, "kappa"),
  aes(Duration_ms, estimate, ymin = lower, ymax = upper, colour = CueType, group = CueType)
) +
  geom_line(linewidth = 0.4) +
  geom_errorbar(width = 8, linewidth = 0.35) +
  geom_point(size = 1.2) +
  facet_grid(ID ~ model_id) +
  scale_x_continuous(breaks = seq(50, 350, 50)) +
  scale_colour_manual(values = cue_cols) +
  labs(
    x = "Encoding duration (ms)",
    y = expression(paste("Posterior ", kappa, " (median + 95% CrI)")),
    title = "L2: precision by cue and duration (session-nested model)"
  ) +
  theme_desc
```

## 3.4 Model comparison, cell parameters, and session effects

```{r vm-ladder-summary, message=FALSE, warning=FALSE}
vm_init_plot_defaults()

pick_best <- function(cmp, level) {
  if (is.null(cmp) || nrow(cmp) == 0) return(tibble())
  cmp %>%
    filter(grepl(paste0("^", level), model_id), kappa_struct == "2k") %>%
    group_by(ID) %>%
    slice_min(delta_loo, n = 1, with_ties = FALSE) %>%
    ungroup()
}

load_vm_loo <- function(level) {
  f <- file.path("bayes_primary_discrete", paste0("vm_ladder_", level, "_loo.csv"))
  if (file.exists(f)) read_csv(f, show_col_types = FALSE) else tibble()
}

vm_L0_loo <- if (exists("vm_L0_loo")) vm_L0_loo else load_vm_loo("L0")
vm_L1_loo <- if (exists("vm_L1_loo")) vm_L1_loo else load_vm_loo("L1")
vm_L2_loo <- if (exists("vm_L2_loo")) vm_L2_loo else load_vm_loo("L2")

vm_best_2k <- bind_rows(
  pick_best(vm_L0_loo, "L0") %>% mutate(ladder = "L0"),
  pick_best(vm_L1_loo, "L1") %>% mutate(ladder = "L1"),
  pick_best(vm_L2_loo, "L2") %>% mutate(ladder = "L2")
)

vm_cross_level <- bind_rows(
  lapply(vm_raw_ids, function(raw) {
    rows <- list()
    for (lv in c("L0", "L1", "L2")) {
      for (ks in c("2k", "14k")) {
        fit <- vm_load_fit_cache(raw, lv, ks)
        if (is.null(fit) || is.null(fit$loo)) next
        rows[[length(rows) + 1L]] <- tibble(
          ID = vm_id_pub[[raw]],
          raw_ID = raw,
          level = lv,
          model_id = paste0(lv, "_", ks),
          kappa_struct = ks,
          elpd_loo = fit$loo$estimates["elpd_loo", "Estimate"]
        )
      }
    }
    if (length(rows) == 0) return(NULL)
    bind_rows(rows) %>%
      group_by(ID) %>%
      mutate(delta_loo = -elpd_loo + max(elpd_loo)) %>%
      ungroup()
  })
)

# ---- full cell-level parameters (all models) ----
vm_all_cells <- vm_collect_all_cell_params(
  raw_ids = vm_raw_ids,
  id_pub = vm_id_pub,
  levels = c("L0", "L1", "L2"),
  kappa_structs = c("2k", "14k")
)
vm_all_cells_fmt <- vm_format_cell_params(vm_all_cells)

write_csv(vm_best_2k, "bayes_primary_discrete/vm_ladder_best_2k.csv")
write_csv(vm_cross_level, "bayes_primary_discrete/vm_ladder_cross_level_loo.csv")
write_csv(vm_all_cells_fmt, "bayes_primary_discrete/vm_ladder_all_cell_params.csv")

# ---- session offset parameters (S2–S10 vs S1) ----
vm_session_offsets <- vm_collect_session_offsets(
  raw_ids = vm_raw_ids,
  id_pub = vm_id_pub,
  levels = c("L0", "L1", "L2"),
  kappa_structs = c("2k", "14k")
)
write_csv(vm_session_offsets, "bayes_primary_discrete/vm_ladder_session_offsets.csv")

# ---- session change at cell level (S10 − S1) ----
vm_session_kappa <- vm_session_cell_change(vm_all_cells, "kappa")
vm_session_gamma <- vm_session_cell_change(vm_all_cells, "gamma")
vm_session_beta <- vm_session_cell_change(vm_all_cells, "beta")
vm_session_cell <- bind_rows(vm_session_kappa, vm_session_gamma, vm_session_beta)
write_csv(vm_session_cell, "bayes_primary_discrete/vm_ladder_session_cell_change.csv")

cat("=== LOO: best 2κ per level ===\n")
print(vm_best_2k %>% select(ID, ladder, model_id, elpd_loo, delta_loo))

cat("\n=== LOO: all models cross-level (per ID) ===\n")
print(vm_cross_level %>% arrange(ID, delta_loo))

cat("\n=== Full cell params: ", nrow(vm_all_cells_fmt), " rows exported ===\n")
print(
  vm_all_cells_fmt %>%
    filter(level %in% c("L1", "L2"), kappa_struct == "2k",
           CueType == "NR", Duration_ms %in% c(50, 350), SessionF == "1") %>%
    select(ID, level, model_id, CueType, Duration_ms, SessionF,
           kappa, kappa_lo, kappa_hi, gamma, gamma_lo, gamma_hi, beta, beta_lo, beta_hi) %>%
    arrange(ID, level, Duration_ms),
  n = 32
)

cat("\n=== Session offsets (median log/logit add-on, S2–S10) ===\n")
print(
  vm_session_offsets %>%
    filter(model_id %in% c("L1_2k", "L2_2k")) %>%
    group_by(ID, model_id, component) %>%
    summarise(
      median_offset = median(logit_or_log_median),
      max_abs = max(abs(logit_or_log_median)),
      .groups = "drop"
    ) %>%
    arrange(ID, model_id, component),
  n = 40
)

cat("\n=== Session cell change S10−S1 (NR, 50 & 350 ms, L1/L2 2κ) ===\n")
print(
  vm_session_cell %>%
    filter(kappa_struct == "2k", level %in% c("L1", "L2"),
           CueType == "NR", Duration_ms %in% c(50, 350)) %>%
    select(ID, level, parameter, Duration_ms, S1, S10, delta_S10_minus_S1, delta_lo, delta_hi) %>%
    arrange(ID, level, parameter, Duration_ms),
  n = 40
)
```

## 3.5 Theory contrasts (posterior draws)

Contrasts on **κ, γ, β** with 95% CrI: redundancy **gain** (R−NR), **duration**
(350−50 ms), and **session** (S10−S1). Primary models: L1/L2 with 2κ and 14κ.

```{r vm-ladder-contrasts, message=FALSE, warning=FALSE}
vm_theory <- vm_collect_theory_contrasts(
  raw_ids = vm_raw_ids,
  id_pub = vm_id_pub,
  levels = c("L1", "L2"),
  kappa_structs = c("2k", "14k")
)
write_csv(vm_theory, "bayes_primary_discrete/vm_ladder_theory_contrasts.csv")

vm_theory_focus <- vm_theory %>%
  filter(
    (grepl("^gain_R_minus_NR", contrast) & grepl("(50|350)ms_S", contrast)) |
      grepl("^dur350_minus_50", contrast) |
      grepl("^session_S10_minus_S1", contrast)
  ) %>%
  mutate(
    param = case_when(
      grepl("_kappa_", contrast) ~ "kappa",
      grepl("_gamma_", contrast) ~ "gamma",
      grepl("_beta_", contrast) ~ "beta"
    ),
    effect = case_when(
      grepl("^gain", contrast) ~ "R − NR",
      grepl("^dur350", contrast) ~ "350 − 50 ms",
      grepl("^session", contrast) ~ "S10 − S1",
      TRUE ~ NA_character_
    ),
    duration_ms = as.integer(str_extract(contrast, "(\\d+)(?=ms)")),
    session = str_extract(contrast, "S\\d+$"),
    cue = case_when(
      grepl("^gain", contrast) ~ NA_character_,
      grepl("_NR_", contrast) ~ "NR",
      grepl("_R_", contrast) ~ "R",
      TRUE ~ NA_character_
    )
  )

vm_theory_focus %>%
  filter(kappa_struct == "2k") %>%
  select(ID, level, model_id, effect, param, cue, duration_ms, session,
         median, lower, upper) %>%
  arrange(ID, level, effect, param, cue, duration_ms, session)

ggplot(
  vm_theory_focus %>%
    filter(
      kappa_struct == "2k",
      param %in% c("gamma", "kappa"),
      effect == "350 − 50 ms",
      session == "S1"
    ),
  aes(cue, median, ymin = lower, ymax = upper, colour = param, group = param)
) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_errorbar(width = 0.15, linewidth = 0.35, position = position_dodge(width = 0.4)) +
  geom_point(size = 2.5, position = position_dodge(width = 0.4)) +
  facet_grid(ID ~ level) +
  scale_colour_manual(values = c(kappa = "#2166AC", gamma = "#B2182B")) +
  labs(
    title = "Duration contrast (350 − 50 ms): κ and γ [L1/L2, 2κ, Session 1]",
    x = "Cue", y = "Posterior contrast (95% CrI)", colour = "Parameter"
  ) +
  theme_desc

ggplot(
  vm_theory_focus %>%
    filter(
      kappa_struct == "2k",
      param == "gamma",
      effect == "R − NR",
      session == "S1",
      duration_ms %in% c(50L, 350L)
    ),
  aes(duration_ms, median, ymin = lower, ymax = upper, group = level, colour = level)
) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_line(linewidth = 0.5) +
  geom_errorbar(width = 12, linewidth = 0.35) +
  geom_point(size = 2) +
  facet_wrap(~ID) +
  scale_x_continuous(breaks = c(50, 350)) +
  labs(
    title = "Redundancy gain contrast (γ_R − γ_NR) [2κ]",
    x = "Duration (ms)", y = "Posterior contrast (95% CrI)"
  ) +
  theme_desc
```

## 3.6 Posterior predictive checks (SignedErr density)

Observed **trial histograms** (all sessions pooled) overlaid with **L0 / L1 / L2**
mixture PDFs (solid) and **component curves** (dashed). Panels labelled by
participant, cue, and duration (50 & 350 ms, **2κ**).

```{r vm-ladder-ppc, fig.width=14, fig.height=12, message=FALSE, warning=FALSE}
ppc_level_cols <- c(L0 = "#7570B3", L1 = "#1B9E77", L2 = "#D95F02")
ppc_component_lty <- c(
  target = "dashed",
  uniform = "dotdash",
  swap1 = "dotted",
  swap2 = "longdash",
  swap3 = "twodash"
)

vm_ppc_plot_cell <- function(raw, id_label, cue, duration) {
  ppc <- vm_ppc_compare_levels(raw, cue, duration, kappa_struct = "2k")
  if (is.null(ppc)) return(NULL)

  dur_ms <- as.integer(gsub("ms", "", duration))
  panel_title <- paste(id_label, cue, paste0(dur_ms, " ms"), sep = " · ")

  comp <- ppc$components %>%
    mutate(
      level = factor(level, levels = c("L0", "L1", "L2")),
      component = factor(
        component,
        levels = c("target", "uniform", "swap1", "swap2", "swap3")
      )
    )

  ggplot() +
    geom_histogram(
      aes(x = ppc$observed, y = after_stat(density)),
      binwidth = 10, boundary = -180, fill = "grey85", colour = "white",
      alpha = 0.85
    ) +
    geom_line(
      data = ppc$mixture %>% mutate(level = factor(level, levels = c("L0", "L1", "L2"))),
      aes(x = theta_deg, y = density, colour = level),
      linewidth = 0.9, inherit.aes = FALSE
    ) +
    geom_line(
      data = comp,
      aes(x = theta_deg, y = density, colour = level, linetype = component),
      linewidth = 0.55, inherit.aes = FALSE
    ) +
    scale_x_continuous(limits = c(-180, 180), breaks = seq(-180, 180, 90)) +
    scale_colour_manual(name = "Model", values = ppc_level_cols) +
    scale_linetype_manual(name = "Component", values = ppc_component_lty) +
    labs(
      title = panel_title,
      x = "Signed error (°)", y = "Density"
    ) +
    theme_desc +
    theme(
      plot.title = element_text(size = 10, face = "bold"),
      legend.position = "none"
    )
}

ppc_specs <- expand_grid(
  raw = vm_raw_ids,
  cue = c("NR", "R"),
  duration = c("50ms", "350ms")
) %>%
  mutate(
    id_label = unname(vm_id_pub[raw]),
    row_key = interaction(cue, duration, lex.order = TRUE)
  ) %>%
  arrange(row_key, id_label)

ppc_plots <- pmap(
  ppc_specs,
  function(raw, cue, duration, id_label, row_key) {
    vm_ppc_plot_cell(raw, id_label, cue, duration)
  }
)

ppc_plots <- ppc_plots[!vapply(ppc_plots, is.null, logical(1))]
if (length(ppc_plots) > 0) {
  for (i in seq_along(ppc_plots)) {
    if (i > 1) {
      ppc_plots[[i]] <- ppc_plots[[i]] + theme(legend.position = "none")
    } else {
      ppc_plots[[i]] <- ppc_plots[[i]] +
        theme(
          legend.position = "bottom",
          legend.box = "vertical",
          legend.text = element_text(size = 8)
        )
    }
  }

  patchwork::wrap_plots(ppc_plots, ncol = 4, byrow = TRUE) +
    patchwork::plot_annotation(
      title = "Posterior predictive checks: all sessions pooled [2κ]",
      subtitle = paste(
        "Rows: NR/R × 50/350 ms · Columns: P01–P04 ·",
        "Solid = mixture PDF · Dashed = components (colour = model level)"
      ),
      theme = theme(
        plot.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 10)
      )
    )
}
```

# 4. Bayesian primary analysis: cirSD (discrete session)

Parallel to AbsErr, but the outcome is **circular SD** of signed errors.
cirSD is a cell summary (one value per Cue × Duration × Session). A saturated
`CueType * DurationF * SessionF` mean model on a single cirSD per cell would
have no residual degrees of freedom. To keep the **same design formula** as
AbsErr and still identify residual variance, we use **B bootstrap cirSD
replicates per cell** (resample trials within the cell, recompute cirSD).

Model (same fixed structure as AbsErr discrete):

`cirSD_deg ~ CueType * DurationF * SessionF`

Mixture modelling is **not** repeated here: the mixture already targets the
signed-error distribution directly (finer than AbsErr or cirSD summaries).

## 4.1 Build bootstrap cirSD dataset and fit / load models

```{r cirsd-bayes-data, message=FALSE, warning=FALSE}
rope_cir <- rope_abs
B_cir <- 40
set.seed(1234)

circ_sd_deg <- function(x) {
  x <- x[is.finite(x)]
  if (length(x) < 5) return(NA_real_)
  as.numeric(sd.circular(circular(x, units = "degrees"))) * (180 / pi)
}

build_cirsd_boot <- function(d_sub, B = B_cir) {
  rows <- list()
  i <- 0
  for (cue in levels(d_sub$CueType)) {
    for (dur in levels(d_sub$DurationF)) {
      for (sess_i in levels(d_sub$SessionF)) {
        idx <- which(
          d_sub$CueType == cue &
            d_sub$DurationF == dur &
            d_sub$SessionF == sess_i
        )
        x <- d_sub$SignedErr[idx]
        x <- x[is.finite(x)]
        stopifnot(length(x) >= 10)
        for (b in seq_len(B)) {
          i <- i + 1
          rows[[i]] <- tibble(
            CueType = cue,
            DurationF = dur,
            SessionF = sess_i,
            boot_id = b,
            cirSD_deg = circ_sd_deg(sample(x, replace = TRUE))
          )
        }
      }
    }
  }
  bind_rows(rows) %>%
    mutate(
      CueType = factor(CueType, levels = levels(d_sub$CueType)),
      DurationF = factor(DurationF, levels = levels(d_sub$DurationF)),
      SessionF = factor(SessionF, levels = levels(d_sub$SessionF))
    )
}

dir.create("bayes_primary_discrete", showWarnings = FALSE, recursive = TRUE)

cir_boot_files <- c(
  AQ = "bayes_primary_discrete/cirsd_boot_AQ.rds",
  HC = "bayes_primary_discrete/cirsd_boot_HC.rds",
  YILIU = "bayes_primary_discrete/cirsd_boot_YILIU.rds",
  YDL = "bayes_primary_discrete/cirsd_boot_YDL.rds"
)
d_cir_AQ <- if (file.exists(cir_boot_files[["AQ"]])) readRDS(cir_boot_files[["AQ"]]) else {
  tmp <- build_cirsd_boot(d_AQ); saveRDS(tmp, cir_boot_files[["AQ"]]); tmp
}
d_cir_HC <- if (file.exists(cir_boot_files[["HC"]])) readRDS(cir_boot_files[["HC"]]) else {
  tmp <- build_cirsd_boot(d_HC); saveRDS(tmp, cir_boot_files[["HC"]]); tmp
}
d_cir_YILIU <- if (file.exists(cir_boot_files[["YILIU"]])) readRDS(cir_boot_files[["YILIU"]]) else {
  tmp <- build_cirsd_boot(d_YILIU); saveRDS(tmp, cir_boot_files[["YILIU"]]); tmp
}
d_cir_YDL <- if (file.exists(cir_boot_files[["YDL"]])) readRDS(cir_boot_files[["YDL"]]) else {
  tmp <- build_cirsd_boot(d_YDL); saveRDS(tmp, cir_boot_files[["YDL"]]); tmp
}

stopifnot(nrow(d_cir_AQ) == 2 * 7 * 10 * B_cir)
stopifnot(nrow(d_cir_YDL) == 2 * 7 * 10 * B_cir)

bayes_priors_cir <- c(
  prior(normal(40, 20), class = "Intercept"),
  prior(normal(0, 15), class = "b"),
  prior(exponential(1 / 15), class = "sigma"),
  prior(gamma(2, 0.05), class = "nu")
)

m_cir_AQ <- brm(
  cirSD_deg ~ CueType * DurationF * SessionF,
  data = d_cir_AQ, family = student(), prior = bayes_priors_cir,
  chains = 4, cores = 4, iter = 4000, warmup = 2000,
  control = list(adapt_delta = 0.95, max_treedepth = 12), seed = 1234,
  file = "bayes_primary_discrete/m_AQ_cirsd", file_refit = "on_change"
)
m_cir_HC <- brm(
  cirSD_deg ~ CueType * DurationF * SessionF,
  data = d_cir_HC, family = student(), prior = bayes_priors_cir,
  chains = 4, cores = 4, iter = 4000, warmup = 2000,
  control = list(adapt_delta = 0.95, max_treedepth = 12), seed = 1234,
  file = "bayes_primary_discrete/m_HC_cirsd", file_refit = "on_change"
)
m_cir_YILIU <- brm(
  cirSD_deg ~ CueType * DurationF * SessionF,
  data = d_cir_YILIU, family = student(), prior = bayes_priors_cir,
  chains = 4, cores = 4, iter = 4000, warmup = 2000,
  control = list(adapt_delta = 0.95, max_treedepth = 12), seed = 1234,
  file = "bayes_primary_discrete/m_YILIU_cirsd", file_refit = "on_change"
)
m_cir_YDL <- brm(
  cirSD_deg ~ CueType * DurationF * SessionF,
  data = d_cir_YDL, family = student(), prior = bayes_priors_cir,
  chains = 4, cores = 4, iter = 4000, warmup = 2000,
  control = list(adapt_delta = 0.95, max_treedepth = 12), seed = 1234,
  file = "bayes_primary_discrete/m_YDL_cirsd", file_refit = "on_change"
)

summary(m_cir_AQ)
summary(m_cir_HC)
summary(m_cir_YILIU)
summary(m_cir_YDL)
```

## 4.2 Theory contrasts (cirSD; same questions as AbsErr)

```{r cirsd-bayes-contrasts, message=FALSE, warning=FALSE}
contrast_rows_cir <- list()
row_i <- 0
models_cir <- list(P01 = m_cir_AQ, P02 = m_cir_HC, P03 = m_cir_YILIU, P04 = m_cir_YDL)

for (pid in names(models_cir)) {
  model <- models_cir[[pid]]

  emm_gain <- emmeans(model, ~ CueType | DurationF)
  draws_gain <- gather_emmeans_draws(
    contrast(emm_gain, method = list("NR - R" = c(1, -1)), by = "DurationF")
  )

  for (dur in levels(d$DurationF)) {
    v <- draws_gain$.value[draws_gain$DurationF == dur]
    v <- v[is.finite(v)]
    stopifnot(length(v) > 0)
    row_i <- row_i + 1
    contrast_rows_cir[[row_i]] <- tibble(
      ID = pid, effect = paste0("gain_", dur),
      median = median(v), lower = as.numeric(quantile(v, 0.025)),
      upper = as.numeric(quantile(v, 0.975)),
      p_lt_0 = mean(v < 0), p_gt_0 = mean(v > 0),
      p_in_rope = mean(abs(v) < rope_cir)
    )
  }
  v <- draws_gain$.value[draws_gain$DurationF == "50ms"]
  v <- v[is.finite(v)]
  row_i <- row_i + 1
  contrast_rows_cir[[row_i]] <- tibble(
    ID = pid, effect = "gain_at_50ms",
    median = median(v), lower = as.numeric(quantile(v, 0.025)),
    upper = as.numeric(quantile(v, 0.975)),
    p_lt_0 = mean(v < 0), p_gt_0 = mean(v > 0),
    p_in_rope = mean(abs(v) < rope_cir)
  )

  draws_dur <- gather_emmeans_draws(contrast(
    emmeans(model, ~ DurationF | CueType),
    method = list("linear_trend" = w_linear, "350ms - 50ms" = w_350_minus_50),
    by = "CueType"
  ))
  for (cue in c("NR", "R")) {
    for (cname in c("linear_trend", "350ms - 50ms")) {
      v <- draws_dur$.value[draws_dur$CueType == cue & draws_dur$contrast == cname]
      v <- v[is.finite(v)]
      stopifnot(length(v) > 0)
      row_i <- row_i + 1
      contrast_rows_cir[[row_i]] <- tibble(
        ID = pid, effect = paste0(cue, ": ", cname),
        median = median(v), lower = as.numeric(quantile(v, 0.025)),
        upper = as.numeric(quantile(v, 0.975)),
        p_lt_0 = mean(v < 0), p_gt_0 = mean(v > 0),
        p_in_rope = mean(abs(v) < rope_cir)
      )
    }
  }

  draws_ds <- gather_emmeans_draws(contrast(
    emmeans(model, ~ DurationF | CueType * SessionF),
    method = list("350ms - 50ms" = w_350_minus_50),
    by = c("CueType", "SessionF")
  )) %>%
    ungroup() %>%
    mutate(SessionF = as.character(SessionF)) %>%
    select(.draw, CueType, SessionF, .value)

  slice_early <- draws_ds %>%
    filter(SessionF %in% early_sessions) %>%
    group_by(.draw, CueType) %>% summarise(.value = mean(.value), .groups = "drop")
  slice_average <- draws_ds %>%
    filter(SessionF %in% sess_levels) %>%
    group_by(.draw, CueType) %>% summarise(.value = mean(.value), .groups = "drop")
  slice_late <- draws_ds %>%
    filter(SessionF %in% late_sessions) %>%
    group_by(.draw, CueType) %>% summarise(.value = mean(.value), .groups = "drop")
  slice_draws <- list(early = slice_early, average = slice_average, late = slice_late)

  for (slice in c("early", "average", "late")) {
    wide <- slice_draws[[slice]] %>% pivot_wider(names_from = CueType, values_from = .value)
    v <- wide$R - wide$NR
    v <- v[is.finite(v)]
    stopifnot(length(v) > 0)
    row_i <- row_i + 1
    contrast_rows_cir[[row_i]] <- tibble(
      ID = pid, effect = paste0("catchup_", slice),
      median = median(v), lower = as.numeric(quantile(v, 0.025)),
      upper = as.numeric(quantile(v, 0.975)),
      p_lt_0 = mean(v < 0), p_gt_0 = mean(v > 0),
      p_in_rope = mean(abs(v) < rope_cir)
    )
  }

  wide_e <- slice_draws$early %>% pivot_wider(names_from = CueType, values_from = .value) %>% mutate(catchup = R - NR)
  wide_l <- slice_draws$late %>% pivot_wider(names_from = CueType, values_from = .value) %>% mutate(catchup = R - NR)
  v <- wide_l$catchup[match(wide_e$.draw, wide_l$.draw)] - wide_e$catchup
  v <- v[is.finite(v)]
  stopifnot(length(v) > 0)
  row_i <- row_i + 1
  contrast_rows_cir[[row_i]] <- tibble(
    ID = pid, effect = "catchup_late_minus_early",
    median = median(v), lower = as.numeric(quantile(v, 0.025)),
    upper = as.numeric(quantile(v, 0.975)),
    p_lt_0 = mean(v < 0), p_gt_0 = mean(v > 0),
    p_in_rope = mean(abs(v) < rope_cir)
  )

  for (cue in c("NR", "R")) {
    early_df <- slice_draws$early %>% filter(CueType == cue) %>% arrange(.draw)
    late_df <- slice_draws$late %>% filter(CueType == cue) %>% arrange(.draw)
    stopifnot(identical(early_df$.draw, late_df$.draw))
    v <- late_df$.value - early_df$.value
    v <- v[is.finite(v)]
    stopifnot(length(v) > 0)
    row_i <- row_i + 1
    contrast_rows_cir[[row_i]] <- tibble(
      ID = pid, effect = paste0(cue, ": duration_effect_late_minus_early"),
      median = median(v), lower = as.numeric(quantile(v, 0.025)),
      upper = as.numeric(quantile(v, 0.975)),
      p_lt_0 = mean(v < 0), p_gt_0 = mean(v > 0),
      p_in_rope = mean(abs(v) < rope_cir)
    )
  }
}

primary_summary_cir <- bind_rows(contrast_rows_cir) %>% arrange(effect, ID)
primary_summary_cir

theory_focus_cir <- primary_summary_cir %>%
  filter(effect %in% c(
    "gain_at_50ms",
    "NR: linear_trend", "R: linear_trend",
    "NR: 350ms - 50ms", "R: 350ms - 50ms",
    "catchup_early", "catchup_average", "catchup_late",
    "catchup_late_minus_early",
    "NR: duration_effect_late_minus_early",
    "R: duration_effect_late_minus_early"
  )) %>%
  mutate(question = case_when(
    effect == "gain_at_50ms" ~ "1. redundancy advantage",
    grepl("linear_trend|350ms - 50ms", effect) ~ "2. duration change by cue",
    grepl("^catchup_", effect) ~ "3. NR catch-up vs R",
    grepl("duration_effect_late_minus_early", effect) ~ "4. practice change of duration slope",
    TRUE ~ "other"
  )) %>%
  arrange(question, effect, ID)

theory_focus_cir

write_csv(theory_focus_cir, "bayes_primary_discrete/theory_focus_cirsd.csv")
write_csv(primary_summary_cir, "bayes_primary_discrete/primary_summary_cirsd.csv")
```

## 4.3 Posterior predictive checks (cirSD)

```{r cirsd-bayes-ppc, fig.width=10, fig.height=4}
(pp_check(m_cir_AQ, ndraws = 50, type = "ecdf_overlay") + labs(title = "AQ: cirSD ECDF", x = "cirSD (deg)")) |
  (pp_check(m_cir_HC, ndraws = 50, type = "ecdf_overlay") + labs(title = "HC: cirSD ECDF", x = "cirSD (deg)")) |
  (pp_check(m_cir_YILIU, ndraws = 50, type = "ecdf_overlay") + labs(title = "YILIU: cirSD ECDF", x = "cirSD (deg)")) |
  (pp_check(m_cir_YDL, ndraws = 50, type = "ecdf_overlay") + labs(title = "YDL: cirSD ECDF", x = "cirSD (deg)"))
```

## 4.4 Contrast plots (cirSD)

```{r cirsd-bayes-plots, fig.width=10, fig.height=8}
plot_duration_cir <- primary_summary_cir %>%
  filter(effect %in% c("NR: 350ms - 50ms", "R: 350ms - 50ms")) %>%
  mutate(CueType = if_else(grepl("^NR", effect), "NR", "R"))
plot_catch_cir <- primary_summary_cir %>%
  filter(effect %in% c("catchup_early", "catchup_average", "catchup_late")) %>%
  mutate(SessionSlice = factor(gsub("catchup_", "", effect), levels = c("early", "average", "late")))
plot_practice_cir <- primary_summary_cir %>%
  filter(grepl("duration_effect_late_minus_early", effect)) %>%
  mutate(CueType = if_else(grepl("^NR", effect), "NR", "R"))

p1c <- ggplot(plot_duration_cir, aes(CueType, median, ymin = lower, ymax = upper)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_hline(yintercept = c(-rope_cir, rope_cir), linetype = "dotted", colour = "grey40") +
  geom_pointrange() + facet_wrap(~ID) +
  labs(title = "50→350 ms cirSD change by cue (avg over sessions)",
       subtitle = "Negative = lower cirSD at 350 ms; dotted = ±ROPE",
       y = "Posterior median (deg)", x = "Cue")
p2c <- ggplot(plot_catch_cir, aes(SessionSlice, median, ymin = lower, ymax = upper, group = 1)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_hline(yintercept = c(-rope_cir, rope_cir), linetype = "dotted", colour = "grey40") +
  geom_line() + geom_pointrange() + facet_wrap(~ID) +
  labs(title = "NR catch-up in cirSD across session slices",
       subtitle = "Positive = NR cirSD fell more than R (50→350)",
       y = "Posterior median (deg)", x = "Session slice")
p3c <- ggplot(plot_practice_cir, aes(CueType, median, ymin = lower, ymax = upper)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_hline(yintercept = c(-rope_cir, rope_cir), linetype = "dotted", colour = "grey40") +
  geom_pointrange() + facet_wrap(~ID) +
  labs(title = "Practice change in 50→350 cirSD duration effect",
       subtitle = "Late (S9–10) − early (S1–2)",
       y = "Posterior median (deg)", x = "Cue")
p1c / p2c / p3c
```

# 5. VM ladder: Session 1 vs Session 10 (γ and κ from L1, 2κ)

Uses **L1_2k** posterior cell parameters. Requires Section 3.2 (L1) to be fit first.

```{r vm-s1-s10, message=FALSE, warning=FALSE}
vm_init_plot_defaults()
if (!exists("vm_L1")) {
  vm_L1 <- readRDS("bayes_primary_discrete/vm_ladder_L1_results.rds")$fits
}

vm_s1s10_rows <- list()
row_i <- 0

for (raw in vm_raw_ids) {
  fit_obj <- vm_L1[[raw]][["L1_2k"]]
  stopifnot(!is.null(fit_obj))
  cp <- extract_vm_cell_params(fit_obj)

  for (cue in c("NR", "R")) {
    for (dur in levels(d$DurationF)) {
      s1 <- cp %>% filter(CueType == cue, DurationF == dur, SessionF == "1")
      s10 <- cp %>% filter(CueType == cue, DurationF == dur, SessionF == "10")
      stopifnot(nrow(s1) == 1, nrow(s10) == 1)
      dur_ms <- as.numeric(gsub("ms", "", dur))
      for (param in c("gamma", "kappa")) {
        row_i <- row_i + 1
        vm_s1s10_rows[[row_i]] <- tibble(
          ID = vm_id_pub[[raw]],
          CueType = cue,
          DurationF = dur,
          Duration_ms = dur_ms,
          parameter = param,
          median = s10[[paste0(param, "_median")]] - s1[[paste0(param, "_median")]],
          lower = s10[[paste0(param, "_lower")]] - s1[[paste0(param, "_upper")]],
          upper = s10[[paste0(param, "_upper")]] - s1[[paste0(param, "_lower")]]
        )
      }
    }
  }
}

vm_s1s10 <- bind_rows(vm_s1s10_rows) %>% arrange(ID, parameter, CueType, Duration_ms)
vm_s1s10

write_csv(vm_s1s10, "bayes_primary_discrete/vm_ladder_s1_vs_s10.csv")
```

```{r vm-s1-s10-plot, fig.width=10, fig.height=8}
ggplot(
  vm_s1s10 %>% filter(parameter == "gamma"),
  aes(Duration_ms, median, ymin = lower, ymax = upper, colour = CueType, group = CueType)
) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_line(linewidth = 0.5) +
  geom_errorbar(width = 8, linewidth = 0.35) +
  geom_point(size = 2) +
  facet_wrap(~ID) +
  scale_x_continuous(breaks = seq(50, 350, by = 50)) +
  scale_colour_manual(values = cue_cols) +
  labs(
    title = "L1 (2κ): Session 10 − Session 1 change in γ",
    x = "Encoding duration (ms)",
    y = "Δγ (median + 95% CrI)"
  ) +
  theme_desc

ggplot(
  vm_s1s10 %>% filter(parameter == "kappa"),
  aes(Duration_ms, median, ymin = lower, ymax = upper, colour = CueType, group = CueType)
) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_line(linewidth = 0.5) +
  geom_errorbar(width = 8, linewidth = 0.35) +
  geom_point(size = 2) +
  facet_wrap(~ID) +
  scale_x_continuous(breaks = seq(50, 350, by = 50)) +
  scale_colour_manual(values = cue_cols) +
  labs(
    title = "L1 (2κ): Session 10 − Session 1 change in κ",
    x = "Encoding duration (ms)",
    y = expression(Delta * kappa ~ "(median + 95% CrI)")
  ) +
  theme_desc
```


# 6. Pre-mixture 2: cirSD Session 1 vs Session 10 (bootstrap)

Companion in circular-SD units (same procedure as the continuous-session
script). For each participant × cue × duration: compute cirSD in Session
1 and Session 10, then bootstrap S1 − S10. Positive = higher circular
variability in Session 1.

```{r cirsd-s1-s10, message=FALSE, warning=FALSE}
B <- 10000  # bootstrap replicates
set.seed(1234)

ids <- participants
cues <- c("NR", "R")
durs <- levels(d$DurationF)

cirsd_rows <- list()
row_i <- 0

for (pid in ids) {
  for (cue in cues) {
    for (dur in durs) {
      # Trials in this cell for Session 1 and Session 10.
      x1 <- d$SignedErr[d$ID == pid & d$CueType == cue & d$DurationF == dur & d$Session == 1]
      x10 <- d$SignedErr[d$ID == pid & d$CueType == cue & d$DurationF == dur & d$Session == 10]
      x1 <- x1[is.finite(x1)]
      x10 <- x10[is.finite(x10)]

      # Need enough trials for a stable circular SD.
      if (length(x1) < 5 || length(x10) < 5) {
        row_i <- row_i + 1
        cirsd_rows[[row_i]] <- tibble(
          ID = pid, CueType = cue, DurationF = dur,
          Duration_ms = as.numeric(gsub("ms", "", dur)),
          estimate = NA_real_, lower = NA_real_, upper = NA_real_,
          boot_se = NA_real_, n1 = length(x1), n10 = length(x10),
          excludes_0 = NA
        )
        next
      }

      # Observed cirSD difference (degrees).
      sd1 <- as.numeric(sd.circular(circular(x1, units = "degrees"))) * (180 / pi)
      sd10 <- as.numeric(sd.circular(circular(x10, units = "degrees"))) * (180 / pi)
      obs <- sd1 - sd10

      # Bootstrap: resample each session's trials with replacement.
      boot_vals <- numeric(B)
      for (b in seq_len(B)) {
        s1_b <- sample(x1, replace = TRUE)
        s10_b <- sample(x10, replace = TRUE)
        sd1_b <- as.numeric(sd.circular(circular(s1_b, units = "degrees"))) * (180 / pi)
        sd10_b <- as.numeric(sd.circular(circular(s10_b, units = "degrees"))) * (180 / pi)
        boot_vals[b] <- sd1_b - sd10_b
      }
      boot_vals <- boot_vals[is.finite(boot_vals)]

      lo <- as.numeric(quantile(boot_vals, 0.025))
      hi <- as.numeric(quantile(boot_vals, 0.975))

      row_i <- row_i + 1
      cirsd_rows[[row_i]] <- tibble(
        ID = pid, CueType = cue, DurationF = dur,
        Duration_ms = as.numeric(gsub("ms", "", dur)),
        estimate = obs,
        lower = lo,
        upper = hi,
        boot_se = sd(boot_vals),
        n1 = length(x1),
        n10 = length(x10),
        excludes_0 = (lo > 0 & hi > 0) | (lo < 0 & hi < 0)
      )
    }
  }
}

cirsd_s1_s10 <- bind_rows(cirsd_rows) %>%
  arrange(ID, CueType, Duration_ms)

cirsd_s1_s10
write_csv(cirsd_s1_s10, "bayes_primary_discrete/cirsd_s1_vs_s10_bootstrap.csv")
```

```{r cirsd-s1-s10-short, message=FALSE, warning=FALSE}
# Short-duration companion: mean of cirSD(50/100/150) in S1 minus same mean in S10.
# Bootstrap resamples trials within each duration × session cell, then re-averages.
cirsd_short_rows <- list()
row_i <- 0
short_durs <- c("50ms", "100ms", "150ms")

for (pid in ids) {
  for (cue in cues) {
    # Observed mean cirSD across the three short durations.
    sd1_vec <- numeric(3)
    sd10_vec <- numeric(3)
    ok <- TRUE
    for (j in seq_along(short_durs)) {
      dur <- short_durs[j]
      x1 <- d$SignedErr[d$ID == pid & d$CueType == cue & d$DurationF == dur & d$Session == 1]
      x10 <- d$SignedErr[d$ID == pid & d$CueType == cue & d$DurationF == dur & d$Session == 10]
      x1 <- x1[is.finite(x1)]
      x10 <- x10[is.finite(x10)]
      if (length(x1) < 5 || length(x10) < 5) {
        ok <- FALSE
        break
      }
      sd1_vec[j] <- as.numeric(sd.circular(circular(x1, units = "degrees"))) * (180 / pi)
      sd10_vec[j] <- as.numeric(sd.circular(circular(x10, units = "degrees"))) * (180 / pi)
    }

    if (!ok) {
      row_i <- row_i + 1
      cirsd_short_rows[[row_i]] <- tibble(
        ID = pid, CueType = cue,
        estimate = NA_real_, lower = NA_real_, upper = NA_real_,
        boot_se = NA_real_, excludes_0 = NA
      )
      next
    }

    obs <- mean(sd1_vec) - mean(sd10_vec)

    boot_vals <- numeric(B)
    for (b in seq_len(B)) {
      sd1_b <- numeric(3)
      sd10_b <- numeric(3)
      for (j in seq_along(short_durs)) {
        dur <- short_durs[j]
        x1 <- d$SignedErr[d$ID == pid & d$CueType == cue & d$DurationF == dur & d$Session == 1]
        x10 <- d$SignedErr[d$ID == pid & d$CueType == cue & d$DurationF == dur & d$Session == 10]
        x1 <- x1[is.finite(x1)]
        x10 <- x10[is.finite(x10)]
        s1_b <- sample(x1, replace = TRUE)
        s10_b <- sample(x10, replace = TRUE)
        sd1_b[j] <- as.numeric(sd.circular(circular(s1_b, units = "degrees"))) * (180 / pi)
        sd10_b[j] <- as.numeric(sd.circular(circular(s10_b, units = "degrees"))) * (180 / pi)
      }
      boot_vals[b] <- mean(sd1_b) - mean(sd10_b)
    }
    boot_vals <- boot_vals[is.finite(boot_vals)]
    lo <- as.numeric(quantile(boot_vals, 0.025))
    hi <- as.numeric(quantile(boot_vals, 0.975))

    row_i <- row_i + 1
    cirsd_short_rows[[row_i]] <- tibble(
      ID = pid, CueType = cue,
      estimate = obs, lower = lo, upper = hi,
      boot_se = sd(boot_vals),
      excludes_0 = (lo > 0 & hi > 0) | (lo < 0 & hi < 0)
    )
  }
}

cirsd_short <- bind_rows(cirsd_short_rows) %>% arrange(ID, CueType)
cirsd_short
write_csv(cirsd_short, "bayes_primary_discrete/cirsd_s1_vs_s10_short_bootstrap.csv")
```

```{r cirsd-s1-s10-plot, fig.width=10, fig.height=5}
ggplot(
  cirsd_s1_s10,
  aes(Duration_ms, estimate, ymin = lower, ymax = upper, colour = CueType, group = CueType)
) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_line() +
  geom_pointrange() +
  facet_wrap(~ID) +
  labs(
    title = "cirSD Session 1 − Session 10 by duration",
    subtitle = "Bootstrap 95% CI; positive = higher circular variability in Session 1",
    x = "Encoding duration (ms)",
    y = "cirSD difference (deg)"
  )
```

# 7. Circular mixture (3-comp): per-cell ML (target + guess + swap)

Bays-style mixture on signed errors, one independent ML fit per cue ×
duration cell (14 cells × participant; sessions pooled). Always 3 unique
non-target colours, so the swap term is three equal von Mises bumps:

`p(θ) = p_target · VM(0, κ) + (β/3) Σ_{j=1..3} VM(μ_j, κ) + γ · Uniform`

- Shared κ for target and colour-swap
- Stick-break order: estimate β (swap) first, then γ from the leftover
  (pure guesses expected rare); p_target = 1 − β − γ
- Fits from `Rscript s1_mixture_cells.R` →
  `mixture_cells/all_cell_params.csv`

```{r mixture-load}
mix_file <- "mixture_cells/all_cell_params.csv"
need_mix <- (!file.exists(mix_file)) ||
  (!all(id_raw %in% read_csv(mix_file, show_col_types = FALSE)$ID))
if (need_mix) {
  message("Running s1_mixture_cells.R ...")
  system("Rscript s1_mixture_cells.R", intern = FALSE)
}
stopifnot(file.exists(mix_file))

mix_cells <- read_csv(mix_file, show_col_types = FALSE) %>%
  mutate(
    ID = factor(id_pub[ID], levels = participants),
    CueType = factor(CueType, levels = c("NR", "R")),
    DurationF = factor(
      DurationF,
      levels = c("50ms", "100ms", "150ms", "200ms", "250ms", "300ms", "350ms")
    ),
    Duration_ms = as.numeric(gsub("ms", "", as.character(DurationF)))
  )

stopifnot(all(participants %in% mix_cells$ID))
stopifnot(all(mix_cells$convergence == 0 | is.na(mix_cells$convergence)))
stopifnot(all(mix_cells$p_target + mix_cells$gamma + mix_cells$beta > 0.99))
stopifnot(all(mix_cells$p_target + mix_cells$gamma + mix_cells$beta < 1.01))

mix_cells %>%
  select(ID, CueType, DurationF, n_trials, kappa, circSD_deg,
         p_target, gamma, beta, nll, convergence) %>%
  mutate(across(c(kappa, circSD_deg, p_target, gamma, beta, nll), ~ round(.x, 3))) %>%
  arrange(ID, CueType, DurationF)
```

```{r mixture-proportion-plot, fig.width=10, fig.height=6}
prop_long <- mix_cells %>%
  select(ID, CueType, Duration_ms, p_target, gamma, beta) %>%
  pivot_longer(
    c(p_target, gamma, beta),
    names_to = "component",
    values_to = "proportion"
  ) %>%
  mutate(
    component = factor(
      component,
      levels = c("p_target", "beta", "gamma"),
      labels = c("Target", "Colour swap", "Guess")
    )
  )

ggplot(prop_long, aes(Duration_ms, proportion, fill = component)) +
  geom_col(width = 35, colour = "grey20", linewidth = 0.2) +
  facet_grid(CueType ~ ID) +
  scale_x_continuous(breaks = seq(50, 350, by = 50)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 1)) +
  labs(
    title = "Mixture response proportions by cue and duration",
    subtitle = "Independent ML fits per cell; sessions pooled",
    x = "Encoding duration (ms)",
    y = "Response proportion",
    fill = "Component"
  ) +
  theme(legend.position = "bottom")
```

```{r mixture-kappa-plot, fig.width=10, fig.height=4}
ggplot(mix_cells, aes(Duration_ms, kappa, colour = CueType, group = CueType)) +
  geom_line() +
  geom_point(size = 2) +
  facet_wrap(~ID, scales = "free_y") +
  scale_x_continuous(breaks = seq(50, 350, by = 50)) +
  labs(
    title = "Shared von Mises precision (κ) by cue and duration",
    x = "Encoding duration (ms)",
    y = "κ"
  )
```

```{r mixture-weight-lines, fig.width=10, fig.height=6}
ggplot(prop_long, aes(Duration_ms, proportion, colour = CueType, group = CueType)) +
  geom_line() +
  geom_point() +
  facet_grid(component ~ ID, scales = "free_y") +
  scale_x_continuous(breaks = seq(50, 350, by = 100)) +
  labs(
    title = "Mixture weights by cue and duration",
    x = "Encoding duration (ms)",
    y = "Proportion"
  )
```

# 8. Circular mixture (2-comp): per-cell ML (target + guess)

Same cell grid as §7, but **no colour-swap**. Large errors can only be
explained as guesses. Useful because swap models often over-attribute
guessing as swaps.

`p(θ) = p_target · VM(0, κ) + γ · Uniform`, with `p_target = 1 − γ`.

Fits from `Rscript s1_mixture_cells.R` (Part D) →
`mixture_cells/all_cell_params_2comp.csv`.

```{r mixture2-load}
mix2_file <- "mixture_cells/all_cell_params_2comp.csv"
need_mix2 <- (!file.exists(mix2_file)) ||
  (!all(id_raw %in% read_csv(mix2_file, show_col_types = FALSE)$ID))
if (need_mix2) {
  message("Running s1_mixture_cells.R (includes 2-comp Part D) ...")
  system("Rscript s1_mixture_cells.R", intern = FALSE)
}
stopifnot(file.exists(mix2_file))

mix2_cells <- read_csv(mix2_file, show_col_types = FALSE) %>%
  mutate(
    ID = factor(id_pub[ID], levels = participants),
    CueType = factor(CueType, levels = c("NR", "R")),
    DurationF = factor(
      DurationF,
      levels = c("50ms", "100ms", "150ms", "200ms", "250ms", "300ms", "350ms")
    ),
    Duration_ms = as.numeric(gsub("ms", "", as.character(DurationF)))
  )

stopifnot(all(participants %in% mix2_cells$ID))
stopifnot(all(mix2_cells$convergence == 0 | is.na(mix2_cells$convergence)))
stopifnot(all(abs(mix2_cells$p_target + mix2_cells$gamma - 1) < 1e-6))

mix2_cells %>%
  select(ID, CueType, DurationF, n_trials, kappa, circSD_deg,
         p_target, gamma, nll, convergence) %>%
  mutate(across(c(kappa, circSD_deg, p_target, gamma, nll), ~ round(.x, 3))) %>%
  arrange(ID, CueType, DurationF)
```

```{r mixture2-proportion-plot, fig.width=10, fig.height=6}
prop2_long <- mix2_cells %>%
  select(ID, CueType, Duration_ms, p_target, gamma) %>%
  pivot_longer(
    c(p_target, gamma),
    names_to = "component",
    values_to = "proportion"
  ) %>%
  mutate(
    component = factor(
      component,
      levels = c("p_target", "gamma"),
      labels = c("Target", "Guess")
    )
  )

ggplot(prop2_long, aes(Duration_ms, proportion, fill = component)) +
  geom_col(width = 35, colour = "grey20", linewidth = 0.2) +
  facet_grid(CueType ~ ID) +
  scale_x_continuous(breaks = seq(50, 350, by = 50)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 1)) +
  labs(
    title = "2-comp mixture proportions by cue and duration",
    subtitle = "Target + guess only (no swap); independent ML fits; sessions pooled",
    x = "Encoding duration (ms)",
    y = "Response proportion",
    fill = "Component"
  ) +
  theme(legend.position = "bottom")
```

```{r mixture2-kappa-plot, fig.width=10, fig.height=4}
ggplot(mix2_cells, aes(Duration_ms, kappa, colour = CueType, group = CueType)) +
  geom_line() +
  geom_point(size = 2) +
  facet_wrap(~ID, scales = "free_y") +
  scale_x_continuous(breaks = seq(50, 350, by = 50)) +
  labs(
    title = "2-comp von Mises precision (κ) by cue and duration",
    x = "Encoding duration (ms)",
    y = "κ"
  )
```

```{r mixture2-weight-lines, fig.width=10, fig.height=5}
ggplot(prop2_long, aes(Duration_ms, proportion, colour = CueType, group = CueType)) +
  geom_line() +
  geom_point() +
  facet_grid(component ~ ID, scales = "free_y") +
  scale_x_continuous(breaks = seq(50, 350, by = 100)) +
  labs(
    title = "2-comp mixture weights by cue and duration",
    x = "Encoding duration (ms)",
    y = "Proportion"
  )
```

# 9. Simple von Mises: 14-κ vs 2-κ (sessions pooled)

Assumption-light variability check (no mixture). Signed errors are treated as
a single von Mises centred at 0 (target). Sessions are pooled within cue ×
duration.

- **14-κ model:** one κ per CueType × DurationF cell (14 parameters)
- **2-κ model:** one κ for all NR trials, one κ for all R trials (2 parameters)

Question: does response variability change enough across encoding durations that
the 14-κ model wins on BIC despite many more parameters? If the 2-κ model wins,
duration modulation of κ is not large enough (under BIC) to justify separate
duration κs — not proof that duration effects are “artefacts.”

`BIC = −2 log L + k log(n)` with `n` = trials for that participant.

```{r simple-vm-14-vs-2, message=FALSE, warning=FALSE}
dir.create("bayes_primary_discrete", showWarnings = FALSE, recursive = TRUE)

# log von Mises density at mu = 0 (vectorised over theta; scalar kappa).
log_vm0 <- function(theta, kappa) {
  log_i0 <- log(max(besselI(kappa, 0, expon.scaled = TRUE), .Machine$double.xmin)) + kappa
  kappa * cos(theta) - log(2 * pi) - log_i0
}

# ML κ for one set of signed-error radians (mean fixed at 0).
fit_kappa_vm <- function(theta) {
  theta <- theta[is.finite(theta)]
  stopifnot(length(theta) > 20)

  nll <- function(log_k) {
    k <- exp(log_k)
    if (!is.finite(k) || k < 1e-6 || k > 5000) return(1e12)
    -sum(log_vm0(theta, k))
  }

  starts <- log(c(1, 2, 4, 8, 16, 32))
  best_nll <- Inf
  best_par <- NA_real_
  best_conv <- NA_integer_
  for (s0 in starts) {
    fit <- tryCatch(
      optim(par = s0, fn = nll, method = "BFGS",
            control = list(maxit = 1000, reltol = 1e-12)),
      error = function(e) NULL
    )
    if (is.null(fit)) next
    if (fit$value < best_nll) {
      best_nll <- fit$value
      best_par <- fit$par
      best_conv <- fit$convergence
    }
  }
  stopifnot(is.finite(best_par), best_conv == 0)

  kappa <- exp(best_par)
  # Circular SD (deg) from κ.
  a <- besselI(kappa, 1, expon.scaled = TRUE) /
    max(besselI(kappa, 0, expon.scaled = TRUE), .Machine$double.xmin)
  a <- min(max(a, 1e-12), 1 - 1e-12)
  list(
    kappa = kappa,
    circSD_deg = sqrt(-2 * log(a)) * (180 / pi),
    logLik = -best_nll,
    nll = best_nll,
    n = length(theta),
    convergence = best_conv
  )
}

# Signed errors in radians for each participant.
d_vm <- d %>%
  mutate(
    theta = atan2(sin(SignedErr * pi / 180), cos(SignedErr * pi / 180))
  )

vm_kappa_rows <- list()
vm_bic_rows <- list()
row_k <- 0
row_b <- 0

for (pid in participants) {
  d_pid <- d_vm %>% filter(ID == pid)
  n_pid <- nrow(d_pid)

  # ---- 14-κ: fit each cue × duration cell ----
  logLik_14 <- 0
  for (cue in c("NR", "R")) {
    for (dur in levels(d$DurationF)) {
      th <- d_pid$theta[d_pid$CueType == cue & d_pid$DurationF == dur]
      fit <- fit_kappa_vm(th)
      logLik_14 <- logLik_14 + fit$logLik
      row_k <- row_k + 1
      vm_kappa_rows[[row_k]] <- tibble(
        ID = pid,
        model = "14k",
        CueType = cue,
        DurationF = dur,
        Duration_ms = as.numeric(gsub("ms", "", dur)),
        n_trials = fit$n,
        kappa = fit$kappa,
        circSD_deg = fit$circSD_deg,
        logLik_cell = fit$logLik,
        convergence = fit$convergence
      )
    }
  }
  k_14 <- 14L
  bic_14 <- -2 * logLik_14 + k_14 * log(n_pid)
  aic_14 <- -2 * logLik_14 + 2 * k_14

  # ---- 2-κ: one κ per cue (all durations pooled) ----
  logLik_2 <- 0
  for (cue in c("NR", "R")) {
    th <- d_pid$theta[d_pid$CueType == cue]
    fit <- fit_kappa_vm(th)
    logLik_2 <- logLik_2 + fit$logLik
    row_k <- row_k + 1
    vm_kappa_rows[[row_k]] <- tibble(
      ID = pid,
      model = "2k",
      CueType = cue,
      DurationF = NA_character_,
      Duration_ms = NA_real_,
      n_trials = fit$n,
      kappa = fit$kappa,
      circSD_deg = fit$circSD_deg,
      logLik_cell = fit$logLik,
      convergence = fit$convergence
    )
  }
  k_2 <- 2L
  bic_2 <- -2 * logLik_2 + k_2 * log(n_pid)
  aic_2 <- -2 * logLik_2 + 2 * k_2

  # Positive delta_BIC_14_minus_2 => 2-κ preferred by BIC
  # (14-κ has higher BIC). Negative => 14-κ wins.
  row_b <- row_b + 1
  vm_bic_rows[[row_b]] <- tibble(
    ID = pid,
    n_trials = n_pid,
    logLik_14k = logLik_14,
    logLik_2k = logLik_2,
    k_14k = k_14,
    k_2k = k_2,
    BIC_14k = bic_14,
    BIC_2k = bic_2,
    AIC_14k = aic_14,
    AIC_2k = aic_2,
    delta_BIC_14_minus_2 = bic_14 - bic_2,
    delta_AIC_14_minus_2 = aic_14 - aic_2,
    BIC_winner = if_else(bic_14 < bic_2, "14k", "2k"),
    AIC_winner = if_else(aic_14 < aic_2, "14k", "2k")
  )
}

vm_kappa_table <- bind_rows(vm_kappa_rows)
vm_bic_table <- bind_rows(vm_bic_rows)

vm_kappa_table %>%
  mutate(across(c(kappa, circSD_deg, logLik_cell), ~ round(.x, 3))) %>%
  arrange(model, ID, CueType, Duration_ms)

vm_bic_table %>%
  mutate(across(where(is.numeric), ~ round(.x, 2)))

write_csv(vm_kappa_table, "bayes_primary_discrete/simple_vm_kappa_table.csv")
write_csv(vm_bic_table, "bayes_primary_discrete/simple_vm_14k_vs_2k_bic.csv")
```

```{r simple-vm-14-vs-2-plot, fig.width=10, fig.height=4}
# 14-κ estimates by duration; horizontal lines = 2-κ cue-level κ.
k14 <- vm_kappa_table %>% filter(model == "14k")
k2 <- vm_kappa_table %>% filter(model == "2k") %>%
  select(ID, CueType, kappa_2k = kappa)

ggplot(k14, aes(Duration_ms, kappa, colour = CueType, group = CueType)) +
  geom_hline(
    data = k2,
    aes(yintercept = kappa_2k, colour = CueType),
    linetype = "dashed", alpha = 0.7
  ) +
  geom_line() +
  geom_point(size = 2) +
  facet_wrap(~ID, scales = "free_y") +
  scale_x_continuous(breaks = seq(50, 350, by = 50)) +
  labs(
    title = "Simple von Mises κ: 14-κ (points/lines) vs 2-κ (dashed)",
    subtitle = "Sessions pooled; mean fixed at 0. Lower BIC wins (see table).",
    x = "Encoding duration (ms)",
    y = "κ"
  )
```

# 10. Mixture 3-comp: κ free per cell vs κ tied within cue

After allowing swap (β) and guess (γ) to explain heavy tails, does **precision κ**
still need to vary with encoding duration?

- **κ-free:** existing per-cell mixture (14 × {κ, β, γ} = 42 parameters).
  logLik = sum of cell logLiks from `mixture_cells/`.
- **κ-tied:** one κ for NR and one κ for R across all durations; β and γ remain
  free in each of the 14 cells (2 + 14 + 14 = 30 parameters).

Sessions pooled. Fit κ-tied offline with `Rscript s1_mixture_kappa_compare.R`
(warm-started from the free cell fits).

`delta_BIC_free_minus_tied < 0` ⇒ κ-free wins (duration still matters for κ
after mixture decomposition).

```{r mixture-kappa-free-vs-tied, message=FALSE, warning=FALSE}
mix_bic_file <- "bayes_primary_discrete/mixture_kappa_free_vs_tied_bic.csv"
need_mix_kappa <- (!file.exists(mix_bic_file)) ||
  (!all(id_raw %in% read_csv(mix_bic_file, show_col_types = FALSE)$ID))
if (need_mix_kappa) {
  message("Running s1_mixture_kappa_compare.R (κ-tied fits) ...")
  system("Rscript s1_mixture_kappa_compare.R", intern = FALSE)
}
stopifnot(file.exists(mix_bic_file))

mix_kappa_bic <- read_csv(mix_bic_file, show_col_types = FALSE) %>%
  mutate(ID = factor(id_pub[ID], levels = participants))
mix_kappa_tied_params <- read_csv(
  "bayes_primary_discrete/mixture_kappa_tied_params.csv",
  show_col_types = FALSE
) %>%
  mutate(ID = factor(id_pub[ID], levels = participants))
mix_kappa_free_params <- read_csv(
  "bayes_primary_discrete/mixture_kappa_free_params.csv",
  show_col_types = FALSE
) %>%
  mutate(ID = factor(id_pub[ID], levels = participants))

mix_kappa_bic %>%
  mutate(across(where(is.numeric), ~ round(.x, 2)))

# Side-by-side κ: free (by duration) vs tied (horizontal)
k_free_plot <- mix_kappa_free_params %>%
  mutate(
    Duration_ms = as.numeric(gsub("ms", "", as.character(DurationF))),
    CueType = factor(CueType, levels = c("NR", "R"))
  )
k_tied_plot <- mix_kappa_bic %>%
  select(ID, kappa_tied_NR, kappa_tied_R) %>%
  pivot_longer(
    c(kappa_tied_NR, kappa_tied_R),
    names_to = "CueType",
    values_to = "kappa_tied"
  ) %>%
  mutate(CueType = if_else(CueType == "kappa_tied_NR", "NR", "R"))
```

```{r mixture-kappa-free-vs-tied-plot, fig.width=10, fig.height=4}
ggplot(k_free_plot, aes(Duration_ms, kappa, colour = CueType, group = CueType)) +
  geom_hline(
    data = k_tied_plot,
    aes(yintercept = kappa_tied, colour = CueType),
    linetype = "dashed", alpha = 0.7
  ) +
  geom_line() +
  geom_point(size = 2) +
  facet_wrap(~ID, scales = "free_y") +
  scale_x_continuous(breaks = seq(50, 350, by = 50)) +
  labs(
    title = "3-comp mixture κ: free per cell (lines) vs tied within cue (dashed)",
    subtitle = "β and γ free per cell in both models; sessions pooled",
    x = "Encoding duration (ms)",
    y = "κ"
  )
```

# 11. Mixture 2-comp: κ free per cell vs κ tied within cue

Same question as §10, but after **target + guess only** (no swap). Does κ still
need to vary with duration once γ can absorb heavy tails?

- **κ-free:** 14 × {κ, γ} = 28 parameters (`mixture_cells/all_cell_params_2comp.csv`).
- **κ-tied:** one κ per cue; γ free per cell (2 + 14 = 16 parameters).

Fit offline with `Rscript s1_mixture_kappa_compare.R` (2-comp section).

```{r mixture2-kappa-free-vs-tied, message=FALSE, warning=FALSE}
mix2_bic_file <- "bayes_primary_discrete/mixture_kappa_2comp_free_vs_tied_bic.csv"
need_mix2_kappa <- (!file.exists(mix2_bic_file)) ||
  (!all(id_raw %in% read_csv(mix2_bic_file, show_col_types = FALSE)$ID))
if (need_mix2_kappa) {
  message("Running s1_mixture_kappa_compare.R ...")
  system("Rscript s1_mixture_kappa_compare.R", intern = FALSE)
}
stopifnot(file.exists(mix2_bic_file))

mix2_kappa_bic <- read_csv(mix2_bic_file, show_col_types = FALSE) %>%
  mutate(ID = factor(id_pub[ID], levels = participants))
mix2_kappa_tied_params <- read_csv(
  "bayes_primary_discrete/mixture_kappa_2comp_tied_params.csv",
  show_col_types = FALSE
) %>%
  mutate(ID = factor(id_pub[ID], levels = participants))
mix2_kappa_free_params <- read_csv(
  "bayes_primary_discrete/mixture_kappa_2comp_free_params.csv",
  show_col_types = FALSE
) %>%
  mutate(ID = factor(id_pub[ID], levels = participants))

mix2_kappa_bic %>%
  mutate(across(where(is.numeric), ~ round(.x, 2)))

k2_free_plot <- mix2_kappa_free_params %>%
  mutate(
    Duration_ms = as.numeric(gsub("ms", "", as.character(DurationF))),
    CueType = factor(CueType, levels = c("NR", "R"))
  )
k2_tied_plot <- mix2_kappa_bic %>%
  select(ID, kappa_tied_NR, kappa_tied_R) %>%
  pivot_longer(
    c(kappa_tied_NR, kappa_tied_R),
    names_to = "CueType",
    values_to = "kappa_tied"
  ) %>%
  mutate(CueType = if_else(CueType == "kappa_tied_NR", "NR", "R"))
```

```{r mixture2-kappa-free-vs-tied-plot, fig.width=10, fig.height=4}
ggplot(k2_free_plot, aes(Duration_ms, kappa, colour = CueType, group = CueType)) +
  geom_hline(
    data = k2_tied_plot,
    aes(yintercept = kappa_tied, colour = CueType),
    linetype = "dashed", alpha = 0.7
  ) +
  geom_line() +
  geom_point(size = 2) +
  facet_wrap(~ID, scales = "free_y") +
  scale_x_continuous(breaks = seq(50, 350, by = 50)) +
  labs(
    title = "2-comp mixture κ: free per cell (lines) vs tied within cue (dashed)",
    subtitle = "γ free per cell in both models; sessions pooled; no swap",
    x = "Encoding duration (ms)",
    y = "κ"
  )
```

## Appendix

Legacy frequentist / exploratory Bayes material lives in
`s1_smallN_analysis.Rmd`. Primary discrete-session inference is Sections
3–11 above.
