---
title: "Study 1 AbsErr Bayes — discrete SessionF"
author: "Yu Liu"
date: "11th May 2026"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
editor_options: 
  markdown: 
    wrap: 72
---

# Overview

Companion to `s1_smallN_analysis.Rmd`, identical pipeline except
**session is categorical** (`SessionF`) instead of continuous
`Session_z`.

Model difference (Sections 3–4):

- Continuous: `AbsErr / cirSD ~ CueType * DurationF * Session_z + (1 | SessionF)`
- This script: `AbsErr / cirSD ~ CueType * DurationF * SessionF`
  (no random session intercept — `SessionF` is already a fixed factor)

Analysis order:

1.  Load and prepare trial data
2.  Descriptive plots (placeholder)
3.  Bayesian AbsErr primary (discrete session)
4.  Bayesian cirSD primary (same design; bootstrap cell replicates)
5.  Pre-mixture: AbsErr Session 1 vs 10 by duration
6.  Pre-mixture: cirSD Session 1 vs 10 (nonparametric bootstrap companion)
7.  Circular mixture cell fits (signed-error distribution; not repeated for cirSD)
8.  Simple von Mises κ comparison: 14-κ (cue × duration) vs 2-κ (cue only)
9.  Mixture κ free per cell vs κ tied within cue (β, γ still free per cell)

Style: linear commented code; short loops over AQ / HC / YILIU only.

# Setup

```{r setup, message=FALSE, warning=FALSE}
# Core packages for this pipeline.
library(tidyverse)   # data wrangling + ggplot2
library(nlme)        # frequentist mixed models (Appendix)
library(brms)        # Bayesian regression
library(bayesplot)   # posterior predictive checks
library(tidybayes)   # gather posterior draws from emmeans
library(emmeans)     # estimated marginal means / contrasts
library(patchwork)   # combine ggplot panels
library(loo)         # Appendix model comparison only
library(circular)    # circular SD for signed colour errors

theme_set(theme_bw())
options(mc.cores = parallel::detectCores())
set.seed(1234)  # reproducible bootstrap / Stan seeds where used
```

# 1. Load and prepare data

```{r load-data}
# Model-ready trial table from MATLAB preprocessing.
d <- read_csv("S1_smallN_model_ready.csv", show_col_types = FALSE)

# Quick sanity checks: expected columns and no empty file.
stopifnot(nrow(d) > 0)
stopifnot(all(c("ID", "CueType", "Duration", "Session", "Session_z",
                "AbsErr", "SignedErr") %in% names(d)))

glimpse(d)
```

```{r prepare-factors}
# Set factor levels explicitly so contrasts and plots keep a fixed order.
d <- d %>%
  mutate(
    ID = factor(ID),
    CueType = factor(CueType, levels = c("NR", "R")),
    DurationF = factor(
      Duration,
      levels = c("50ms", "100ms", "150ms", "200ms", "250ms", "300ms", "350ms")
    ),
    Grouping = factor(Grouping),
    SessionF = factor(Session),
    SessionPhase = factor(SessionPhase, levels = c("Early", "Middle", "Late")),
    logRT = log(RT)
  )

# One data frame per currently complete participant.
# droplevels() drops unused ID levels so factor levels stay clean.
d_AQ <- droplevels(filter(d, ID == "AQ"))
d_HC <- droplevels(filter(d, ID == "HC"))
d_YILIU <- droplevels(filter(d, ID == "YILIU"))

# Confirm design levels.
levels(d$CueType)
levels(d$DurationF)
levels(d$SessionF)
```

# 2. Descriptive figures (placeholder)

These plots are for exploration only. Manuscript figures will be updated
later.

```{r descriptives-cirsd}
# Circular SD of signed errors by participant × cue × duration.
# sd.circular returns radians when units="degrees"; convert to degrees.
circ_summary <- d %>%
  group_by(ID, CueType, DurationF) %>%
  summarise(
    circSD = as.numeric(sd.circular(circular(SignedErr, units = "degrees"))) * (180 / pi),
    meanAbs = mean(AbsErr),
    .groups = "drop"
  )

ggplot(circ_summary, aes(DurationF, circSD, colour = CueType, group = CueType)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  facet_wrap(~ID) +
  labs(y = "Circular SD (deg)", x = "Encoding duration")
# Same summary pooled across participants (descriptive only; not primary inference).
group_circ <- d %>%
  group_by(CueType, DurationF) %>%
  summarise(
    circSD = as.numeric(sd.circular(circular(SignedErr, units = "degrees"))) * (180 / pi),
    .groups = "drop"
  )

ggplot(group_circ, aes(DurationF, circSD, colour = CueType, group = CueType)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  labs(y = "Circular SD (deg)", x = "Encoding duration")
```

# 3. Bayesian primary analysis: AbsErr (discrete session)

One Student-t model per participant:

`AbsErr ~ CueType * DurationF * SessionF`

- Cue, duration, and session are all categorical (cell means).
- No `(1 | SessionF)`: session is already a fixed factor, so a random
  session intercept would be redundant / unidentified.
- “Average session” contrasts = equal-weight average over all `SessionF`
  levels (emmeans default when SessionF is omitted from the specs).
- Early = sessions 1–2; late = sessions 9–10 (mean of those session
  contrasts).
- ROPE for AbsErr differences: ±3 degrees.

## 3.1 Fit / load models

```{r bayes-fit, message=FALSE, warning=FALSE}
# Region of practical equivalence for AbsErr contrasts (degrees).
rope_abs <- 3

# Session factor levels used in early / late / S1 / S10 contrasts.
sess_levels <- levels(d$SessionF)
stopifnot(all(c("1", "2", "9", "10") %in% sess_levels))
early_sessions <- c("1", "2")
late_sessions  <- c("9", "10")

# Ordered-duration linear contrast weights (7 levels: 50 ... 350 ms).
# Negative AbsErr contrast = error falls as duration increases.
w_linear <- c(-3, -2, -1, 0, 1, 2, 3)
w_350_minus_50 <- c(-1, 0, 0, 0, 0, 0, 1)

# Weights for Session 1 − Session 10 on the SessionF factor (order = sess_levels).
w_s1_minus_s10 <- setNames(rep(0, length(sess_levels)), sess_levels)
w_s1_minus_s10["1"]  <- 1
w_s1_minus_s10["10"] <- -1

# Mildly regularising priors. No class = "sd" (no random effects in this model).
bayes_priors <- c(
  prior(normal(30, 20), class = "Intercept"),
  prior(normal(0, 15), class = "b"),
  prior(exponential(1), class = "sigma"),
  prior(gamma(2, 0.1), class = "nu")
)

dir.create("bayes_primary_discrete", showWarnings = FALSE, recursive = TRUE)

# Fit / cache each participant. Saved under bayes_primary_discrete/ so they do
# not overwrite the continuous Session_z models.
m_AQ <- brm(
  AbsErr ~ CueType * DurationF * SessionF,
  data = d_AQ,
  family = student(),
  prior = bayes_priors,
  chains = 4, cores = 4, iter = 4000, warmup = 2000,
  control = list(adapt_delta = 0.95, max_treedepth = 12),
  seed = 1234,
  file = "bayes_primary_discrete/m_AQ_abserr",
  file_refit = "on_change"
)

m_HC <- brm(
  AbsErr ~ CueType * DurationF * SessionF,
  data = d_HC,
  family = student(),
  prior = bayes_priors,
  chains = 4, cores = 4, iter = 4000, warmup = 2000,
  control = list(adapt_delta = 0.95, max_treedepth = 12),
  seed = 1234,
  file = "bayes_primary_discrete/m_HC_abserr",
  file_refit = "on_change"
)

m_YILIU <- brm(
  AbsErr ~ CueType * DurationF * SessionF,
  data = d_YILIU,
  family = student(),
  prior = bayes_priors,
  chains = 4, cores = 4, iter = 4000, warmup = 2000,
  control = list(adapt_delta = 0.95, max_treedepth = 12),
  seed = 1234,
  file = "bayes_primary_discrete/m_YILIU_abserr",
  file_refit = "on_change"
)

summary(m_AQ)
summary(m_HC)
summary(m_YILIU)
```

## 3.2 Theory contrasts (written out)

Same scientific questions as the continuous script; session slices use
discrete session labels instead of Session_z landmarks.

1.  Redundancy gain at 50 ms (averaged over sessions): AbsErr(NR) −
    AbsErr(R)
2.  Duration change by cue (averaged over sessions): 350−50 and linear
    trend
3.  Catch-up early / average / late; average = all sessions; early =
    1–2; late = 9–10
4.  Practice: late − early change in the 50→350 duration effect, by cue

```{r bayes-contrasts, message=FALSE, warning=FALSE}
contrast_rows <- list()
row_i <- 0
models <- list(AQ = m_AQ, HC = m_HC, YILIU = m_YILIU)

for (pid in names(models)) {
  model <- models[[pid]]

  # ---- 1) Redundancy gain by duration, averaged over SessionF ----
  emm_gain <- emmeans(model, ~ CueType | DurationF)
  ct_gain <- contrast(emm_gain, method = list("NR - R" = c(1, -1)), by = "DurationF")
  draws_gain <- gather_emmeans_draws(ct_gain)

  for (dur in levels(d$DurationF)) {
    v <- draws_gain$.value[draws_gain$DurationF == dur]
    v <- v[is.finite(v)]
    stopifnot(length(v) > 0)
    row_i <- row_i + 1
    contrast_rows[[row_i]] <- tibble(
      ID = pid,
      effect = paste0("gain_", dur),
      median = median(v),
      lower = as.numeric(quantile(v, 0.025)),
      upper = as.numeric(quantile(v, 0.975)),
      p_lt_0 = mean(v < 0),
      p_gt_0 = mean(v > 0),
      p_in_rope = mean(abs(v) < rope_abs)
    )
  }

  v <- draws_gain$.value[draws_gain$DurationF == "50ms"]
  v <- v[is.finite(v)]
  row_i <- row_i + 1
  contrast_rows[[row_i]] <- tibble(
    ID = pid,
    effect = "gain_at_50ms",
    median = median(v),
    lower = as.numeric(quantile(v, 0.025)),
    upper = as.numeric(quantile(v, 0.975)),
    p_lt_0 = mean(v < 0),
    p_gt_0 = mean(v > 0),
    p_in_rope = mean(abs(v) < rope_abs)
  )

  # ---- 2) Duration effects by cue, averaged over SessionF ----
  emm_dur <- emmeans(model, ~ DurationF | CueType)
  ct_dur <- contrast(
    emm_dur,
    method = list(
      "linear_trend" = w_linear,
      "350ms - 50ms" = w_350_minus_50
    ),
    by = "CueType"
  )
  draws_dur <- gather_emmeans_draws(ct_dur)

  for (cue in c("NR", "R")) {
    for (cname in c("linear_trend", "350ms - 50ms")) {
      v <- draws_dur$.value[draws_dur$CueType == cue & draws_dur$contrast == cname]
      v <- v[is.finite(v)]
      stopifnot(length(v) > 0)
      row_i <- row_i + 1
      contrast_rows[[row_i]] <- tibble(
        ID = pid,
        effect = paste0(cue, ": ", cname),
        median = median(v),
        lower = as.numeric(quantile(v, 0.025)),
        upper = as.numeric(quantile(v, 0.975)),
        p_lt_0 = mean(v < 0),
        p_gt_0 = mean(v > 0),
        p_in_rope = mean(abs(v) < rope_abs)
      )
    }
  }

  # ---- 3) 50→350 duration change at every session, then slice ----
  emm_dur_sess <- emmeans(model, ~ DurationF | CueType * SessionF)
  ct_dur_sess <- contrast(
    emm_dur_sess,
    method = list("350ms - 50ms" = w_350_minus_50),
    by = c("CueType", "SessionF")
  )
  draws_ds <- gather_emmeans_draws(ct_dur_sess)
  draws_ds$SessionF <- as.character(draws_ds$SessionF)

  # Average the 50→350 change across sessions within each posterior draw.
  slice_early <- draws_ds %>%
    filter(SessionF %in% early_sessions) %>%
    group_by(.draw, CueType) %>%
    summarise(.value = mean(.value), .groups = "drop")
  slice_average <- draws_ds %>%
    filter(SessionF %in% sess_levels) %>%
    group_by(.draw, CueType) %>%
    summarise(.value = mean(.value), .groups = "drop")
  slice_late <- draws_ds %>%
    filter(SessionF %in% late_sessions) %>%
    group_by(.draw, CueType) %>%
    summarise(.value = mean(.value), .groups = "drop")

  slice_draws <- list(
    early = slice_early,
    average = slice_average,
    late = slice_late
  )

  for (slice in c("early", "average", "late")) {
    wide <- slice_draws[[slice]] %>%
      pivot_wider(names_from = CueType, values_from = .value)
    v <- wide$R - wide$NR
    v <- v[is.finite(v)]
    stopifnot(length(v) > 0)
    row_i <- row_i + 1
    contrast_rows[[row_i]] <- tibble(
      ID = pid,
      effect = paste0("catchup_", slice),
      median = median(v),
      lower = as.numeric(quantile(v, 0.025)),
      upper = as.numeric(quantile(v, 0.975)),
      p_lt_0 = mean(v < 0),
      p_gt_0 = mean(v > 0),
      p_in_rope = mean(abs(v) < rope_abs)
    )
  }

  # Catch-up late − early
  wide_e <- slice_draws$early %>%
    pivot_wider(names_from = CueType, values_from = .value) %>%
    mutate(catchup = R - NR)
  wide_l <- slice_draws$late %>%
    pivot_wider(names_from = CueType, values_from = .value) %>%
    mutate(catchup = R - NR)
  v <- wide_l$catchup[match(wide_e$.draw, wide_l$.draw)] - wide_e$catchup
  v <- v[is.finite(v)]
  row_i <- row_i + 1
  contrast_rows[[row_i]] <- tibble(
    ID = pid,
    effect = "catchup_late_minus_early",
    median = median(v),
    lower = as.numeric(quantile(v, 0.025)),
    upper = as.numeric(quantile(v, 0.975)),
    p_lt_0 = mean(v < 0),
    p_gt_0 = mean(v > 0),
    p_in_rope = mean(abs(v) < rope_abs)
  )

  # ---- 4) Practice: late − early duration effect, by cue ----
  for (cue in c("NR", "R")) {
    early_df <- slice_draws$early %>% filter(CueType == cue) %>% arrange(.draw)
    late_df  <- slice_draws$late  %>% filter(CueType == cue) %>% arrange(.draw)
    stopifnot(identical(early_df$.draw, late_df$.draw))
    v <- late_df$.value - early_df$.value
    v <- v[is.finite(v)]
    stopifnot(length(v) > 0)
    row_i <- row_i + 1
    contrast_rows[[row_i]] <- tibble(
      ID = pid,
      effect = paste0(cue, ": duration_effect_late_minus_early"),
      median = median(v),
      lower = as.numeric(quantile(v, 0.025)),
      upper = as.numeric(quantile(v, 0.975)),
      p_lt_0 = mean(v < 0),
      p_gt_0 = mean(v > 0),
      p_in_rope = mean(abs(v) < rope_abs)
    )
  }
}

primary_summary <- bind_rows(contrast_rows) %>%
  arrange(effect, ID)

primary_summary

theory_focus <- primary_summary %>%
  filter(
    effect %in% c(
      "gain_at_50ms",
      "NR: linear_trend", "R: linear_trend",
      "NR: 350ms - 50ms", "R: 350ms - 50ms",
      "catchup_early", "catchup_average", "catchup_late",
      "catchup_late_minus_early",
      "NR: duration_effect_late_minus_early",
      "R: duration_effect_late_minus_early"
    )
  ) %>%
  mutate(
    question = case_when(
      effect == "gain_at_50ms" ~ "1. redundancy advantage",
      grepl("linear_trend|350ms - 50ms", effect) ~ "2. duration change by cue",
      grepl("^catchup_", effect) ~ "3. NR catch-up vs R",
      grepl("duration_effect_late_minus_early", effect) ~ "4. practice change of duration slope",
      TRUE ~ "other"
    )
  ) %>%
  arrange(question, effect, ID)

theory_focus

write_csv(theory_focus, "bayes_primary_discrete/theory_focus.csv")
write_csv(primary_summary, "bayes_primary_discrete/primary_summary.csv")
```

## 3.3 Posterior predictive checks

```{r bayes-ppc, fig.width=10, fig.height=7}
p_ecdf_AQ <- pp_check(m_AQ, ndraws = 50, type = "ecdf_overlay") +
  coord_cartesian(xlim = c(0, 180)) +
  labs(title = "AQ: ECDF", x = "AbsErr (deg)")
p_ecdf_HC <- pp_check(m_HC, ndraws = 50, type = "ecdf_overlay") +
  coord_cartesian(xlim = c(0, 180)) +
  labs(title = "HC: ECDF", x = "AbsErr (deg)")
p_ecdf_YILIU <- pp_check(m_YILIU, ndraws = 50, type = "ecdf_overlay") +
  coord_cartesian(xlim = c(0, 180)) +
  labs(title = "YILIU: ECDF", x = "AbsErr (deg)")

yrep_AQ <- pmin(pmax(posterior_predict(m_AQ, ndraws = 8), 0), 180)
yrep_HC <- pmin(pmax(posterior_predict(m_HC, ndraws = 8), 0), 180)
yrep_YILIU <- pmin(pmax(posterior_predict(m_YILIU, ndraws = 8), 0), 180)

p_hist_AQ <- bayesplot::ppc_hist(get_y(m_AQ), yrep_AQ, binwidth = 5) +
  coord_cartesian(xlim = c(0, 180)) +
  labs(title = "AQ: hist", x = "AbsErr (deg)")
p_hist_HC <- bayesplot::ppc_hist(get_y(m_HC), yrep_HC, binwidth = 5) +
  coord_cartesian(xlim = c(0, 180)) +
  labs(title = "HC: hist", x = "AbsErr (deg)")
p_hist_YILIU <- bayesplot::ppc_hist(get_y(m_YILIU), yrep_YILIU, binwidth = 5) +
  coord_cartesian(xlim = c(0, 180)) +
  labs(title = "YILIU: hist", x = "AbsErr (deg)")

(p_ecdf_AQ | p_ecdf_HC | p_ecdf_YILIU) /
  (p_hist_AQ | p_hist_HC | p_hist_YILIU)
```

## 3.4 Contrast plots

```{r bayes-contrast-plots, fig.width=10, fig.height=8}
plot_duration <- primary_summary %>%
  filter(effect %in% c("NR: 350ms - 50ms", "R: 350ms - 50ms")) %>%
  mutate(CueType = if_else(grepl("^NR", effect), "NR", "R"))

plot_catch <- primary_summary %>%
  filter(effect %in% c("catchup_early", "catchup_average", "catchup_late")) %>%
  mutate(
    SessionSlice = factor(
      gsub("catchup_", "", effect),
      levels = c("early", "average", "late")
    )
  )

plot_practice <- primary_summary %>%
  filter(grepl("duration_effect_late_minus_early", effect)) %>%
  mutate(CueType = if_else(grepl("^NR", effect), "NR", "R"))

p1 <- ggplot(plot_duration, aes(CueType, median, ymin = lower, ymax = upper)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_hline(yintercept = c(-rope_abs, rope_abs), linetype = "dotted", colour = "grey40") +
  geom_pointrange() +
  facet_wrap(~ID) +
  labs(
    title = "50→350 ms AbsErr change by cue (avg over sessions)",
    subtitle = "Negative = lower error at 350 ms; dotted = ±3° ROPE",
    y = "Posterior median (deg)", x = "Cue"
  )

p2 <- ggplot(plot_catch, aes(SessionSlice, median, ymin = lower, ymax = upper, group = 1)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_hline(yintercept = c(-rope_abs, rope_abs), linetype = "dotted", colour = "grey40") +
  geom_line() +
  geom_pointrange() +
  facet_wrap(~ID) +
  labs(
    title = "NR catch-up across session slices (discrete)",
    subtitle = "Early = S1–2; average = all sessions; late = S9–10",
    y = "Posterior median (deg)", x = "Session slice"
  )

p3 <- ggplot(plot_practice, aes(CueType, median, ymin = lower, ymax = upper)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_hline(yintercept = c(-rope_abs, rope_abs), linetype = "dotted", colour = "grey40") +
  geom_pointrange() +
  facet_wrap(~ID) +
  labs(
    title = "Practice change in 50→350 duration effect",
    subtitle = "Late (S9–10) − early (S1–2)",
    y = "Posterior median (deg)", x = "Cue"
  )

p1 / p2 / p3
```

# 4. Bayesian primary analysis: cirSD (discrete session)

Parallel to AbsErr, but the outcome is **circular SD** of signed errors.
cirSD is a cell summary (one value per Cue × Duration × Session). A saturated
`CueType * DurationF * SessionF` mean model on a single cirSD per cell would
have no residual degrees of freedom. To keep the **same design formula** as
AbsErr and still identify residual variance, we use **B bootstrap cirSD
replicates per cell** (resample trials within the cell, recompute cirSD).

Model (same fixed structure as AbsErr discrete):

`cirSD_deg ~ CueType * DurationF * SessionF`

Mixture modelling is **not** repeated here: the mixture already targets the
signed-error distribution directly (finer than AbsErr or cirSD summaries).

## 4.1 Build bootstrap cirSD dataset and fit / load models

```{r cirsd-bayes-data, message=FALSE, warning=FALSE}
rope_cir <- rope_abs
B_cir <- 40
set.seed(1234)

circ_sd_deg <- function(x) {
  x <- x[is.finite(x)]
  if (length(x) < 5) return(NA_real_)
  as.numeric(sd.circular(circular(x, units = "degrees"))) * (180 / pi)
}

build_cirsd_boot <- function(d_sub, B = B_cir) {
  rows <- list()
  i <- 0
  for (cue in levels(d_sub$CueType)) {
    for (dur in levels(d_sub$DurationF)) {
      for (sess_i in levels(d_sub$SessionF)) {
        idx <- which(
          d_sub$CueType == cue &
            d_sub$DurationF == dur &
            d_sub$SessionF == sess_i
        )
        x <- d_sub$SignedErr[idx]
        x <- x[is.finite(x)]
        stopifnot(length(x) >= 10)
        for (b in seq_len(B)) {
          i <- i + 1
          rows[[i]] <- tibble(
            CueType = cue,
            DurationF = dur,
            SessionF = sess_i,
            boot_id = b,
            cirSD_deg = circ_sd_deg(sample(x, replace = TRUE))
          )
        }
      }
    }
  }
  bind_rows(rows) %>%
    mutate(
      CueType = factor(CueType, levels = levels(d_sub$CueType)),
      DurationF = factor(DurationF, levels = levels(d_sub$DurationF)),
      SessionF = factor(SessionF, levels = levels(d_sub$SessionF))
    )
}

dir.create("bayes_primary_discrete", showWarnings = FALSE, recursive = TRUE)

cir_boot_files <- c(
  AQ = "bayes_primary_discrete/cirsd_boot_AQ.rds",
  HC = "bayes_primary_discrete/cirsd_boot_HC.rds",
  YILIU = "bayes_primary_discrete/cirsd_boot_YILIU.rds"
)
d_cir_AQ <- if (file.exists(cir_boot_files[["AQ"]])) readRDS(cir_boot_files[["AQ"]]) else {
  tmp <- build_cirsd_boot(d_AQ); saveRDS(tmp, cir_boot_files[["AQ"]]); tmp
}
d_cir_HC <- if (file.exists(cir_boot_files[["HC"]])) readRDS(cir_boot_files[["HC"]]) else {
  tmp <- build_cirsd_boot(d_HC); saveRDS(tmp, cir_boot_files[["HC"]]); tmp
}
d_cir_YILIU <- if (file.exists(cir_boot_files[["YILIU"]])) readRDS(cir_boot_files[["YILIU"]]) else {
  tmp <- build_cirsd_boot(d_YILIU); saveRDS(tmp, cir_boot_files[["YILIU"]]); tmp
}

stopifnot(nrow(d_cir_AQ) == 2 * 7 * 10 * B_cir)

bayes_priors_cir <- c(
  prior(normal(40, 20), class = "Intercept"),
  prior(normal(0, 15), class = "b"),
  prior(exponential(1), class = "sigma"),
  prior(gamma(2, 0.1), class = "nu")
)

m_cir_AQ <- brm(
  cirSD_deg ~ CueType * DurationF * SessionF,
  data = d_cir_AQ, family = student(), prior = bayes_priors_cir,
  chains = 4, cores = 4, iter = 4000, warmup = 2000,
  control = list(adapt_delta = 0.95, max_treedepth = 12), seed = 1234,
  file = "bayes_primary_discrete/m_AQ_cirsd", file_refit = "on_change"
)
m_cir_HC <- brm(
  cirSD_deg ~ CueType * DurationF * SessionF,
  data = d_cir_HC, family = student(), prior = bayes_priors_cir,
  chains = 4, cores = 4, iter = 4000, warmup = 2000,
  control = list(adapt_delta = 0.95, max_treedepth = 12), seed = 1234,
  file = "bayes_primary_discrete/m_HC_cirsd", file_refit = "on_change"
)
m_cir_YILIU <- brm(
  cirSD_deg ~ CueType * DurationF * SessionF,
  data = d_cir_YILIU, family = student(), prior = bayes_priors_cir,
  chains = 4, cores = 4, iter = 4000, warmup = 2000,
  control = list(adapt_delta = 0.95, max_treedepth = 12), seed = 1234,
  file = "bayes_primary_discrete/m_YILIU_cirsd", file_refit = "on_change"
)

summary(m_cir_AQ)
summary(m_cir_HC)
summary(m_cir_YILIU)
```

## 4.2 Theory contrasts (cirSD; same questions as AbsErr)

```{r cirsd-bayes-contrasts, message=FALSE, warning=FALSE}
contrast_rows_cir <- list()
row_i <- 0
models_cir <- list(AQ = m_cir_AQ, HC = m_cir_HC, YILIU = m_cir_YILIU)

for (pid in names(models_cir)) {
  model <- models_cir[[pid]]

  emm_gain <- emmeans(model, ~ CueType | DurationF)
  draws_gain <- gather_emmeans_draws(
    contrast(emm_gain, method = list("NR - R" = c(1, -1)), by = "DurationF")
  )

  for (dur in levels(d$DurationF)) {
    v <- draws_gain$.value[draws_gain$DurationF == dur]
    v <- v[is.finite(v)]
    stopifnot(length(v) > 0)
    row_i <- row_i + 1
    contrast_rows_cir[[row_i]] <- tibble(
      ID = pid, effect = paste0("gain_", dur),
      median = median(v), lower = as.numeric(quantile(v, 0.025)),
      upper = as.numeric(quantile(v, 0.975)),
      p_lt_0 = mean(v < 0), p_gt_0 = mean(v > 0),
      p_in_rope = mean(abs(v) < rope_cir)
    )
  }
  v <- draws_gain$.value[draws_gain$DurationF == "50ms"]
  v <- v[is.finite(v)]
  row_i <- row_i + 1
  contrast_rows_cir[[row_i]] <- tibble(
    ID = pid, effect = "gain_at_50ms",
    median = median(v), lower = as.numeric(quantile(v, 0.025)),
    upper = as.numeric(quantile(v, 0.975)),
    p_lt_0 = mean(v < 0), p_gt_0 = mean(v > 0),
    p_in_rope = mean(abs(v) < rope_cir)
  )

  draws_dur <- gather_emmeans_draws(contrast(
    emmeans(model, ~ DurationF | CueType),
    method = list("linear_trend" = w_linear, "350ms - 50ms" = w_350_minus_50),
    by = "CueType"
  ))
  for (cue in c("NR", "R")) {
    for (cname in c("linear_trend", "350ms - 50ms")) {
      v <- draws_dur$.value[draws_dur$CueType == cue & draws_dur$contrast == cname]
      v <- v[is.finite(v)]
      stopifnot(length(v) > 0)
      row_i <- row_i + 1
      contrast_rows_cir[[row_i]] <- tibble(
        ID = pid, effect = paste0(cue, ": ", cname),
        median = median(v), lower = as.numeric(quantile(v, 0.025)),
        upper = as.numeric(quantile(v, 0.975)),
        p_lt_0 = mean(v < 0), p_gt_0 = mean(v > 0),
        p_in_rope = mean(abs(v) < rope_cir)
      )
    }
  }

  draws_ds <- gather_emmeans_draws(contrast(
    emmeans(model, ~ DurationF | CueType * SessionF),
    method = list("350ms - 50ms" = w_350_minus_50),
    by = c("CueType", "SessionF")
  )) %>%
    ungroup() %>%
    mutate(SessionF = as.character(SessionF)) %>%
    select(.draw, CueType, SessionF, .value)

  slice_early <- draws_ds %>%
    filter(SessionF %in% early_sessions) %>%
    group_by(.draw, CueType) %>% summarise(.value = mean(.value), .groups = "drop")
  slice_average <- draws_ds %>%
    filter(SessionF %in% sess_levels) %>%
    group_by(.draw, CueType) %>% summarise(.value = mean(.value), .groups = "drop")
  slice_late <- draws_ds %>%
    filter(SessionF %in% late_sessions) %>%
    group_by(.draw, CueType) %>% summarise(.value = mean(.value), .groups = "drop")
  slice_draws <- list(early = slice_early, average = slice_average, late = slice_late)

  for (slice in c("early", "average", "late")) {
    wide <- slice_draws[[slice]] %>% pivot_wider(names_from = CueType, values_from = .value)
    v <- wide$R - wide$NR
    v <- v[is.finite(v)]
    stopifnot(length(v) > 0)
    row_i <- row_i + 1
    contrast_rows_cir[[row_i]] <- tibble(
      ID = pid, effect = paste0("catchup_", slice),
      median = median(v), lower = as.numeric(quantile(v, 0.025)),
      upper = as.numeric(quantile(v, 0.975)),
      p_lt_0 = mean(v < 0), p_gt_0 = mean(v > 0),
      p_in_rope = mean(abs(v) < rope_cir)
    )
  }

  wide_e <- slice_draws$early %>% pivot_wider(names_from = CueType, values_from = .value) %>% mutate(catchup = R - NR)
  wide_l <- slice_draws$late %>% pivot_wider(names_from = CueType, values_from = .value) %>% mutate(catchup = R - NR)
  v <- wide_l$catchup[match(wide_e$.draw, wide_l$.draw)] - wide_e$catchup
  v <- v[is.finite(v)]
  stopifnot(length(v) > 0)
  row_i <- row_i + 1
  contrast_rows_cir[[row_i]] <- tibble(
    ID = pid, effect = "catchup_late_minus_early",
    median = median(v), lower = as.numeric(quantile(v, 0.025)),
    upper = as.numeric(quantile(v, 0.975)),
    p_lt_0 = mean(v < 0), p_gt_0 = mean(v > 0),
    p_in_rope = mean(abs(v) < rope_cir)
  )

  for (cue in c("NR", "R")) {
    early_df <- slice_draws$early %>% filter(CueType == cue) %>% arrange(.draw)
    late_df <- slice_draws$late %>% filter(CueType == cue) %>% arrange(.draw)
    stopifnot(identical(early_df$.draw, late_df$.draw))
    v <- late_df$.value - early_df$.value
    v <- v[is.finite(v)]
    stopifnot(length(v) > 0)
    row_i <- row_i + 1
    contrast_rows_cir[[row_i]] <- tibble(
      ID = pid, effect = paste0(cue, ": duration_effect_late_minus_early"),
      median = median(v), lower = as.numeric(quantile(v, 0.025)),
      upper = as.numeric(quantile(v, 0.975)),
      p_lt_0 = mean(v < 0), p_gt_0 = mean(v > 0),
      p_in_rope = mean(abs(v) < rope_cir)
    )
  }
}

primary_summary_cir <- bind_rows(contrast_rows_cir) %>% arrange(effect, ID)
primary_summary_cir

theory_focus_cir <- primary_summary_cir %>%
  filter(effect %in% c(
    "gain_at_50ms",
    "NR: linear_trend", "R: linear_trend",
    "NR: 350ms - 50ms", "R: 350ms - 50ms",
    "catchup_early", "catchup_average", "catchup_late",
    "catchup_late_minus_early",
    "NR: duration_effect_late_minus_early",
    "R: duration_effect_late_minus_early"
  )) %>%
  mutate(question = case_when(
    effect == "gain_at_50ms" ~ "1. redundancy advantage",
    grepl("linear_trend|350ms - 50ms", effect) ~ "2. duration change by cue",
    grepl("^catchup_", effect) ~ "3. NR catch-up vs R",
    grepl("duration_effect_late_minus_early", effect) ~ "4. practice change of duration slope",
    TRUE ~ "other"
  )) %>%
  arrange(question, effect, ID)

theory_focus_cir

write_csv(theory_focus_cir, "bayes_primary_discrete/theory_focus_cirsd.csv")
write_csv(primary_summary_cir, "bayes_primary_discrete/primary_summary_cirsd.csv")
```

## 4.3 Posterior predictive checks (cirSD)

```{r cirsd-bayes-ppc, fig.width=10, fig.height=4}
(pp_check(m_cir_AQ, ndraws = 50, type = "ecdf_overlay") + labs(title = "AQ: cirSD ECDF", x = "cirSD (deg)")) |
  (pp_check(m_cir_HC, ndraws = 50, type = "ecdf_overlay") + labs(title = "HC: cirSD ECDF", x = "cirSD (deg)")) |
  (pp_check(m_cir_YILIU, ndraws = 50, type = "ecdf_overlay") + labs(title = "YILIU: cirSD ECDF", x = "cirSD (deg)"))
```

## 4.4 Contrast plots (cirSD)

```{r cirsd-bayes-plots, fig.width=10, fig.height=8}
plot_duration_cir <- primary_summary_cir %>%
  filter(effect %in% c("NR: 350ms - 50ms", "R: 350ms - 50ms")) %>%
  mutate(CueType = if_else(grepl("^NR", effect), "NR", "R"))
plot_catch_cir <- primary_summary_cir %>%
  filter(effect %in% c("catchup_early", "catchup_average", "catchup_late")) %>%
  mutate(SessionSlice = factor(gsub("catchup_", "", effect), levels = c("early", "average", "late")))
plot_practice_cir <- primary_summary_cir %>%
  filter(grepl("duration_effect_late_minus_early", effect)) %>%
  mutate(CueType = if_else(grepl("^NR", effect), "NR", "R"))

p1c <- ggplot(plot_duration_cir, aes(CueType, median, ymin = lower, ymax = upper)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_hline(yintercept = c(-rope_cir, rope_cir), linetype = "dotted", colour = "grey40") +
  geom_pointrange() + facet_wrap(~ID) +
  labs(title = "50→350 ms cirSD change by cue (avg over sessions)",
       subtitle = "Negative = lower cirSD at 350 ms; dotted = ±ROPE",
       y = "Posterior median (deg)", x = "Cue")
p2c <- ggplot(plot_catch_cir, aes(SessionSlice, median, ymin = lower, ymax = upper, group = 1)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_hline(yintercept = c(-rope_cir, rope_cir), linetype = "dotted", colour = "grey40") +
  geom_line() + geom_pointrange() + facet_wrap(~ID) +
  labs(title = "NR catch-up in cirSD across session slices",
       subtitle = "Positive = NR cirSD fell more than R (50→350)",
       y = "Posterior median (deg)", x = "Session slice")
p3c <- ggplot(plot_practice_cir, aes(CueType, median, ymin = lower, ymax = upper)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_hline(yintercept = c(-rope_cir, rope_cir), linetype = "dotted", colour = "grey40") +
  geom_pointrange() + facet_wrap(~ID) +
  labs(title = "Practice change in 50→350 cirSD duration effect",
       subtitle = "Late (S9–10) − early (S1–2)",
       y = "Posterior median (deg)", x = "Cue")
p1c / p2c / p3c
```

# 5. Pre-mixture 1: AbsErr Session 1 vs Session 10 by duration

With discrete `SessionF`, this is a direct session contrast (not a
Session_z slice).

```{r abserr-s1-s10, message=FALSE, warning=FALSE}
s1s10_rows <- list()
row_i <- 0

for (pid in names(models)) {
  model <- models[[pid]]

  emm <- emmeans(model, ~ SessionF | CueType * DurationF)
  ct <- contrast(
    emm,
    method = list("S1 - S10" = as.numeric(w_s1_minus_s10)),
    by = c("CueType", "DurationF")
  )
  draws <- gather_emmeans_draws(ct)

  for (cue in c("NR", "R")) {
    for (dur in levels(d$DurationF)) {
      v <- draws$.value[draws$CueType == cue & draws$DurationF == dur]
      v <- v[is.finite(v)]
      stopifnot(length(v) > 0)
      dur_ms <- as.numeric(gsub("ms", "", dur))
      row_i <- row_i + 1
      s1s10_rows[[row_i]] <- tibble(
        ID = pid,
        CueType = cue,
        DurationF = dur,
        Duration_ms = dur_ms,
        short_duration = dur_ms <= 150,
        median = median(v),
        lower = as.numeric(quantile(v, 0.025)),
        upper = as.numeric(quantile(v, 0.975)),
        p_gt_0 = mean(v > 0),
        p_lt_0 = mean(v < 0),
        p_in_rope = mean(abs(v) < rope_abs)
      )
    }

    short <- draws %>%
      filter(CueType == cue, DurationF %in% c("50ms", "100ms", "150ms")) %>%
      group_by(.draw) %>%
      summarise(.value = mean(.value), .groups = "drop")
    v <- short$.value[is.finite(short$.value)]
    row_i <- row_i + 1
    s1s10_rows[[row_i]] <- tibble(
      ID = pid,
      CueType = cue,
      DurationF = "short_mean",
      Duration_ms = NA_real_,
      short_duration = TRUE,
      median = median(v),
      lower = as.numeric(quantile(v, 0.025)),
      upper = as.numeric(quantile(v, 0.975)),
      p_gt_0 = mean(v > 0),
      p_lt_0 = mean(v < 0),
      p_in_rope = mean(abs(v) < rope_abs)
    )
  }
}

abserr_s1_s10 <- bind_rows(s1s10_rows)
abserr_s1_s10_cell <- abserr_s1_s10 %>%
  filter(DurationF != "short_mean") %>%
  arrange(ID, CueType, Duration_ms)
abserr_s1_s10_short <- abserr_s1_s10 %>%
  filter(DurationF == "short_mean") %>%
  arrange(ID, CueType)

abserr_s1_s10_cell
abserr_s1_s10_short

write_csv(abserr_s1_s10_cell, "bayes_primary_discrete/abserr_s1_vs_s10_by_duration.csv")
write_csv(abserr_s1_s10_short, "bayes_primary_discrete/abserr_s1_vs_s10_short.csv")
```

```{r abserr-s1-s10-plot, fig.width=10, fig.height=5}
ggplot(
  abserr_s1_s10_cell,
  aes(Duration_ms, median, ymin = lower, ymax = upper, colour = CueType, group = CueType)
) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_hline(yintercept = c(-rope_abs, rope_abs), linetype = "dotted", colour = "grey40") +
  geom_line() +
  geom_pointrange() +
  facet_wrap(~ID) +
  labs(
    title = "AbsErr Session 1 − Session 10 by duration (discrete SessionF)",
    subtitle = "Positive = higher AbsErr in Session 1; dotted = ±3° ROPE",
    x = "Encoding duration (ms)",
    y = "Posterior median difference (deg)"
  )
```

# 6. Pre-mixture 2: cirSD Session 1 vs Session 10 (bootstrap)

Companion in circular-SD units (same procedure as the continuous-session
script). For each participant × cue × duration: compute cirSD in Session
1 and Session 10, then bootstrap S1 − S10. Positive = higher circular
variability in Session 1.

```{r cirsd-s1-s10, message=FALSE, warning=FALSE}
B <- 10000  # bootstrap replicates
set.seed(1234)

ids <- c("AQ", "HC", "YILIU")
cues <- c("NR", "R")
durs <- levels(d$DurationF)

cirsd_rows <- list()
row_i <- 0

for (pid in ids) {
  for (cue in cues) {
    for (dur in durs) {
      # Trials in this cell for Session 1 and Session 10.
      x1 <- d$SignedErr[d$ID == pid & d$CueType == cue & d$DurationF == dur & d$Session == 1]
      x10 <- d$SignedErr[d$ID == pid & d$CueType == cue & d$DurationF == dur & d$Session == 10]
      x1 <- x1[is.finite(x1)]
      x10 <- x10[is.finite(x10)]

      # Need enough trials for a stable circular SD.
      if (length(x1) < 5 || length(x10) < 5) {
        row_i <- row_i + 1
        cirsd_rows[[row_i]] <- tibble(
          ID = pid, CueType = cue, DurationF = dur,
          Duration_ms = as.numeric(gsub("ms", "", dur)),
          estimate = NA_real_, lower = NA_real_, upper = NA_real_,
          boot_se = NA_real_, n1 = length(x1), n10 = length(x10),
          excludes_0 = NA
        )
        next
      }

      # Observed cirSD difference (degrees).
      sd1 <- as.numeric(sd.circular(circular(x1, units = "degrees"))) * (180 / pi)
      sd10 <- as.numeric(sd.circular(circular(x10, units = "degrees"))) * (180 / pi)
      obs <- sd1 - sd10

      # Bootstrap: resample each session's trials with replacement.
      boot_vals <- numeric(B)
      for (b in seq_len(B)) {
        s1_b <- sample(x1, replace = TRUE)
        s10_b <- sample(x10, replace = TRUE)
        sd1_b <- as.numeric(sd.circular(circular(s1_b, units = "degrees"))) * (180 / pi)
        sd10_b <- as.numeric(sd.circular(circular(s10_b, units = "degrees"))) * (180 / pi)
        boot_vals[b] <- sd1_b - sd10_b
      }
      boot_vals <- boot_vals[is.finite(boot_vals)]

      lo <- as.numeric(quantile(boot_vals, 0.025))
      hi <- as.numeric(quantile(boot_vals, 0.975))

      row_i <- row_i + 1
      cirsd_rows[[row_i]] <- tibble(
        ID = pid, CueType = cue, DurationF = dur,
        Duration_ms = as.numeric(gsub("ms", "", dur)),
        estimate = obs,
        lower = lo,
        upper = hi,
        boot_se = sd(boot_vals),
        n1 = length(x1),
        n10 = length(x10),
        excludes_0 = (lo > 0 & hi > 0) | (lo < 0 & hi < 0)
      )
    }
  }
}

cirsd_s1_s10 <- bind_rows(cirsd_rows) %>%
  arrange(ID, CueType, Duration_ms)

cirsd_s1_s10
write_csv(cirsd_s1_s10, "bayes_primary_discrete/cirsd_s1_vs_s10_bootstrap.csv")
```

```{r cirsd-s1-s10-short, message=FALSE, warning=FALSE}
# Short-duration companion: mean of cirSD(50/100/150) in S1 minus same mean in S10.
# Bootstrap resamples trials within each duration × session cell, then re-averages.
cirsd_short_rows <- list()
row_i <- 0
short_durs <- c("50ms", "100ms", "150ms")

for (pid in ids) {
  for (cue in cues) {
    # Observed mean cirSD across the three short durations.
    sd1_vec <- numeric(3)
    sd10_vec <- numeric(3)
    ok <- TRUE
    for (j in seq_along(short_durs)) {
      dur <- short_durs[j]
      x1 <- d$SignedErr[d$ID == pid & d$CueType == cue & d$DurationF == dur & d$Session == 1]
      x10 <- d$SignedErr[d$ID == pid & d$CueType == cue & d$DurationF == dur & d$Session == 10]
      x1 <- x1[is.finite(x1)]
      x10 <- x10[is.finite(x10)]
      if (length(x1) < 5 || length(x10) < 5) {
        ok <- FALSE
        break
      }
      sd1_vec[j] <- as.numeric(sd.circular(circular(x1, units = "degrees"))) * (180 / pi)
      sd10_vec[j] <- as.numeric(sd.circular(circular(x10, units = "degrees"))) * (180 / pi)
    }

    if (!ok) {
      row_i <- row_i + 1
      cirsd_short_rows[[row_i]] <- tibble(
        ID = pid, CueType = cue,
        estimate = NA_real_, lower = NA_real_, upper = NA_real_,
        boot_se = NA_real_, excludes_0 = NA
      )
      next
    }

    obs <- mean(sd1_vec) - mean(sd10_vec)

    boot_vals <- numeric(B)
    for (b in seq_len(B)) {
      sd1_b <- numeric(3)
      sd10_b <- numeric(3)
      for (j in seq_along(short_durs)) {
        dur <- short_durs[j]
        x1 <- d$SignedErr[d$ID == pid & d$CueType == cue & d$DurationF == dur & d$Session == 1]
        x10 <- d$SignedErr[d$ID == pid & d$CueType == cue & d$DurationF == dur & d$Session == 10]
        x1 <- x1[is.finite(x1)]
        x10 <- x10[is.finite(x10)]
        s1_b <- sample(x1, replace = TRUE)
        s10_b <- sample(x10, replace = TRUE)
        sd1_b[j] <- as.numeric(sd.circular(circular(s1_b, units = "degrees"))) * (180 / pi)
        sd10_b[j] <- as.numeric(sd.circular(circular(s10_b, units = "degrees"))) * (180 / pi)
      }
      boot_vals[b] <- mean(sd1_b) - mean(sd10_b)
    }
    boot_vals <- boot_vals[is.finite(boot_vals)]
    lo <- as.numeric(quantile(boot_vals, 0.025))
    hi <- as.numeric(quantile(boot_vals, 0.975))

    row_i <- row_i + 1
    cirsd_short_rows[[row_i]] <- tibble(
      ID = pid, CueType = cue,
      estimate = obs, lower = lo, upper = hi,
      boot_se = sd(boot_vals),
      excludes_0 = (lo > 0 & hi > 0) | (lo < 0 & hi < 0)
    )
  }
}

cirsd_short <- bind_rows(cirsd_short_rows) %>% arrange(ID, CueType)
cirsd_short
write_csv(cirsd_short, "bayes_primary_discrete/cirsd_s1_vs_s10_short_bootstrap.csv")
```

```{r cirsd-s1-s10-plot, fig.width=10, fig.height=5}
ggplot(
  cirsd_s1_s10,
  aes(Duration_ms, estimate, ymin = lower, ymax = upper, colour = CueType, group = CueType)
) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_line() +
  geom_pointrange() +
  facet_wrap(~ID) +
  labs(
    title = "cirSD Session 1 − Session 10 by duration",
    subtitle = "Bootstrap 95% CI; positive = higher circular variability in Session 1",
    x = "Encoding duration (ms)",
    y = "cirSD difference (deg)"
  )
```

# 7. Circular mixture: per-cell ML (target + guess + swap)

Bays-style mixture on signed errors, one independent ML fit per cue ×
duration cell (14 cells × participant; sessions pooled). Always 3 unique
non-target colours, so the swap term is three equal von Mises bumps:

`p(θ) = p_target · VM(0, κ) + (β/3) Σ_{j=1..3} VM(μ_j, κ) + γ · Uniform`

- Shared κ for target and colour-swap
- Stick-break order: estimate β (swap) first, then γ from the leftover
  (pure guesses expected rare); p_target = 1 − β − γ
- Fits from `Rscript s1_mixture_cells.R` →
  `mixture_cells/all_cell_params.csv`

```{r mixture-load}
mix_file <- "mixture_cells/all_cell_params.csv"
stopifnot(file.exists(mix_file))

mix_cells <- read_csv(mix_file, show_col_types = FALSE) %>%
  mutate(
    CueType = factor(CueType, levels = c("NR", "R")),
    DurationF = factor(
      DurationF,
      levels = c("50ms", "100ms", "150ms", "200ms", "250ms", "300ms", "350ms")
    ),
    Duration_ms = as.numeric(gsub("ms", "", as.character(DurationF)))
  )

stopifnot(all(c("AQ", "HC", "YILIU") %in% mix_cells$ID))
stopifnot(all(mix_cells$convergence == 0 | is.na(mix_cells$convergence)))
stopifnot(all(mix_cells$p_target + mix_cells$gamma + mix_cells$beta > 0.99))
stopifnot(all(mix_cells$p_target + mix_cells$gamma + mix_cells$beta < 1.01))

mix_cells %>%
  select(ID, CueType, DurationF, n_trials, kappa, circSD_deg,
         p_target, gamma, beta, nll, convergence) %>%
  mutate(across(c(kappa, circSD_deg, p_target, gamma, beta, nll), ~ round(.x, 3))) %>%
  arrange(ID, CueType, DurationF)
```

```{r mixture-proportion-plot, fig.width=10, fig.height=6}
prop_long <- mix_cells %>%
  select(ID, CueType, Duration_ms, p_target, gamma, beta) %>%
  pivot_longer(
    c(p_target, gamma, beta),
    names_to = "component",
    values_to = "proportion"
  ) %>%
  mutate(
    component = factor(
      component,
      levels = c("p_target", "beta", "gamma"),
      labels = c("Target", "Colour swap", "Guess")
    )
  )

ggplot(prop_long, aes(Duration_ms, proportion, fill = component)) +
  geom_col(width = 35, colour = "grey20", linewidth = 0.2) +
  facet_grid(CueType ~ ID) +
  scale_x_continuous(breaks = seq(50, 350, by = 50)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 1)) +
  labs(
    title = "Mixture response proportions by cue and duration",
    subtitle = "Independent ML fits per cell; sessions pooled",
    x = "Encoding duration (ms)",
    y = "Response proportion",
    fill = "Component"
  ) +
  theme(legend.position = "bottom")
```

```{r mixture-kappa-plot, fig.width=10, fig.height=4}
ggplot(mix_cells, aes(Duration_ms, kappa, colour = CueType, group = CueType)) +
  geom_line() +
  geom_point(size = 2) +
  facet_wrap(~ID, scales = "free_y") +
  scale_x_continuous(breaks = seq(50, 350, by = 50)) +
  labs(
    title = "Shared von Mises precision (κ) by cue and duration",
    x = "Encoding duration (ms)",
    y = "κ"
  )
```

```{r mixture-weight-lines, fig.width=10, fig.height=6}
ggplot(prop_long, aes(Duration_ms, proportion, colour = CueType, group = CueType)) +
  geom_line() +
  geom_point() +
  facet_grid(component ~ ID, scales = "free_y") +
  scale_x_continuous(breaks = seq(50, 350, by = 100)) +
  labs(
    title = "Mixture weights by cue and duration",
    x = "Encoding duration (ms)",
    y = "Proportion"
  )
```

# 8. Simple von Mises: 14-κ vs 2-κ (sessions pooled)

Assumption-light variability check (no mixture). Signed errors are treated as
a single von Mises centred at 0 (target). Sessions are pooled within cue ×
duration.

- **14-κ model:** one κ per CueType × DurationF cell (14 parameters)
- **2-κ model:** one κ for all NR trials, one κ for all R trials (2 parameters)

Question: does response variability change enough across encoding durations that
the 14-κ model wins on BIC despite many more parameters? If the 2-κ model wins,
duration modulation of κ is not large enough (under BIC) to justify separate
duration κs — not proof that duration effects are “artefacts.”

`BIC = −2 log L + k log(n)` with `n` = trials for that participant.

```{r simple-vm-14-vs-2, message=FALSE, warning=FALSE}
dir.create("bayes_primary_discrete", showWarnings = FALSE, recursive = TRUE)

# log von Mises density at mu = 0 (vectorised over theta; scalar kappa).
log_vm0 <- function(theta, kappa) {
  log_i0 <- log(max(besselI(kappa, 0, expon.scaled = TRUE), .Machine$double.xmin)) + kappa
  kappa * cos(theta) - log(2 * pi) - log_i0
}

# ML κ for one set of signed-error radians (mean fixed at 0).
fit_kappa_vm <- function(theta) {
  theta <- theta[is.finite(theta)]
  stopifnot(length(theta) > 20)

  nll <- function(log_k) {
    k <- exp(log_k)
    if (!is.finite(k) || k < 1e-6 || k > 5000) return(1e12)
    -sum(log_vm0(theta, k))
  }

  starts <- log(c(1, 2, 4, 8, 16, 32))
  best_nll <- Inf
  best_par <- NA_real_
  best_conv <- NA_integer_
  for (s0 in starts) {
    fit <- tryCatch(
      optim(par = s0, fn = nll, method = "BFGS",
            control = list(maxit = 1000, reltol = 1e-12)),
      error = function(e) NULL
    )
    if (is.null(fit)) next
    if (fit$value < best_nll) {
      best_nll <- fit$value
      best_par <- fit$par
      best_conv <- fit$convergence
    }
  }
  stopifnot(is.finite(best_par), best_conv == 0)

  kappa <- exp(best_par)
  # Circular SD (deg) from κ.
  a <- besselI(kappa, 1, expon.scaled = TRUE) /
    max(besselI(kappa, 0, expon.scaled = TRUE), .Machine$double.xmin)
  a <- min(max(a, 1e-12), 1 - 1e-12)
  list(
    kappa = kappa,
    circSD_deg = sqrt(-2 * log(a)) * (180 / pi),
    logLik = -best_nll,
    nll = best_nll,
    n = length(theta),
    convergence = best_conv
  )
}

# Signed errors in radians for each participant.
d_vm <- d %>%
  mutate(
    theta = atan2(sin(SignedErr * pi / 180), cos(SignedErr * pi / 180))
  )

vm_kappa_rows <- list()
vm_bic_rows <- list()
row_k <- 0
row_b <- 0

for (pid in c("AQ", "HC", "YILIU")) {
  d_pid <- d_vm %>% filter(ID == pid)
  n_pid <- nrow(d_pid)

  # ---- 14-κ: fit each cue × duration cell ----
  logLik_14 <- 0
  for (cue in c("NR", "R")) {
    for (dur in levels(d$DurationF)) {
      th <- d_pid$theta[d_pid$CueType == cue & d_pid$DurationF == dur]
      fit <- fit_kappa_vm(th)
      logLik_14 <- logLik_14 + fit$logLik
      row_k <- row_k + 1
      vm_kappa_rows[[row_k]] <- tibble(
        ID = pid,
        model = "14k",
        CueType = cue,
        DurationF = dur,
        Duration_ms = as.numeric(gsub("ms", "", dur)),
        n_trials = fit$n,
        kappa = fit$kappa,
        circSD_deg = fit$circSD_deg,
        logLik_cell = fit$logLik,
        convergence = fit$convergence
      )
    }
  }
  k_14 <- 14L
  bic_14 <- -2 * logLik_14 + k_14 * log(n_pid)
  aic_14 <- -2 * logLik_14 + 2 * k_14

  # ---- 2-κ: one κ per cue (all durations pooled) ----
  logLik_2 <- 0
  for (cue in c("NR", "R")) {
    th <- d_pid$theta[d_pid$CueType == cue]
    fit <- fit_kappa_vm(th)
    logLik_2 <- logLik_2 + fit$logLik
    row_k <- row_k + 1
    vm_kappa_rows[[row_k]] <- tibble(
      ID = pid,
      model = "2k",
      CueType = cue,
      DurationF = NA_character_,
      Duration_ms = NA_real_,
      n_trials = fit$n,
      kappa = fit$kappa,
      circSD_deg = fit$circSD_deg,
      logLik_cell = fit$logLik,
      convergence = fit$convergence
    )
  }
  k_2 <- 2L
  bic_2 <- -2 * logLik_2 + k_2 * log(n_pid)
  aic_2 <- -2 * logLik_2 + 2 * k_2

  # Positive delta_BIC_14_minus_2 => 2-κ preferred by BIC
  # (14-κ has higher BIC). Negative => 14-κ wins.
  row_b <- row_b + 1
  vm_bic_rows[[row_b]] <- tibble(
    ID = pid,
    n_trials = n_pid,
    logLik_14k = logLik_14,
    logLik_2k = logLik_2,
    k_14k = k_14,
    k_2k = k_2,
    BIC_14k = bic_14,
    BIC_2k = bic_2,
    AIC_14k = aic_14,
    AIC_2k = aic_2,
    delta_BIC_14_minus_2 = bic_14 - bic_2,
    delta_AIC_14_minus_2 = aic_14 - aic_2,
    BIC_winner = if_else(bic_14 < bic_2, "14k", "2k"),
    AIC_winner = if_else(aic_14 < aic_2, "14k", "2k")
  )
}

vm_kappa_table <- bind_rows(vm_kappa_rows)
vm_bic_table <- bind_rows(vm_bic_rows)

vm_kappa_table %>%
  mutate(across(c(kappa, circSD_deg, logLik_cell), ~ round(.x, 3))) %>%
  arrange(model, ID, CueType, Duration_ms)

vm_bic_table %>%
  mutate(across(where(is.numeric), ~ round(.x, 2)))

write_csv(vm_kappa_table, "bayes_primary_discrete/simple_vm_kappa_table.csv")
write_csv(vm_bic_table, "bayes_primary_discrete/simple_vm_14k_vs_2k_bic.csv")
```

```{r simple-vm-14-vs-2-plot, fig.width=10, fig.height=4}
# 14-κ estimates by duration; horizontal lines = 2-κ cue-level κ.
k14 <- vm_kappa_table %>% filter(model == "14k")
k2 <- vm_kappa_table %>% filter(model == "2k") %>%
  select(ID, CueType, kappa_2k = kappa)

ggplot(k14, aes(Duration_ms, kappa, colour = CueType, group = CueType)) +
  geom_hline(
    data = k2,
    aes(yintercept = kappa_2k, colour = CueType),
    linetype = "dashed", alpha = 0.7
  ) +
  geom_line() +
  geom_point(size = 2) +
  facet_wrap(~ID, scales = "free_y") +
  scale_x_continuous(breaks = seq(50, 350, by = 50)) +
  labs(
    title = "Simple von Mises κ: 14-κ (points/lines) vs 2-κ (dashed)",
    subtitle = "Sessions pooled; mean fixed at 0. Lower BIC wins (see table).",
    x = "Encoding duration (ms)",
    y = "κ"
  )
```

# 9. Mixture: κ free per cell vs κ tied within cue

After allowing swap (β) and guess (γ) to explain heavy tails, does **precision κ**
still need to vary with encoding duration?

- **κ-free:** existing per-cell mixture (14 × {κ, β, γ} = 42 parameters).
  logLik = sum of cell logLiks from `mixture_cells/`.
- **κ-tied:** one κ for NR and one κ for R across all durations; β and γ remain
  free in each of the 14 cells (2 + 14 + 14 = 30 parameters).

Sessions pooled. Fit κ-tied offline with `Rscript s1_mixture_kappa_compare.R`
(warm-started from the free cell fits).

`delta_BIC_free_minus_tied < 0` ⇒ κ-free wins (duration still matters for κ
after mixture decomposition).

```{r mixture-kappa-free-vs-tied, message=FALSE, warning=FALSE}
mix_bic_file <- "bayes_primary_discrete/mixture_kappa_free_vs_tied_bic.csv"
if (!file.exists(mix_bic_file)) {
  message("Running s1_mixture_kappa_compare.R (κ-tied fits) ...")
  system("Rscript s1_mixture_kappa_compare.R", intern = FALSE)
}
stopifnot(file.exists(mix_bic_file))

mix_kappa_bic <- read_csv(mix_bic_file, show_col_types = FALSE)
mix_kappa_tied_params <- read_csv(
  "bayes_primary_discrete/mixture_kappa_tied_params.csv",
  show_col_types = FALSE
)
mix_kappa_free_params <- read_csv(
  "bayes_primary_discrete/mixture_kappa_free_params.csv",
  show_col_types = FALSE
)

mix_kappa_bic %>%
  mutate(across(where(is.numeric), ~ round(.x, 2)))

# Side-by-side κ: free (by duration) vs tied (horizontal)
k_free_plot <- mix_kappa_free_params %>%
  mutate(
    Duration_ms = as.numeric(gsub("ms", "", as.character(DurationF))),
    CueType = factor(CueType, levels = c("NR", "R"))
  )
k_tied_plot <- mix_kappa_bic %>%
  select(ID, kappa_tied_NR, kappa_tied_R) %>%
  pivot_longer(
    c(kappa_tied_NR, kappa_tied_R),
    names_to = "CueType",
    values_to = "kappa_tied"
  ) %>%
  mutate(CueType = if_else(CueType == "kappa_tied_NR", "NR", "R"))
```

```{r mixture-kappa-free-vs-tied-plot, fig.width=10, fig.height=4}
ggplot(k_free_plot, aes(Duration_ms, kappa, colour = CueType, group = CueType)) +
  geom_hline(
    data = k_tied_plot,
    aes(yintercept = kappa_tied, colour = CueType),
    linetype = "dashed", alpha = 0.7
  ) +
  geom_line() +
  geom_point(size = 2) +
  facet_wrap(~ID, scales = "free_y") +
  scale_x_continuous(breaks = seq(50, 350, by = 50)) +
  labs(
    title = "Mixture κ: free per cell (lines) vs tied within cue (dashed)",
    subtitle = "β and γ free per cell in both models; sessions pooled",
    x = "Encoding duration (ms)",
    y = "κ"
  )
```

## Appendix

Legacy frequentist / exploratory Bayes material lives in
`s1_smallN_analysis.Rmd`. Primary discrete-session inference is Sections
3–9 above.
