---
title: "popcdm_fitting_1 — all participants"
output: html_document
---

Fit the redundancy POPCDM design model **once per participant** (`uid`).  
When you knit or run chunks, set the working directory to this folder (same directory as `data2_final.csv` and `fit_popcdm_redundancy.R`).

**Parameters (edit one chunk below):** use **named** starts / lower / upper for the **seven** optimised coordinates (`redundancy_Pvar_free_names()`). Set **fixed** CDM tail values (`eta1`, `eta2`, `ter`, `st`) in `fixed_eta_ter`; **`a` is free** here (seventh coordinate). `redundancy_fit_inputs_seven()` builds `P_var_start`, box bounds, and `P_fix` for the fitter.

**Grid / RT (step 1):** Before fitting, run `redundancy_rt_grid_audit()` with your candidate `tmax` and `h`. If `frac_rt_ge_tmax` is non-negligible, many trial log-likelihoods sit on the **interpolation floor** and parameters can be pulled to **bounds**.

**Multistart + logit β:** `beta_link = "logit"` maps **only** `beta1` and `beta2` (positions 4 and 5 in the seven-vector) through a logistic; `P_var_hat` is always physical.

**Iteration speed:** For pilots, set `TRIAL_CAP`; set `TRIAL_CAP <- NULL` for full data.

# Setup

```{r}
source("fit_popcdm_redundancy.R")
load_popcdm_sources()

data_csv <- "data2_final.csv"
if (!file.exists(data_csv)) {
  stop("Cannot find ", data_csv, " — set working directory to Data/Redundancy 2024")
}

uids <- sort(unique(utils::read.csv(data_csv, stringsAsFactors = FALSE)$uid))
uids
```

# RT and grid audit (tune `tmax`, `h`, `nw`)

Use the **same** `tmax`, `h`, and `nw` you will pass into the fitter.

```{r}
fit_df_all <- prepare_redundancy_fit_frame(data_csv, uid_filter = NULL, drop_na_cond = TRUE)

nw <- 100L
h <- 5 / 100
tmax <- 3.5

redundancy_rt_grid_audit(fit_df_all, tmax = tmax, h = h, nw = nw)
```

# Parameters (free seven-vector + fixed eta / ter / st)

Edit **`starts`**, **`lower`**, **`upper`** (all **named**; names must match `redundancy_Pvar_free_names()`). Edit **`fixed_eta_ter`** for the four slots that are **not** in `P_var` (everything except the seven free names is fixed via `P_fix`; `a` is taken from `starts["a"]`).

```{r}
fixed_eta_ter <- list(eta1 = 1, eta2 = 1e-6, ter = 0.2, st = 0.1)

starts <- c(
  alphaEst = 5.0,
  vnorm_R = 2.5,
  vnorm_NR = 2.2,
  beta1 = 0.5,
  beta2 = 0.5,
  kappa = 5.0,
  a = 2.0
)
lower <- c(
  alphaEst = 0.05,
  vnorm_R = 1.0,
  vnorm_NR = 1.0,
  beta1 = 0.05,
  beta2 = 0.05,
  kappa = 0.2,
  a = 0.1
)
upper <- c(
  alphaEst = 20.0,
  vnorm_R = 10.0,
  vnorm_NR = 10.0,
  beta1 = 1.0,
  beta2 = 1.0,
  kappa = 20.0,
  a = 10.0
)

fit_inp <- redundancy_fit_inputs_seven(starts, lower, upper, fixed_eta_ter = fixed_eta_ter)
P_var_start <- fit_inp$P_var_start
lower_b <- fit_inp$lower
upper_b <- fit_inp$upper
P_fix <- fit_inp$P_fix

fit_control <- list(maxit = 500L, factr = 1e7)

TRIAL_CAP <- 50L
n_starts <- 3L
ms_seed <- 1L
beta_link <- "logit"
```

# Fit each participant (multistart + logit β)

```{r, message=TRUE, warning=TRUE}
fits <- vector("list", length(uids))
names(fits) <- uids

for (uid in uids) {
  message("Fitting participant: ", uid)
  fit_df_full <- prepare_redundancy_fit_frame(data_csv, uid_filter = uid, drop_na_cond = TRUE)
  fit_df <- if (is.null(TRIAL_CAP)) {
    fit_df_full
  } else {
    redundancy_thin_trials_by_cond(fit_df_full, max_per_cond = TRIAL_CAP, seed = ms_seed)
  }
  fits[[uid]] <- fit_redundancy_popcdm_multistart(
    fit_df,
    P_var_start = P_var_start,
    lower = lower_b,
    upper = upper_b,
    P_fix = P_fix,
    n_starts = n_starts,
    seed = ms_seed + match(uid, uids),
    beta_link = beta_link,
    beta_latent_limit = 15,
    nw = nw,
    h = h,
    tmax = tmax,
    control = fit_control
  )
}
```

# Summary table

```{r}
nms <- redundancy_Pvar_free_names()

fit_summary <- do.call(
  rbind,
  lapply(uids, function(u) {
    f <- fits[[u]]
    pv <- unname(as.numeric(f$P_var_hat))
    if (length(pv) != length(nms)) {
      stop("Unexpected P_var length for ", u)
    }
    names(pv) <- nms
    eps <- 1e-5
    il <- which(abs(pv - lower_b) < eps)
    iu <- which(abs(upper_b - pv) < eps)
    on_lower <- if (length(il)) paste(nms[il], collapse = ",") else NA_character_
    on_upper <- if (length(iu)) paste(nms[iu], collapse = ",") else NA_character_
    ms <- f$multistart_detail
    data.frame(
      uid = u,
      n_trials = f$n_trials,
      nll = f$nll,
      n_starts = ms$n_starts,
      best_start_index = ms$best_index,
      nll_spread = max(ms$nll_each) - min(ms$nll_each),
      np_free = f$np_free,
      convergence = f$optim$convergence,
      alphaEst = pv["alphaEst"],
      vnorm_R = pv["vnorm_R"],
      vnorm_NR = pv["vnorm_NR"],
      beta1 = pv["beta1"],
      beta2 = pv["beta2"],
      kappa = pv["kappa"],
      a = pv["a"],
      at_lower = on_lower,
      at_upper = on_upper,
      stringsAsFactors = FALSE
    )
  })
)
rownames(fit_summary) <- NULL
fit_summary
```

# When parameters still hit bounds (diagnostics)

Pass the **same** `P_fix` as in the fit. Alpha / multiplier logic lives in **`redundancy_nll()`** in `fit_popcdm_redundancy.R` (function `redundancy_branch_m` sits directly above it).

```{r, eval=FALSE}
uid <- uids[[1]]
fit_df <- prepare_redundancy_fit_frame(data_csv, uid_filter = uid, drop_na_cond = TRUE)
fit_df <- redundancy_thin_trials_by_cond(fit_df, max_per_cond = 80L, seed = 1L)
P0 <- P_var_start
redundancy_mesh_hit_rates(P0, fit_df, P_fix = P_fix, nw = 48L, h = 0.008, tmax = 5)
redundancy_curve_nll_1d(1L, seq(0.2, 8, length.out = 25L), P0, fit_df, P_fix = P_fix, nw = 48L, h = 0.008, tmax = 5)
```

# Per-participant objects

`fits[["YL"]]` holds `optim`, `P_hat`, **`P_var_hat`** (physical seven-vector in order: alphaEst, vnorm_R, vnorm_NR, beta1, beta2, kappa, a), `nll`, `multistart_detail`, etc.

```{r, eval=FALSE}
fits[["YL"]]$P_var_hat
fits[["YL"]]$P_hat
```
