---
title: "POPCDM Figure Walkthrough (ggplot)"
output: html_document
---

This notebook recreates the original POPCDM walkthrough figure using `ggplot2`, with each panel in a separate chunk so you can run and inspect them one-by-one.

```{r setup}
source("popcdm300.R")
library(ggplot2)
```

## Shared helpers

```{r helpers}
marginal_rt_from_gt <- function(Theta, Tvec, Gt) {
  w <- 2 * pi / length(Theta)
  h <- if (length(Tvec) > 1L) diff(Tvec)[1] else 1
  pt <- colSums(Gt) * w
  pt <- pmax(pt, 0)
  pt <- pt / sum(pt * h)
  data.frame(t = Tvec, dens = pt)
}
```

## Run one POPCDM configuration

```{r model-run}
nw <- 50L
h <- 2.5 / 300
tmax <- 2.5

P_pop <- c(alpha = 1.0, kappa = 20.0)
pc <- popcode(P_pop, nw = nw)

P_full <- c(
  vnorm = 5.0, eta1 = 0.5, eta2 = 0.001, a = 5.0,
  alpha = 10.0, kappa = 20.0, ter = 0.0, st = 0.0
)
out <- popcdm300(P_full, nw = nw, h = h, tmax = tmax, return_components = TRUE)

pang_n <- pc$pang / max(pc$pang)
Ptheta_n <- out$Ptheta / max(out$Ptheta)
rt_df <- marginal_rt_from_gt(out$Theta, out$T, out$Gt)

polar_pang <- data.frame(
  theta = pc$th,
  xend = pang_n * cos(pc$th),
  yend = pang_n * sin(pc$th),
  value = pang_n
)

circle_df <- data.frame(
  phi = seq(0, 2 * pi, length.out = 361)
)
circle_df$x <- cos(circle_df$phi)
circle_df$y <- sin(circle_df$phi)

theta_deg <- out$Theta * 180 / pi
comp_df <- do.call(
  rbind,
  lapply(seq_len(nw), function(i) {
    data.frame(
      source_bin = i,
      theta_deg = theta_deg,
      ptheta = out$Ptheta_components[i, ],
      pang_weight = out$Pang[i],
      ptheta_weighted = out$Pang[i] * out$Ptheta_components[i, ]
    )
  })
)

# RT component decomposition:
# under current shift-mix construction, shifting is along theta rows only, so each source-bin
# component has the same RT shape before Pang weighting.
sigma <- 1.0
Theta_model <- out$Theta
v1 <- P_full["vnorm"] * cos(Theta_model[1])
v2 <- P_full["vnorm"] * sin(Theta_model[1])
Pi <- c(v1, v2, P_full["eta1"], P_full["eta2"], sigma, P_full["a"])
base <- cdm_core(Pi, nw = nw, h = h, tmax = tmax)
rt_component <- marginal_rt_from_gt(base$Theta, base$T, base$Gt)
rt_comp_df <- do.call(
  rbind,
  lapply(seq_len(nw), function(i) {
    data.frame(
      source_bin = i,
      t = rt_component$t,
      rt_comp = rt_component$dens,
      pang_weight = out$Pang[i],
      rt_comp_weighted = out$Pang[i] * rt_component$dens
    )
  })
)

gt_df <- expand.grid(
  theta_deg = theta_deg,
  t = out$T
)
gt_df$gt <- as.vector(out$Gt)
```

## Panel 1: `pang` (population drift-direction weights)

**Figure title:** `pang` (population drift-direction weights)  
**Figure subtitle:** Population-layer PMF over drift directions.  
**Figure description:** Spokes show normalized `pang` values on the circular angle grid before CDM mixing.

```{r panel-1-pang}
ggplot() +
  geom_path(data = circle_df, aes(x = x, y = y), color = "grey80", linewidth = 0.6) +
  geom_hline(yintercept = 0, color = "grey92") +
  geom_vline(xintercept = 0, color = "grey92") +
  geom_segment(
    data = polar_pang,
    aes(x = 0, y = 0, xend = xend, yend = yend),
    color = "#3366CC", linewidth = 0.8
    ) +
  coord_equal(xlim = c(-1.1, 1.1), ylim = c(-1.1, 1.1)) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank()
  )
```

## Panel 2: `Ptheta` components (all bins, colored)

**Figure title:** `Ptheta` components (all bins, colored)  
**Figure subtitle:** Per-source-bin angle marginals before `Pang` weighting.  
**Figure description:** Each colored line is one shifted component `Ptheta_i(theta)` from `out$Ptheta_components`.

```{r panel-5-components-top}
comp_all <- comp_df
comp_all$source_bin <- factor(comp_all$source_bin)

ggplot(comp_all, aes(x = theta_deg, y = ptheta, color = source_bin, group = source_bin)) +
  geom_line(linewidth = 0.6, alpha = 0.8) +
  labs(
    x = "Theta (deg)",
    y = "Component Ptheta",
    color = "Source bin i"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none")
```

## Panel 3: weighted components for all bins

**Figure title:** Weighted `Ptheta` components (all bins)  
**Figure subtitle:** Source-bin contributions after `Pang` weighting.  
**Figure description:** Each line is `Pang[i] * Ptheta_i(theta)`; these contributions sum to final `out$Ptheta`.

```{r panel-6-components-weighted-all}
comp_all_w <- comp_df
comp_all_w$source_bin <- factor(comp_all_w$source_bin)

ggplot(comp_all_w, aes(x = theta_deg, y = ptheta_weighted, color = source_bin, group = source_bin)) +
  geom_line(linewidth = 0.55, alpha = 0.8) +
  labs(
    x = "Theta (deg)",
    y = "Weighted component contribution",
    color = "Source bin i"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none")
```

## Panel 4: weighted RT components for all bins

**Figure title:** Weighted RT components (all bins)  
**Figure subtitle:** RT-side contributions after `Pang` weighting.  
**Figure description:** Under current shift-mix implementation, RT component shape is shared across bins and scaled by `Pang[i]`.

```{r panel-9-rt-components-weighted}
rt_comp_all <- rt_comp_df
rt_comp_all$source_bin <- factor(rt_comp_all$source_bin)

ggplot(rt_comp_all, aes(x = t, y = rt_comp_weighted, color = source_bin, group = source_bin)) +
  geom_line(linewidth = 0.55, alpha = 0.8) +
  labs(
    x = "Time (s)",
    y = "Weighted RT contribution",
    color = "Source bin i"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none")
```

## Panel 5: weighted contributions (all bins) + final `Ptheta`

**Figure title:** Weighted `Ptheta` contributions and final `Ptheta`  
**Figure subtitle:** Sum-of-components reconstruction in angle space.  
**Figure description:** Colored lines are `Pang[i] * Ptheta_i(theta)`; black line is the final mixed `out$Ptheta`.

```{r panel-7-components-weighted-top}
comp_all_w <- comp_df
comp_all_w$source_bin <- factor(comp_all_w$source_bin)
ptheta_final_df <- data.frame(theta_deg = theta_deg, ptheta_final = out$Ptheta)

ggplot() +
  geom_line(
    data = comp_all_w,
    aes(x = theta_deg, y = ptheta_weighted, color = source_bin, group = source_bin),
    linewidth = 0.5, alpha = 0.45
  ) +
  geom_line(
    data = ptheta_final_df,
    aes(x = theta_deg, y = ptheta_final),
    color = "black", linewidth = 1.1
  ) +
  labs(
    x = "Theta (deg)",
    y = "Contribution / final Ptheta",
    color = "Source bin i"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none")
```

## Panel 6: weighted RT contributions + final RT marginal

**Figure title:** Weighted RT contributions and final RT marginal  
**Figure subtitle:** Sum-of-components reconstruction in time space.  
**Figure description:** Colored lines are `Pang[i] * RT_component_i(t)`; black line is the final RT marginal from mixed `Gt`.

```{r panel-10-rt-components-weighted-plus-final}
rt_final_df <- data.frame(t = rt_df$t, rt_final = rt_df$dens)

ggplot() +
  geom_line(
    data = rt_comp_all,
    aes(x = t, y = rt_comp_weighted, color = source_bin, group = source_bin),
    linewidth = 0.5, alpha = 0.45
  ) +
  geom_line(
    data = rt_final_df,
    aes(x = t, y = rt_final),
    color = "black", linewidth = 1.1
  ) +
  labs(
    x = "Time (s)",
    y = "Contribution / final RT marginal",
    color = "Source bin i"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none")
```

## Panel 7: `Gt(theta, t)` in 3D

**Figure title:** 3D surface of `Gt(theta, t)`  
**Figure subtitle:** Joint density over response angle and decision time.  
**Figure description:** Height encodes the joint density value at each `(theta, t)` grid location.

```{r panel-4b-gt-3d, fig.width=11, fig.height=8, out.width='100%'}
# persp() expects z with rows aligned to x and columns to y.
# out$Gt is [theta_index, time_index], so we can pass it directly.
theta_deg_vec <- theta_deg
time_vec <- out$T
z_mat <- out$Gt

oldpar <- par(no.readonly = TRUE)
on.exit(par(oldpar), add = TRUE)
par(bg = "white", mar = c(3.5, 3.5, 2, 1))

persp(
  x = theta_deg_vec,
  y = time_vec,
  z = z_mat,
  theta = 35,
  phi = 30,
  expand = 0.4,
  col = "lightblue",
  shade = 0.45,
  border = NA,
  ticktype = "detailed",
  xlab = "Theta (deg)",
  ylab = "Time (s)",
  zlab = "Gt"
)
```

## Appendix: Additional `Gt(theta, t)` Views

### `Gt(theta, t)` as a heatmap

**Figure title:** `Gt(theta, t)` heatmap  
**Figure subtitle:** Joint density in 2D color map form.  
**Figure description:** Color indicates density magnitude across theta-time grid cells.

```{r panel-a1-gt-heatmap}
ggplot(gt_df, aes(x = theta_deg, y = t, fill = gt)) +
  geom_raster(interpolate = TRUE) +
  scale_fill_viridis_c(option = "C") +
  labs(
    x = "Theta (deg)",
    y = "Time (s)",
    fill = "Gt"
  ) +
  theme_minimal(base_size = 12)
```

### `Gt(theta, t)` contour view

**Figure title:** `Gt(theta, t)` contours  
**Figure subtitle:** Iso-density contours of the same joint surface.  
**Figure description:** Filled contour bands show equal-density regions and highlight the geometry of the joint.

```{r panel-a2-gt-contour}
ggplot(gt_df, aes(x = theta_deg, y = t, z = gt)) +
  geom_contour_filled(bins = 12) +
  labs(
    x = "Theta (deg)",
    y = "Time (s)",
    fill = "Gt level"
  ) +
  theme_minimal(base_size = 12)
```

## Optional: print parameter values used above

```{r params}
P_pop
P_full
```

