---
title: "fit2"
output: html_document
---

# CI-Grouped Redundancy POPCDM Fit

Clean POPCDM fitting workflow for `data2_final.csv`.

Switch **`beta_variant`** in the config chunk to compare beta schemes:

- `down_up`: shared down-scaling (`beta_down`), separate up-scaling for R (`beta_up`)
- `base_nr_r`: separate exponents for baseline, NR, and R (`beta_baseline`, `beta_nr`, `beta_r`)

Other constraints (your current setup): singleton `kappa`, `a`, `eta1`; `vnorm` and `ter` grouped by colour-count.

**Amplitude scaling:** `alphaEst` is the theoretical resource baseline \(B\). Trial precision uses  
`alpha_trial = alphaEst * (ColorNi / color_count_ref)^(-beta) * ...` with default `color_count_ref = 2` (smallest set in the design), matching `amplitude_test.Rmd` method 2.

```{r setup}
rm(list = ls())

data_csv <- "data2_final.csv"
helper_file <- "fit_popcdm_redundancy.R"

POPCDM_R_DIR <- Sys.getenv(
  "POPCDM_R_DIR",
  "C:/Users/Yuda/Documents/GitHub/DazExperiments/Models/CDM/POPCDM_R"
)

load_popcdm_sources <- function(dir = POPCDM_R_DIR) {
  if (!dir.exists(dir)) {
    stop("POPCDM_R_DIR not found: ", dir, call. = FALSE)
  }
  owd <- getwd()
  on.exit(setwd(owd), add = TRUE)
  setwd(dir)
  source("popcode.R", local = FALSE)
  source("besselFPT.R", local = FALSE)
  source("popcdm300.R", local = FALSE)
}

if (!file.exists(data_csv)) {
  stop("Cannot find ", data_csv, ". Set the working directory to the POPCDM folder.")
}
if (!file.exists(helper_file)) {
  stop("Cannot find ", helper_file, ".")
}

source(helper_file)
load_popcdm_sources()
```

# User-editable settings

```{r config}
# --- Beta model switch (change this line to compare variants) ---
beta_variant <- "base_nr_r"   # "down_up" | "base_nr_r"

# Use NULL for all participants, or e.g. c("YL", "PG") for a subset.
participant_ids <- "AQ"

trial_cap_per_condition <- 200L
random_seed <- 1L

grid <- list(
  nw = 50L,
  h = 3.4 / 200,
  tmax = 3.4
)

optim_settings <- list(
  n_starts = 1L,
  # If betas keep hitting bounds with logit, try beta_link = "identity".
  beta_link = "logit",
  beta_latent_limit = 15,
  progress = TRUE,
  progress_every = 5L,
  control = list(maxit = 200L, factr = 1e7)
)

parallel_settings <- list(
  use_parallel = TRUE,
  n_cores = min(parallel::detectCores(logical = FALSE), 5L, na.rm = TRUE)
)
```

# Build model inputs

```{r model-spec}
P_start <- redundancy_default_P(
  beta_variant = beta_variant,
  alphaEst = 2.0,
  vnorm = 2.5,
  kappa = 5.0,
  a = 2.0,
  ter = 0.2,
  eta1 = 1.0,
  eta2 = 1e-6,
  st = 0.2
)

bounds <- redundancy_default_bounds(beta_variant = beta_variant)

Sel <- redundancy_default_sel(
  beta_variant = beta_variant,
  free_eta2 = FALSE,
  free_st = TRUE
)

ci2_conditions <- c(
  "set2_baseline",
  "set4_R_cue",
  "set4_NR_cue",
  "set6_R_cue_ci2",
  "set6_NR_cue_ci2"
)
ci4_conditions <- c(
  "set4_baseline",
  "set6_R_cue_ci4",
  "set6_NR_cue_ci4"
)

m_ci_t <- redundancy_model_spec(
  label = paste0("ci_vet_", beta_variant),
  beta_variant = beta_variant,
  color_count_ref = 2L,
  equal_groups = list(
    redundancy_equal_group("a", COND_LEVELS),
    redundancy_equal_group("kappa", COND_LEVELS),
    redundancy_equal_group("vnorm", COND_LEVELS),
    redundancy_equal_group("ter", ci2_conditions),
    redundancy_equal_group("ter", ci4_conditions),
    redundancy_equal_group("eta1", COND_LEVELS)
  )
)

fit_inputs <- redundancy_fit_inputs(
  P_start = P_start,
  lower = bounds$lower,
  upper = bounds$upper,
  Sel = Sel,
  model_spec = m_ci_t
)

length(P_start)
length(fit_inputs$P_var_start)
names(fit_inputs$P_var_start)
length(fit_inputs$P_fix)
stopifnot(
  beta_variant == redundancy_resolve_beta_variant(fit_inputs$model_spec),
  length(fit_inputs$P_var_start) == sum(fit_inputs$Sel)
)

# Pre-fit: every condition must get a valid alpha multiplier m at starting values
branch_m_audit <- redundancy_branch_m_audit(P_start, model_spec = m_ci_t)
branch_m_audit
if (!all(branch_m_audit$ok)) {
  stop(
    "Invalid branch multiplier m at P_start for some conditions (see branch_m_audit). ",
    "Re-source fit_popcdm_redundancy.R and check beta_variant matches P_start/Sel.",
    call. = FALSE
  )
}
```

# Prepare and audit data

```{r data-audit}
raw_ids <- sort(unique(utils::read.csv(data_csv, stringsAsFactors = FALSE)$uid))
uids_to_fit <- if (is.null(participant_ids)) raw_ids else participant_ids
uids_to_fit

fit_df_all <- prepare_redundancy_fit_frame(
  data_csv,
  uid_filter = uids_to_fit,
  drop_na_cond = TRUE
)

table(fit_df_all$uid, fit_df_all$cond_label)

redundancy_rt_grid_audit(
  fit_df_all,
  tmax = grid$tmax,
  h = grid$h,
  nw = grid$nw
)
```

# Fit selected participants

```{r fit-pilot, message=TRUE, warning=TRUE}
fit_one_participant <- function(uid) {
  message("Fitting participant: ", uid, " (beta_variant=", beta_variant, ")")
  fit_df_full <- prepare_redundancy_fit_frame(
    data_csv,
    uid_filter = uid,
    drop_na_cond = TRUE
  )
  fit_df <- if (is.null(trial_cap_per_condition)) {
    fit_df_full
  } else {
    redundancy_thin_trials_by_cond(
      fit_df_full,
      max_per_cond = trial_cap_per_condition,
      seed = random_seed
    )
  }

  fit_redundancy_popcdm_multistart(
    fit_df,
    P_var_start = fit_inputs$P_var_start,
    lower = fit_inputs$lower,
    upper = fit_inputs$upper,
    P_fix = fit_inputs$P_fix,
    Sel = fit_inputs$Sel,
    model_spec = fit_inputs$model_spec,
    n_starts = optim_settings$n_starts,
    seed = random_seed + match(uid, uids_to_fit),
    beta_link = optim_settings$beta_link,
    beta_latent_limit = optim_settings$beta_latent_limit,
    progress = optim_settings$progress,
    progress_every = optim_settings$progress_every,
    nw = grid$nw,
    h = grid$h,
    tmax = grid$tmax,
    control = optim_settings$control
  )
}

if (isTRUE(parallel_settings$use_parallel) && length(uids_to_fit) > 1L) {
  n_cores <- min(parallel_settings$n_cores, length(uids_to_fit))
  message("Fitting ", length(uids_to_fit), " participants using ", n_cores, " cores.")
  fits <- parallel::mclapply(
    uids_to_fit,
    fit_one_participant,
    mc.cores = n_cores,
    mc.set.seed = TRUE
  )
  names(fits) <- uids_to_fit
} else {
  fits <- lapply(uids_to_fit, fit_one_participant)
  names(fits) <- uids_to_fit
}
```

# Summarise fitted parameters

```{r summary}
fit_summary <- do.call(
  rbind,
  lapply(uids_to_fit, function(uid) {
    f <- fits[[uid]]
    data.frame(
      uid = uid,
      beta_variant = redundancy_resolve_beta_variant(f$model_spec),
      model = f$model_spec$label,
      n_trials = f$n_trials,
      np_free = f$np_free,
      nll = f$nll,
      AIC = 2 * f$np_free + 2 * f$nll,
      convergence = f$optim$convergence,
      stringsAsFactors = FALSE
    )
  })
)
fit_summary

P_hats <- lapply(fits, `[[`, "P_hat")
P_hats[[uids_to_fit[1]]]

# Beta bound diagnostics (physical scale)
beta_bound_reports <- lapply(uids_to_fit, function(uid) {
  f <- fits[[uid]]
  br <- f$beta_bound_report
  if (nrow(br)) {
    br$uid <- uid
  }
  br
})
beta_bound_reports <- do.call(rbind, beta_bound_reports)
if (is.null(beta_bound_reports) || !nrow(beta_bound_reports)) {
  message("No beta parameters on/near bounds (good).")
} else {
  beta_bound_reports
}
```

# Condition-specific parameter table

```{r condition-parameter-table}
beta_cols <- redundancy_beta_names(beta_variant)

condition_parameter_table <- do.call(
  rbind,
  lapply(uids_to_fit, function(uid) {
    P_hat <- P_hats[[uid]]
    do.call(
      rbind,
      lapply(COND_LEVELS, function(cl) {
        row <- data.frame(
          uid = uid,
          cond_label = cl,
          trial_kind = redundancy_trial_kind(cl),
          alphaEst = P_hat["alphaEst"],
          vnorm = P_hat[paste0("vnorm_", cl)],
          kappa = P_hat[paste0("kappa_", cl)],
          a = P_hat[paste0("a_", cl)],
          ter = P_hat[paste0("ter_", cl)],
          eta1 = P_hat[paste0("eta1_", cl)],
          stringsAsFactors = FALSE,
          row.names = NULL
        )
        for (bn in beta_cols) {
          row[[bn]] <- P_hat[bn]
        }
        row
      })
    )
  })
)
numeric_cols <- vapply(condition_parameter_table, is.numeric, logical(1))
condition_parameter_table[numeric_cols] <- lapply(
  condition_parameter_table[numeric_cols],
  round,
  digits = 3
)
condition_parameter_table
```

# Plot predictions against data

```{r plot-data-source}
source(helper_file)
```

```{r plot-error-marginals, fig.width=10, fig.height=8, results='asis'}
for (plot_uid in uids_to_fit) {
  cat("\n\n### Error Marginals: ", plot_uid, "\n\n", sep = "")
  plot_df <- prepare_redundancy_fit_frame(
    data_csv,
    uid_filter = plot_uid,
    drop_na_cond = TRUE
  )
  plot_redundancy_error_marginals(
    plot_df,
    P_hats[[plot_uid]],
    model_spec = fits[[plot_uid]]$model_spec,
    nw = grid$nw,
    h = grid$h,
    tmax = grid$tmax,
    n_bins = 30L
  )
}
```

```{r plot-rt-marginals, fig.width=10, fig.height=8, results='asis'}
for (plot_uid in uids_to_fit) {
  cat("\n\n### RT Marginals: ", plot_uid, "\n\n", sep = "")
  plot_df <- prepare_redundancy_fit_frame(
    data_csv,
    uid_filter = plot_uid,
    drop_na_cond = TRUE
  )
  plot_redundancy_rt_marginals(
    plot_df,
    P_hats[[plot_uid]],
    model_spec = fits[[plot_uid]]$model_spec,
    nw = grid$nw,
    h = grid$h,
    tmax = grid$tmax,
    n_bins = 30L
  )
}
```
