---
title: "fit3_scaling_branch"
output: html_document
---

# Alpha/Kappa Scaling Branch

Work note with full vector:

- `alpha_*` x 9 condition values
- `kappa_*` x 9 condition values
- `eta2_*` x 9 condition values

Use `scaling_mode` to decide whether beta scaling applies to `alpha`, to
`kappa`, or to neither:

- `alpha_beta`: `alpha_trial = alpha_condition * m(beta)`, `kappa` unscaled
- `kappa_beta`: `kappa_trial = kappa_condition * m(beta)`, `alpha` unscaled
- `none`: no beta scaling; both `alpha` and `kappa` are direct condition values

```{r setup}
rm(list = ls())

library(beepr)

data_csv <- "data2_final.csv"
base_helper_file <- "fit_popcdm_redundancy.R"
branch_helper_file <- "fit_popcdm_redundancy_scaling_branch.R"

POPCDM_R_DIR <- Sys.getenv(
  "POPCDM_R_DIR",
  "/Users/prefabteam_ysl/Documents/GitHub/DazExperiments/Models/CDM/POPCDM_R"
)

load_popcdm_sources <- function(dir = POPCDM_R_DIR) {
  if (!dir.exists(dir)) {
    stop("POPCDM_R_DIR not found: ", dir, call. = FALSE)
  }
  owd <- getwd()
  on.exit(setwd(owd), add = TRUE)
  setwd(dir)
  source("popcode.R", local = FALSE)
  source("besselFPT.R", local = FALSE)
  source("popcdm300.R", local = FALSE)
}

stopifnot(file.exists(data_csv), file.exists(base_helper_file), file.exists(branch_helper_file))
source(base_helper_file)
source(branch_helper_file)
load_popcdm_sources()
```

# User-editable settings

```{r config}
beta_variant <- "base_nr_r"  # "down_up" | "base_nr_r"
scaling_mode <- "alpha_beta"      # "alpha_beta" | "kappa_beta" | "none"

# NULL fits all participants.
participant_ids <- "PG"
trial_cap_per_condition <- 450L
random_seed <- 1L

grid <- list(
  nw = 50L,
  h = 3.3 / 300,
  tmax = 3.3
)

optim_settings <- list(
  n_starts = 1L,
  beta_link = "logit",
  beta_latent_limit = 15,
  progress = TRUE,
  progress_every = 5L,
  control = list(maxit = 500L, factr = 1e9)
)

parallel_settings <- list(
  use_parallel = FALSE,
  n_cores = min(parallel::detectCores(logical = FALSE), 5L, na.rm = TRUE)
)

plot_settings <- list(
  error_n_bins = 180L,
  rt_n_bins = 200L
)
```

# Build model inputs

```{r model-spec}
P_start <- redundancy_default_P(
  beta_variant = beta_variant,
  alpha = 50.0,
  vnorm = 2.0,
  kappa = 5.0,
  a = 2.0,
  ter = 0.2,
  eta1 = 1.0,
  eta2 = 1.0,
  st = 0.2
)

bounds <- redundancy_default_bounds(beta_variant = beta_variant)

Sel <- redundancy_default_sel(
  beta_variant = beta_variant,
  free_eta2 = TRUE,
  free_st = FALSE
)

ci2_conditions <- c(
  "set2_baseline",
  "set4_R_cue",
  "set4_NR_cue",
  "set6_R_cue_ci2",
  "set6_NR_cue_ci2"
)
ci4_conditions <- c(
  "set4_baseline",
  "set6_R_cue_ci4",
  "set6_NR_cue_ci4"
)

# Current example: alpha-beta scaling with one shared base alpha,
# grouped eta2s, and one shared kappa.
m_scaling_branch <- redundancy_model_spec(
  label = paste0("scaling_branch_", scaling_mode, "_", beta_variant),
  beta_variant = beta_variant,
  scaling_mode = scaling_mode,
  color_count_ref = 2L,
  equal_groups = list(
    redundancy_equal_group("alpha", COND_LEVELS),
    redundancy_equal_group("kappa", COND_LEVELS),
    redundancy_equal_group("a", COND_LEVELS),
    redundancy_equal_group("eta1", COND_LEVELS),
    redundancy_equal_group("eta2", ci2_conditions),
    redundancy_equal_group("eta2", ci4_conditions),
    redundancy_equal_group("vnorm", ci2_conditions),
    redundancy_equal_group("vnorm", ci4_conditions),
    redundancy_equal_group("ter", ci2_conditions),
    redundancy_equal_group("ter", ci4_conditions)
  )
)

fit_inputs <- redundancy_fit_inputs(
  P_start = P_start,
  lower = bounds$lower,
  upper = bounds$upper,
  Sel = Sel,
  model_spec = m_scaling_branch
)

length(P_start)
length(fit_inputs$P_var_start)
names(fit_inputs$P_var_start)
names(fit_inputs$P_fix)
```

# Prepare and audit data

```{r data-audit} 
raw_ids <- sort(unique(utils::read.csv(data_csv, stringsAsFactors = FALSE)$uid))
uids_to_fit <- if (is.null(participant_ids)) raw_ids else participant_ids
uids_to_fit

fit_df_all <- prepare_redundancy_fit_frame(
  data_csv,
  uid_filter = uids_to_fit,
  drop_na_cond = TRUE
)

table(fit_df_all$uid, fit_df_all$cond_label)

redundancy_rt_grid_audit(
  fit_df_all,
  tmax = grid$tmax,
  h = grid$h,
  nw = grid$nw
)
```

# Fit selected participants

```{r fit, message=TRUE, warning=TRUE}
fit_one_participant <- function(uid) {
  message("Fitting participant: ", uid)
  fit_df_full <- prepare_redundancy_fit_frame(
    data_csv,
    uid_filter = uid,
    drop_na_cond = TRUE
  )
  fit_df <- if (is.null(trial_cap_per_condition)) {
    fit_df_full
  } else {
    redundancy_thin_trials_by_cond(
      fit_df_full,
      max_per_cond = trial_cap_per_condition,
      seed = random_seed
    )
  }

  fit_redundancy_popcdm_multistart(
    fit_df,
    P_var_start = fit_inputs$P_var_start,
    lower = fit_inputs$lower,
    upper = fit_inputs$upper,
    P_fix = fit_inputs$P_fix,
    Sel = fit_inputs$Sel,
    model_spec = fit_inputs$model_spec,
    n_starts = optim_settings$n_starts,
    seed = random_seed + match(uid, uids_to_fit),
    beta_link = optim_settings$beta_link,
    beta_latent_limit = optim_settings$beta_latent_limit,
    progress = optim_settings$progress,
    progress_every = optim_settings$progress_every,
    nw = grid$nw,
    h = grid$h,
    tmax = grid$tmax,
    control = optim_settings$control
  )
}

if (isTRUE(parallel_settings$use_parallel) && length(uids_to_fit) > 1L) {
  n_cores <- min(parallel_settings$n_cores, length(uids_to_fit))
  message("Fitting ", length(uids_to_fit), " participants using ", n_cores, " cores.")
  fits <- parallel::mclapply(
    uids_to_fit,
    fit_one_participant,
    mc.cores = n_cores,
    mc.set.seed = TRUE
  )
  names(fits) <- uids_to_fit
} else {
  fits <- lapply(uids_to_fit, fit_one_participant)
  names(fits) <- uids_to_fit
}

beep(4)
```

# Summaries

```{r summary}  
fit_summary <- do.call(
  rbind,
  lapply(uids_to_fit, function(uid) {
    f <- fits[[uid]]
    data.frame(
      uid = uid,
      model = f$model_spec$label,
      n_trials = f$n_trials,
      np_free = f$np_free,
      nll = f$nll,
      AIC = 2 * f$nll + 2 * f$np_free,
      BIC = 2 * f$nll + f$np_free * log(f$n_trials),
      convergence = f$optim$convergence,
      stringsAsFactors = FALSE
    )
  })
)
fit_summary

P_hats <- lapply(fits, `[[`, "P_hat")
P_hats[[uids_to_fit[1]]]
```

```{r full-parameter-table}
full_parameter_table <- do.call(
  rbind,
  lapply(uids_to_fit, function(uid) {
    P_hat <- P_hats[[uid]]
    data.frame(
      uid = uid,
      parameter = names(P_hat),
      estimate = as.numeric(P_hat),
      is_free = as.logical(fits[[uid]]$Sel[names(P_hat)]),
      stringsAsFactors = FALSE,
      row.names = NULL
    )
  })
)
full_parameter_table$estimate <- round(full_parameter_table$estimate, 4)
print(full_parameter_table)
```

```{r condition-parameter-table}
beta_cols <- if (identical(scaling_mode, "none")) {
  character(0)
} else {
  redundancy_beta_names(beta_variant)
}

condition_parameter_table <- do.call(
  rbind,
  lapply(uids_to_fit, function(uid) {
    P_hat <- P_hats[[uid]]
    do.call(
      rbind,
      lapply(COND_LEVELS, function(cl) {
        pred <- redundancy_predict_condition(
          P_hat,
          cl,
          nw = grid$nw,
          h = grid$h,
          tmax = grid$tmax,
          model_spec = fit_inputs$model_spec
        )
        row <- data.frame(
          uid = uid,
          cond_label = cl,
          alpha = P_hat[paste0("alpha_", cl)],
          kappa = P_hat[paste0("kappa_", cl)],
          alpha_trial = pred$alpha_trial,
          kappa_trial = pred$kappa_trial,
          vnorm = P_hat[paste0("vnorm_", cl)],
          a = P_hat[paste0("a_", cl)],
          ter = P_hat[paste0("ter_", cl)],
          eta1 = P_hat[paste0("eta1_", cl)],
          eta2 = P_hat[paste0("eta2_", cl)],
          stringsAsFactors = FALSE,
          row.names = NULL
        )
        for (bn in beta_cols) {
          row[[bn]] <- P_hat[bn]
        }
        row
      })
    )
  })
)
numeric_cols <- vapply(condition_parameter_table, is.numeric, logical(1))
condition_parameter_table[numeric_cols] <- lapply(
  condition_parameter_table[numeric_cols],
  round,
  digits = 2
)
condition_parameter_table
```

```{r diagnostic-helper-source}
source(base_helper_file)
source(branch_helper_file)
```

# Alpha-kappa curves against final model error marginals

These checks use fitted `alpha_trial` and `kappa_trial` to reconstruct the
population-code input layer for each condition, and compare it with the final
model-predicted error-angle marginal. The red curve is
`popcode(c(alpha_trial, kappa_trial))`; the blue dotted curve is the raw von
Mises density from `vm(kappa_trial)`; the black dashed curve is the final
response-error density from `popcdm300()` after the CDM layer and other
variability sources. The red/blue curves are therefore input-layer curves, while
the black curve is the model output that was previously plotted against data.

```{r plot-population-layers, fig.width=10, fig.height=8, results='asis'}
for (plot_uid in uids_to_fit) {
  cat("\n\n### Population Layer: ", plot_uid, "\n\n", sep = "")
  plot_redundancy_population_layers(
    P_hats[[plot_uid]],
    model_spec = fit_inputs$model_spec,
    nw = 360L,
    h = grid$h,
    tmax = grid$tmax
  )
}
```

```{r population-layer-cirsd-table}
population_cirsd_table <- do.call(
  rbind,
  lapply(uids_to_fit, function(uid) {
    tab <- redundancy_population_cirsd_table(
      P_hats[[uid]],
      model_spec = fit_inputs$model_spec,
      nw = 360L
    )
    tab$uid <- uid
    tab[, c("uid", setdiff(names(tab), "uid"))]
  })
)

numeric_cols <- vapply(population_cirsd_table, is.numeric, logical(1))
population_cirsd_table[numeric_cols] <- lapply(
  population_cirsd_table[numeric_cols],
  round,
  digits = 2
)
population_cirsd_table
```

```{r equivalent-vm-from-pang-table}
equivalent_vm_from_pang_table <- do.call(
  rbind,
  lapply(uids_to_fit, function(uid) {
    tab <- redundancy_population_equiv_vm_table(
      P_hats[[uid]],
      model_spec = fit_inputs$model_spec,
      nw = 360L
    )
    tab$uid <- uid
    tab[, c("uid", setdiff(names(tab), "uid"))]
  })
)

numeric_cols <- vapply(equivalent_vm_from_pang_table, is.numeric, logical(1))
equivalent_vm_from_pang_table[numeric_cols] <- lapply(
  equivalent_vm_from_pang_table[numeric_cols],
  round,
  digits = 4
)
equivalent_vm_from_pang_table
```

# Plot predictions against data

```{r plot-data-source}
# Helpers are sourced in `diagnostic-helper-source` above.
if (!exists("plot_settings")) {
  plot_settings <- list(
    error_n_bins = 360L,
    rt_n_bins = 100L
  )
}
```

```{r plot-error-marginals, fig.width=10, fig.height=8, results='asis'}
for (plot_uid in uids_to_fit) {
  cat("\n\n### Error Marginals: ", plot_uid, "\n\n", sep = "")
  plot_df <- prepare_redundancy_fit_frame(
    data_csv,
    uid_filter = plot_uid,
    drop_na_cond = TRUE
  )
  plot_redundancy_error_marginals(
    plot_df,
    P_hats[[plot_uid]],
    nw = grid$nw,
    h = grid$h,
    tmax = grid$tmax,
    n_bins = plot_settings$error_n_bins,
    model_spec = fit_inputs$model_spec
  )
}
```

```{r plot-rt-marginals, fig.width=10, fig.height=8, results='asis'}
for (plot_uid in uids_to_fit) {
  cat("\n\n### RT Marginals: ", plot_uid, "\n\n", sep = "")
  plot_df <- prepare_redundancy_fit_frame(
    data_csv,
    uid_filter = plot_uid,
    drop_na_cond = TRUE
  )
  plot_redundancy_rt_marginals(
    plot_df,
    P_hats[[plot_uid]],
    nw = grid$nw,
    h = grid$h,
    tmax = grid$tmax,
    n_bins = plot_settings$rt_n_bins,
    model_spec = fit_inputs$model_spec
  )
}
```
